### 2012-05-30 paper/demo.Rnw
started paper

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage[round]{natbib}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}
\usepackage{breakurl}
\begin{document}

<<setup, include=FALSE, cache=FALSE>>=
# this is equivalent to \SweaveOpts{...}
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', fig.show='hold', fig.lp='fig:')
options(replace.assign=TRUE,width=90)
@

\title{A Minimal Demo of glyphmaps}

\author{Garrett Grolemund}

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below. Each was made with \texttt{glyphmaps}. The first two plots recreate plots from \citet{wickham2011split}. The last two recreate plots from \citet{wickham2012glyph}.

<<plyr star glyphs>>=
@

<<plyr filled star glyphs>>=
@

<<glyph maps grided>>=
@

<<glyph maps merged>>=
@

To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with

<<>>=
library(devtools)
load_all("glyphmaps")
@

I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \Sexpr{install_github()}.

\section{Glyphs}
The above graphs are similar in that they are all built around glyphs. I define a glyph as a geom that visually summarises information about a subset of data. Glyphs can be quite simple, or very complex. Often glyphs contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

<<star-glyphs, fig.cap = 'Individual star glyphs comprise a complete plot with an internal coordinate system. Multiple star glyphs can be organized with an external coordinate sytem to reveal second order effects.'>>=
@

\section{glyphmaps geoms}
\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs. These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots. \texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.

<<geom_star>>=
@

<<geom_radar>>=
@

<<geom_scatterplots>>=
@

\subsection{new arguments}
\label{sec:args}
Each \texttt{glyphmaps} geom takes two new arguments, \Sexpr{glyphs} and \Sexpr{reference}. \Sexpr{glyphs} controls how the data will be divided into subsets to be paired with individual glyphs. \Sexpr{reference} lets the user draw reference features with the glyphs to facilitate comparisons.

\Sexpr{glyphs} takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs. The \Sexpr{glyphs} argument works just like a \Sexpr{.vars} argument in a \Sexpr{ddply} call. (In fact, the \Sexpr{glyphs} argument in a \texttt{glyphmaps} geom will be passed into a \Sexpr{ddply} call as the \Sexpr{.vars} argument.)

\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points. These objects help viewers make comparisons between data inside different glyphs.

<<ref-type-examples>>=
@

<<comparison-with-boxes>>=
@

\section{Generalized glyphs}
Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \Sexpr{glyph} and \Sexpr{plyr} functions.

<<bar-chart-glyph>>=
@

<<point-overplotting-ex>>=
@

The plyr function also allows users to key aesthetics to subgroups of data.

<<groupwise-aesthetics>>=
@

\section{Discussion - glyphs, embedded graphics, hierarchies}
Glyphs are geometric objects (i.e, geoms) designed to display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.

Glyphs reveal a hierarchical structure of graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot. Sometimes these subplots are not very interesting, as in the subplot created by a single point geom. At other times, thee subplots are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.

Graphs inherit this hierarchical structure from the data they describe. Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data. Humans innately perform this process when collecting data; it is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.

Glyph maps simultaneously expose data from multiple levels of the hierarchy. As a set of geoms, the glyphs reveal relationships between data points in the higher level, compact data set. As individual plots, each glyph retains information about the data points in the lower level group of data that it summarises. This dual display makes glyph maps particularly useful for certain data analysis tasks. It also provides two different approaches to constructing glyph maps.

Glyph maps can be built from the top down by treating each glyph as a geom within the plot of interest. \texttt{ggplyr} provides new geoms \texttt{geom_star}, \texttt{geom_radar}, \texttt{geom_dart}, and \texttt{geom_plyr} which allow a user to quickly incorporate geoms into an existing plot.

Alternatively, glyph maps can be built from the bottom up by treating each glyph as an existing plot and then combining these plots together with an external coordinate system. This approach parallels the split-apply-combine strategy of data analysis \citep{wickham2011split}. \texttt{ggplyr} provides a set of functions such as \texttt{ggply}, \texttt{dgply}, \texttt{grid}, and \texttt{nest} that help users use the familiar split-apply-combine process to create complex glyph maps. This approach will illustrate that any graph can be turned into a glyph to be used in a higher order graph.

The remainder of this paper will demonstrate how users can use \texttt{ggplyr} tools to build glyph maps from either the bottom up or the top down.

Although we have chosen to describe such plots as glyph maps, they really represent a general class of graphs that have been difficult to construct until now. Because each glyph is equivalent to an independent plot, glyph maps describe any plot that embeds other plots within itself. Embedded plots occur in a variety of contexts, and I have a stack of photocopies to help demonstrate this. Embedded plots also fit into the grammar of graphics, where they represent an extension of facetting. When plots are embedded into two categorical axes, a facetted plot results. When a graph is facetted on continuous axes (which is rarely done), a glyph map results.

(Interesetingly most embedded plots have settled to the stage of embedding plots within a geographic map. This resembles how Wainer, Tufte and others have traced the development of graphs through a concrete stage where cartesian axes seldom refered to anything other than latitude and longitude. Eventually graphs became more sophisticated and axes were used to represent  abstract concepts. Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one order to have a concrete interpretation.)

\section{top down methods}
\texttt{ggplyr}'s top down methods help users fit glyphs into existing methods of visualization. These methods provide new geoms (often based on new grobs) that can be used alongside existing geoms in \texttt{ggplot2}.

Top down methods are difficult to implement in \texttt{ggplot2} for two reasons
\begin{ennumerate}
  \item First, \texttt{ggplot2} calculates aesthetics on the entire data set at once. However, glyphs contain aesthetics that must be keyed to subsets of the data.
  \item Second, the final width and height of individual glyphs often depends on non-position aesthetics, such as angle and length. In the \texttt{ggplot2} pipeline, these aesthetics are scaled right before the plot ranges are trained. As a result, the final widths of glyphs must be computed at draw time and frequently place parts of the glyph outside the plot window.
\end{ennumerate}

\subsection{geom_plyr}
\texttt{ggplyr} provides a way for users to compute aesthetics by

<<boring-random>>=
set.seed(1121)
(x=rnorm(20))
mean(x);var(x)
@

The first element of \texttt{x} is \Sexpr{x[1]}. Boring boxplots
and histograms recorded by the PDF device:

<<boring-plots, fig.width=4, fig.height=4, out.width='.4\\linewidth'>>=
## two plots side by side (option fig.show='hold')
par(mar=c(4,4,.1,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3,las=1)
boxplot(x)
hist(x,main='')
@

Do the above chunks work? You should be able to compile the \TeX{}
document and get a PDF file like this one: \url{https://github.com/downloads/yihui/knitr/knitr-minimal.pdf}.
The Rnw source of this document is at \url{https://github.com/yihui/knitr/blob/master/inst/examples/knitr-minimal.Rnw}.

\bibliographystyle{plainnat}
\bibliography{ggplyr}
\end{document}

### 2012-05-30 paper/demo.tex
started paper

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage{graphicx, color}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\definecolor{fgcolor}{rgb}{0.267, 0.267, 0.267}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.266666666666667,0.266666666666667,0.266666666666667}{#1}}%

\usepackage{color}%

\newsavebox{\hlnormalsizeboxclosebrace}%
\newsavebox{\hlnormalsizeboxopenbrace}%
\newsavebox{\hlnormalsizeboxbackslash}%
\newsavebox{\hlnormalsizeboxlessthan}%
\newsavebox{\hlnormalsizeboxgreaterthan}%
\newsavebox{\hlnormalsizeboxdollar}%
\newsavebox{\hlnormalsizeboxunderscore}%
\newsavebox{\hlnormalsizeboxand}%
\newsavebox{\hlnormalsizeboxhash}%
\newsavebox{\hlnormalsizeboxat}%
\newsavebox{\hlnormalsizeboxpercent}% 
\newsavebox{\hlnormalsizeboxhat}%
\newsavebox{\hlnormalsizeboxsinglequote}%
\newsavebox{\hlnormalsizeboxbacktick}%

\setbox\hlnormalsizeboxopenbrace=\hbox{\begin{normalsize}\verb.{.\end{normalsize}}%
\setbox\hlnormalsizeboxclosebrace=\hbox{\begin{normalsize}\verb.}.\end{normalsize}}%
\setbox\hlnormalsizeboxlessthan=\hbox{\begin{normalsize}\verb.<.\end{normalsize}}%
\setbox\hlnormalsizeboxdollar=\hbox{\begin{normalsize}\verb.$.\end{normalsize}}%
\setbox\hlnormalsizeboxunderscore=\hbox{\begin{normalsize}\verb._.\end{normalsize}}%
\setbox\hlnormalsizeboxand=\hbox{\begin{normalsize}\verb.&.\end{normalsize}}%
\setbox\hlnormalsizeboxhash=\hbox{\begin{normalsize}\verb.#.\end{normalsize}}%
\setbox\hlnormalsizeboxat=\hbox{\begin{normalsize}\verb.@.\end{normalsize}}%
\setbox\hlnormalsizeboxbackslash=\hbox{\begin{normalsize}\verb.\.\end{normalsize}}%
\setbox\hlnormalsizeboxgreaterthan=\hbox{\begin{normalsize}\verb.>.\end{normalsize}}%
\setbox\hlnormalsizeboxpercent=\hbox{\begin{normalsize}\verb.%.\end{normalsize}}%
\setbox\hlnormalsizeboxhat=\hbox{\begin{normalsize}\verb.^.\end{normalsize}}%
\setbox\hlnormalsizeboxsinglequote=\hbox{\begin{normalsize}\verb.'.\end{normalsize}}%
\setbox\hlnormalsizeboxbacktick=\hbox{\begin{normalsize}\verb.`.\end{normalsize}}%
\setbox\hlnormalsizeboxhat=\hbox{\begin{normalsize}\verb.^.\end{normalsize}}%

\newsavebox{\hltinyboxclosebrace}%
\newsavebox{\hltinyboxopenbrace}%
\newsavebox{\hltinyboxbackslash}%
\newsavebox{\hltinyboxlessthan}%
\newsavebox{\hltinyboxgreaterthan}%
\newsavebox{\hltinyboxdollar}%
\newsavebox{\hltinyboxunderscore}%
\newsavebox{\hltinyboxand}%
\newsavebox{\hltinyboxhash}%
\newsavebox{\hltinyboxat}%
\newsavebox{\hltinyboxpercent}% 
\newsavebox{\hltinyboxhat}%
\newsavebox{\hltinyboxsinglequote}%
\newsavebox{\hltinyboxbacktick}%

\setbox\hltinyboxopenbrace=\hbox{\begin{tiny}\verb.{.\end{tiny}}%
\setbox\hltinyboxclosebrace=\hbox{\begin{tiny}\verb.}.\end{tiny}}%
\setbox\hltinyboxlessthan=\hbox{\begin{tiny}\verb.<.\end{tiny}}%
\setbox\hltinyboxdollar=\hbox{\begin{tiny}\verb.$.\end{tiny}}%
\setbox\hltinyboxunderscore=\hbox{\begin{tiny}\verb._.\end{tiny}}%
\setbox\hltinyboxand=\hbox{\begin{tiny}\verb.&.\end{tiny}}%
\setbox\hltinyboxhash=\hbox{\begin{tiny}\verb.#.\end{tiny}}%
\setbox\hltinyboxat=\hbox{\begin{tiny}\verb.@.\end{tiny}}%
\setbox\hltinyboxbackslash=\hbox{\begin{tiny}\verb.\.\end{tiny}}%
\setbox\hltinyboxgreaterthan=\hbox{\begin{tiny}\verb.>.\end{tiny}}%
\setbox\hltinyboxpercent=\hbox{\begin{tiny}\verb.%.\end{tiny}}%
\setbox\hltinyboxhat=\hbox{\begin{tiny}\verb.^.\end{tiny}}%
\setbox\hltinyboxsinglequote=\hbox{\begin{tiny}\verb.'.\end{tiny}}%
\setbox\hltinyboxbacktick=\hbox{\begin{tiny}\verb.`.\end{tiny}}%
\setbox\hltinyboxhat=\hbox{\begin{tiny}\verb.^.\end{tiny}}%

\newsavebox{\hlscriptsizeboxclosebrace}%
\newsavebox{\hlscriptsizeboxopenbrace}%
\newsavebox{\hlscriptsizeboxbackslash}%
\newsavebox{\hlscriptsizeboxlessthan}%
\newsavebox{\hlscriptsizeboxgreaterthan}%
\newsavebox{\hlscriptsizeboxdollar}%
\newsavebox{\hlscriptsizeboxunderscore}%
\newsavebox{\hlscriptsizeboxand}%
\newsavebox{\hlscriptsizeboxhash}%
\newsavebox{\hlscriptsizeboxat}%
\newsavebox{\hlscriptsizeboxpercent}% 
\newsavebox{\hlscriptsizeboxhat}%
\newsavebox{\hlscriptsizeboxsinglequote}%
\newsavebox{\hlscriptsizeboxbacktick}%

\setbox\hlscriptsizeboxopenbrace=\hbox{\begin{scriptsize}\verb.{.\end{scriptsize}}%
\setbox\hlscriptsizeboxclosebrace=\hbox{\begin{scriptsize}\verb.}.\end{scriptsize}}%
\setbox\hlscriptsizeboxlessthan=\hbox{\begin{scriptsize}\verb.<.\end{scriptsize}}%
\setbox\hlscriptsizeboxdollar=\hbox{\begin{scriptsize}\verb.$.\end{scriptsize}}%
\setbox\hlscriptsizeboxunderscore=\hbox{\begin{scriptsize}\verb._.\end{scriptsize}}%
\setbox\hlscriptsizeboxand=\hbox{\begin{scriptsize}\verb.&.\end{scriptsize}}%
\setbox\hlscriptsizeboxhash=\hbox{\begin{scriptsize}\verb.#.\end{scriptsize}}%
\setbox\hlscriptsizeboxat=\hbox{\begin{scriptsize}\verb.@.\end{scriptsize}}%
\setbox\hlscriptsizeboxbackslash=\hbox{\begin{scriptsize}\verb.\.\end{scriptsize}}%
\setbox\hlscriptsizeboxgreaterthan=\hbox{\begin{scriptsize}\verb.>.\end{scriptsize}}%
\setbox\hlscriptsizeboxpercent=\hbox{\begin{scriptsize}\verb.%.\end{scriptsize}}%
\setbox\hlscriptsizeboxhat=\hbox{\begin{scriptsize}\verb.^.\end{scriptsize}}%
\setbox\hlscriptsizeboxsinglequote=\hbox{\begin{scriptsize}\verb.'.\end{scriptsize}}%
\setbox\hlscriptsizeboxbacktick=\hbox{\begin{scriptsize}\verb.`.\end{scriptsize}}%
\setbox\hlscriptsizeboxhat=\hbox{\begin{scriptsize}\verb.^.\end{scriptsize}}%

\newsavebox{\hlfootnotesizeboxclosebrace}%
\newsavebox{\hlfootnotesizeboxopenbrace}%
\newsavebox{\hlfootnotesizeboxbackslash}%
\newsavebox{\hlfootnotesizeboxlessthan}%
\newsavebox{\hlfootnotesizeboxgreaterthan}%
\newsavebox{\hlfootnotesizeboxdollar}%
\newsavebox{\hlfootnotesizeboxunderscore}%
\newsavebox{\hlfootnotesizeboxand}%
\newsavebox{\hlfootnotesizeboxhash}%
\newsavebox{\hlfootnotesizeboxat}%
\newsavebox{\hlfootnotesizeboxpercent}% 
\newsavebox{\hlfootnotesizeboxhat}%
\newsavebox{\hlfootnotesizeboxsinglequote}%
\newsavebox{\hlfootnotesizeboxbacktick}%

\setbox\hlfootnotesizeboxopenbrace=\hbox{\begin{footnotesize}\verb.{.\end{footnotesize}}%
\setbox\hlfootnotesizeboxclosebrace=\hbox{\begin{footnotesize}\verb.}.\end{footnotesize}}%
\setbox\hlfootnotesizeboxlessthan=\hbox{\begin{footnotesize}\verb.<.\end{footnotesize}}%
\setbox\hlfootnotesizeboxdollar=\hbox{\begin{footnotesize}\verb.$.\end{footnotesize}}%
\setbox\hlfootnotesizeboxunderscore=\hbox{\begin{footnotesize}\verb._.\end{footnotesize}}%
\setbox\hlfootnotesizeboxand=\hbox{\begin{footnotesize}\verb.&.\end{footnotesize}}%
\setbox\hlfootnotesizeboxhash=\hbox{\begin{footnotesize}\verb.#.\end{footnotesize}}%
\setbox\hlfootnotesizeboxat=\hbox{\begin{footnotesize}\verb.@.\end{footnotesize}}%
\setbox\hlfootnotesizeboxbackslash=\hbox{\begin{footnotesize}\verb.\.\end{footnotesize}}%
\setbox\hlfootnotesizeboxgreaterthan=\hbox{\begin{footnotesize}\verb.>.\end{footnotesize}}%
\setbox\hlfootnotesizeboxpercent=\hbox{\begin{footnotesize}\verb.%.\end{footnotesize}}%
\setbox\hlfootnotesizeboxhat=\hbox{\begin{footnotesize}\verb.^.\end{footnotesize}}%
\setbox\hlfootnotesizeboxsinglequote=\hbox{\begin{footnotesize}\verb.'.\end{footnotesize}}%
\setbox\hlfootnotesizeboxbacktick=\hbox{\begin{footnotesize}\verb.`.\end{footnotesize}}%
\setbox\hlfootnotesizeboxhat=\hbox{\begin{footnotesize}\verb.^.\end{footnotesize}}%

\newsavebox{\hlsmallboxclosebrace}%
\newsavebox{\hlsmallboxopenbrace}%
\newsavebox{\hlsmallboxbackslash}%
\newsavebox{\hlsmallboxlessthan}%
\newsavebox{\hlsmallboxgreaterthan}%
\newsavebox{\hlsmallboxdollar}%
\newsavebox{\hlsmallboxunderscore}%
\newsavebox{\hlsmallboxand}%
\newsavebox{\hlsmallboxhash}%
\newsavebox{\hlsmallboxat}%
\newsavebox{\hlsmallboxpercent}% 
\newsavebox{\hlsmallboxhat}%
\newsavebox{\hlsmallboxsinglequote}%
\newsavebox{\hlsmallboxbacktick}%

\setbox\hlsmallboxopenbrace=\hbox{\begin{small}\verb.{.\end{small}}%
\setbox\hlsmallboxclosebrace=\hbox{\begin{small}\verb.}.\end{small}}%
\setbox\hlsmallboxlessthan=\hbox{\begin{small}\verb.<.\end{small}}%
\setbox\hlsmallboxdollar=\hbox{\begin{small}\verb.$.\end{small}}%
\setbox\hlsmallboxunderscore=\hbox{\begin{small}\verb._.\end{small}}%
\setbox\hlsmallboxand=\hbox{\begin{small}\verb.&.\end{small}}%
\setbox\hlsmallboxhash=\hbox{\begin{small}\verb.#.\end{small}}%
\setbox\hlsmallboxat=\hbox{\begin{small}\verb.@.\end{small}}%
\setbox\hlsmallboxbackslash=\hbox{\begin{small}\verb.\.\end{small}}%
\setbox\hlsmallboxgreaterthan=\hbox{\begin{small}\verb.>.\end{small}}%
\setbox\hlsmallboxpercent=\hbox{\begin{small}\verb.%.\end{small}}%
\setbox\hlsmallboxhat=\hbox{\begin{small}\verb.^.\end{small}}%
\setbox\hlsmallboxsinglequote=\hbox{\begin{small}\verb.'.\end{small}}%
\setbox\hlsmallboxbacktick=\hbox{\begin{small}\verb.`.\end{small}}%
\setbox\hlsmallboxhat=\hbox{\begin{small}\verb.^.\end{small}}%

\newsavebox{\hllargeboxclosebrace}%
\newsavebox{\hllargeboxopenbrace}%
\newsavebox{\hllargeboxbackslash}%
\newsavebox{\hllargeboxlessthan}%
\newsavebox{\hllargeboxgreaterthan}%
\newsavebox{\hllargeboxdollar}%
\newsavebox{\hllargeboxunderscore}%
\newsavebox{\hllargeboxand}%
\newsavebox{\hllargeboxhash}%
\newsavebox{\hllargeboxat}%
\newsavebox{\hllargeboxpercent}% 
\newsavebox{\hllargeboxhat}%
\newsavebox{\hllargeboxsinglequote}%
\newsavebox{\hllargeboxbacktick}%

\setbox\hllargeboxopenbrace=\hbox{\begin{large}\verb.{.\end{large}}%
\setbox\hllargeboxclosebrace=\hbox{\begin{large}\verb.}.\end{large}}%
\setbox\hllargeboxlessthan=\hbox{\begin{large}\verb.<.\end{large}}%
\setbox\hllargeboxdollar=\hbox{\begin{large}\verb.$.\end{large}}%
\setbox\hllargeboxunderscore=\hbox{\begin{large}\verb._.\end{large}}%
\setbox\hllargeboxand=\hbox{\begin{large}\verb.&.\end{large}}%
\setbox\hllargeboxhash=\hbox{\begin{large}\verb.#.\end{large}}%
\setbox\hllargeboxat=\hbox{\begin{large}\verb.@.\end{large}}%
\setbox\hllargeboxbackslash=\hbox{\begin{large}\verb.\.\end{large}}%
\setbox\hllargeboxgreaterthan=\hbox{\begin{large}\verb.>.\end{large}}%
\setbox\hllargeboxpercent=\hbox{\begin{large}\verb.%.\end{large}}%
\setbox\hllargeboxhat=\hbox{\begin{large}\verb.^.\end{large}}%
\setbox\hllargeboxsinglequote=\hbox{\begin{large}\verb.'.\end{large}}%
\setbox\hllargeboxbacktick=\hbox{\begin{large}\verb.`.\end{large}}%
\setbox\hllargeboxhat=\hbox{\begin{large}\verb.^.\end{large}}%

\newsavebox{\hlLargeboxclosebrace}%
\newsavebox{\hlLargeboxopenbrace}%
\newsavebox{\hlLargeboxbackslash}%
\newsavebox{\hlLargeboxlessthan}%
\newsavebox{\hlLargeboxgreaterthan}%
\newsavebox{\hlLargeboxdollar}%
\newsavebox{\hlLargeboxunderscore}%
\newsavebox{\hlLargeboxand}%
\newsavebox{\hlLargeboxhash}%
\newsavebox{\hlLargeboxat}%
\newsavebox{\hlLargeboxpercent}% 
\newsavebox{\hlLargeboxhat}%
\newsavebox{\hlLargeboxsinglequote}%
\newsavebox{\hlLargeboxbacktick}%

\setbox\hlLargeboxopenbrace=\hbox{\begin{Large}\verb.{.\end{Large}}%
\setbox\hlLargeboxclosebrace=\hbox{\begin{Large}\verb.}.\end{Large}}%
\setbox\hlLargeboxlessthan=\hbox{\begin{Large}\verb.<.\end{Large}}%
\setbox\hlLargeboxdollar=\hbox{\begin{Large}\verb.$.\end{Large}}%
\setbox\hlLargeboxunderscore=\hbox{\begin{Large}\verb._.\end{Large}}%
\setbox\hlLargeboxand=\hbox{\begin{Large}\verb.&.\end{Large}}%
\setbox\hlLargeboxhash=\hbox{\begin{Large}\verb.#.\end{Large}}%
\setbox\hlLargeboxat=\hbox{\begin{Large}\verb.@.\end{Large}}%
\setbox\hlLargeboxbackslash=\hbox{\begin{Large}\verb.\.\end{Large}}%
\setbox\hlLargeboxgreaterthan=\hbox{\begin{Large}\verb.>.\end{Large}}%
\setbox\hlLargeboxpercent=\hbox{\begin{Large}\verb.%.\end{Large}}%
\setbox\hlLargeboxhat=\hbox{\begin{Large}\verb.^.\end{Large}}%
\setbox\hlLargeboxsinglequote=\hbox{\begin{Large}\verb.'.\end{Large}}%
\setbox\hlLargeboxbacktick=\hbox{\begin{Large}\verb.`.\end{Large}}%
\setbox\hlLargeboxhat=\hbox{\begin{Large}\verb.^.\end{Large}}%

\newsavebox{\hlLARGEboxclosebrace}%
\newsavebox{\hlLARGEboxopenbrace}%
\newsavebox{\hlLARGEboxbackslash}%
\newsavebox{\hlLARGEboxlessthan}%
\newsavebox{\hlLARGEboxgreaterthan}%
\newsavebox{\hlLARGEboxdollar}%
\newsavebox{\hlLARGEboxunderscore}%
\newsavebox{\hlLARGEboxand}%
\newsavebox{\hlLARGEboxhash}%
\newsavebox{\hlLARGEboxat}%
\newsavebox{\hlLARGEboxpercent}% 
\newsavebox{\hlLARGEboxhat}%
\newsavebox{\hlLARGEboxsinglequote}%
\newsavebox{\hlLARGEboxbacktick}%

\setbox\hlLARGEboxopenbrace=\hbox{\begin{LARGE}\verb.{.\end{LARGE}}%
\setbox\hlLARGEboxclosebrace=\hbox{\begin{LARGE}\verb.}.\end{LARGE}}%
\setbox\hlLARGEboxlessthan=\hbox{\begin{LARGE}\verb.<.\end{LARGE}}%
\setbox\hlLARGEboxdollar=\hbox{\begin{LARGE}\verb.$.\end{LARGE}}%
\setbox\hlLARGEboxunderscore=\hbox{\begin{LARGE}\verb._.\end{LARGE}}%
\setbox\hlLARGEboxand=\hbox{\begin{LARGE}\verb.&.\end{LARGE}}%
\setbox\hlLARGEboxhash=\hbox{\begin{LARGE}\verb.#.\end{LARGE}}%
\setbox\hlLARGEboxat=\hbox{\begin{LARGE}\verb.@.\end{LARGE}}%
\setbox\hlLARGEboxbackslash=\hbox{\begin{LARGE}\verb.\.\end{LARGE}}%
\setbox\hlLARGEboxgreaterthan=\hbox{\begin{LARGE}\verb.>.\end{LARGE}}%
\setbox\hlLARGEboxpercent=\hbox{\begin{LARGE}\verb.%.\end{LARGE}}%
\setbox\hlLARGEboxhat=\hbox{\begin{LARGE}\verb.^.\end{LARGE}}%
\setbox\hlLARGEboxsinglequote=\hbox{\begin{LARGE}\verb.'.\end{LARGE}}%
\setbox\hlLARGEboxbacktick=\hbox{\begin{LARGE}\verb.`.\end{LARGE}}%
\setbox\hlLARGEboxhat=\hbox{\begin{LARGE}\verb.^.\end{LARGE}}%

\newsavebox{\hlhugeboxclosebrace}%
\newsavebox{\hlhugeboxopenbrace}%
\newsavebox{\hlhugeboxbackslash}%
\newsavebox{\hlhugeboxlessthan}%
\newsavebox{\hlhugeboxgreaterthan}%
\newsavebox{\hlhugeboxdollar}%
\newsavebox{\hlhugeboxunderscore}%
\newsavebox{\hlhugeboxand}%
\newsavebox{\hlhugeboxhash}%
\newsavebox{\hlhugeboxat}%
\newsavebox{\hlhugeboxpercent}% 
\newsavebox{\hlhugeboxhat}%
\newsavebox{\hlhugeboxsinglequote}%
\newsavebox{\hlhugeboxbacktick}%

\setbox\hlhugeboxopenbrace=\hbox{\begin{huge}\verb.{.\end{huge}}%
\setbox\hlhugeboxclosebrace=\hbox{\begin{huge}\verb.}.\end{huge}}%
\setbox\hlhugeboxlessthan=\hbox{\begin{huge}\verb.<.\end{huge}}%
\setbox\hlhugeboxdollar=\hbox{\begin{huge}\verb.$.\end{huge}}%
\setbox\hlhugeboxunderscore=\hbox{\begin{huge}\verb._.\end{huge}}%
\setbox\hlhugeboxand=\hbox{\begin{huge}\verb.&.\end{huge}}%
\setbox\hlhugeboxhash=\hbox{\begin{huge}\verb.#.\end{huge}}%
\setbox\hlhugeboxat=\hbox{\begin{huge}\verb.@.\end{huge}}%
\setbox\hlhugeboxbackslash=\hbox{\begin{huge}\verb.\.\end{huge}}%
\setbox\hlhugeboxgreaterthan=\hbox{\begin{huge}\verb.>.\end{huge}}%
\setbox\hlhugeboxpercent=\hbox{\begin{huge}\verb.%.\end{huge}}%
\setbox\hlhugeboxhat=\hbox{\begin{huge}\verb.^.\end{huge}}%
\setbox\hlhugeboxsinglequote=\hbox{\begin{huge}\verb.'.\end{huge}}%
\setbox\hlhugeboxbacktick=\hbox{\begin{huge}\verb.`.\end{huge}}%
\setbox\hlhugeboxhat=\hbox{\begin{huge}\verb.^.\end{huge}}%

\newsavebox{\hlHugeboxclosebrace}%
\newsavebox{\hlHugeboxopenbrace}%
\newsavebox{\hlHugeboxbackslash}%
\newsavebox{\hlHugeboxlessthan}%
\newsavebox{\hlHugeboxgreaterthan}%
\newsavebox{\hlHugeboxdollar}%
\newsavebox{\hlHugeboxunderscore}%
\newsavebox{\hlHugeboxand}%
\newsavebox{\hlHugeboxhash}%
\newsavebox{\hlHugeboxat}%
\newsavebox{\hlHugeboxpercent}% 
\newsavebox{\hlHugeboxhat}%
\newsavebox{\hlHugeboxsinglequote}%
\newsavebox{\hlHugeboxbacktick}%

\setbox\hlHugeboxopenbrace=\hbox{\begin{Huge}\verb.{.\end{Huge}}%
\setbox\hlHugeboxclosebrace=\hbox{\begin{Huge}\verb.}.\end{Huge}}%
\setbox\hlHugeboxlessthan=\hbox{\begin{Huge}\verb.<.\end{Huge}}%
\setbox\hlHugeboxdollar=\hbox{\begin{Huge}\verb.$.\end{Huge}}%
\setbox\hlHugeboxunderscore=\hbox{\begin{Huge}\verb._.\end{Huge}}%
\setbox\hlHugeboxand=\hbox{\begin{Huge}\verb.&.\end{Huge}}%
\setbox\hlHugeboxhash=\hbox{\begin{Huge}\verb.#.\end{Huge}}%
\setbox\hlHugeboxat=\hbox{\begin{Huge}\verb.@.\end{Huge}}%
\setbox\hlHugeboxbackslash=\hbox{\begin{Huge}\verb.\.\end{Huge}}%
\setbox\hlHugeboxgreaterthan=\hbox{\begin{Huge}\verb.>.\end{Huge}}%
\setbox\hlHugeboxpercent=\hbox{\begin{Huge}\verb.%.\end{Huge}}%
\setbox\hlHugeboxhat=\hbox{\begin{Huge}\verb.^.\end{Huge}}%
\setbox\hlHugeboxsinglequote=\hbox{\begin{Huge}\verb.'.\end{Huge}}%
\setbox\hlHugeboxbacktick=\hbox{\begin{Huge}\verb.`.\end{Huge}}%
\setbox\hlHugeboxhat=\hbox{\begin{Huge}\verb.^.\end{Huge}}%

\def\urltilda{\kern -.15em\lower .7ex\hbox{\~{}}\kern .04em}%

\newcommand{\hlstd}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.16,0.16,1}{#1}}
\newcommand{\hlesc}[1]{\textcolor[rgb]{1,0,1}{#1}}
\newcommand{\hlstr}[1]{\textcolor[rgb]{1,0,0}{#1}}
\newcommand{\hldstr}[1]{\textcolor[rgb]{0.51,0.51,0}{#1}}
\newcommand{\hlslc}[1]{\textcolor[rgb]{0.51,0.51,0.51}{\it{#1}}}
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.51,0.51,0.51}{\it{#1}}}
\newcommand{\hldir}[1]{\textcolor[rgb]{0,0.51,0}{#1}}
\newcommand{\hlsym}[1]{\textcolor[rgb]{0,0,0}{#1}}
\newcommand{\hlline}[1]{\textcolor[rgb]{0.33,0.33,0.33}{#1}}
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0,0,0}{\bf{#1}}}
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.51,0,0}{#1}}
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0,0,0}{\bf{#1}}}
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0,0,0.51}{#1}}

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage[round]{natbib}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}
\usepackage{breakurl}
\begin{document}

\title{A Minimal Demo of glyphmaps}

\author{Garrett Grolemund}

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below. Each was made with \texttt{glyphmaps}. The first two plots recreate plots from \citet{wickham2011split}. The last two recreate plots from \citet{wickham2012glyph}.

To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{load\usebox{\hlnormalsizeboxunderscore}all}\hlkeyword{(}\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## Error: Can't find package glyphmaps
\end{verbatim}
\end{kframe}
\end{knitrout}

I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \begin{verbatim}

Error in paste(repo, branch, sep = "/", collapse = ", ") : 
  argument "repo" is missing, with no default

\end{verbatim}
.

\section{Glyphs}
The above graphs are similar in that they are all built around glyphs. I define a glyph as a geom that visually summarises information about a subset of data. Glyphs can be quite simple, or very complex. Often glyphs contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

\section{glyphmaps geoms}
\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs. These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots. \texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.

\subsection{new arguments}
\label{sec:args}
Each \texttt{glyphmaps} geom takes two new arguments, \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 and \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'reference' not found

\end{verbatim}
. \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 controls how the data will be divided into subsets to be paired with individual glyphs. \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'reference' not found

\end{verbatim}
 lets the user draw reference features with the glyphs to facilitate comparisons.

\begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs. The \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 argument works just like a \begin{verbatim}

Error in eval(expr, envir, enclos) : object '.vars' not found

\end{verbatim}
 argument in a \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'ddply' not found

\end{verbatim}
 call. (In fact, the \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyphs' not found

\end{verbatim}
 argument in a \texttt{glyphmaps} geom will be passed into a \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'ddply' not found

\end{verbatim}
 call as the \begin{verbatim}

Error in eval(expr, envir, enclos) : object '.vars' not found

\end{verbatim}
 argument.)

\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points. These objects help viewers make comparisons between data inside different glyphs.

\section{Generalized glyphs}
Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'glyph' not found

\end{verbatim}
 and \begin{verbatim}

Error in eval(expr, envir, enclos) : object 'plyr' not found

\end{verbatim}
 functions.

The plyr function also allows users to key aesthetics to subgroups of data.

\section{Discussion - glyphs, embedded graphics, hierarchies}
Glyphs are geometric objects (i.e, geoms) designed to display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.

Glyphs reveal a hierarchical structure of graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot. Sometimes these subplots are not very interesting, as in the subplot created by a single point geom. At other times, thee subplots are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.

Graphs inherit this hierarchical structure from the data they describe. Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data. Humans innately perform this process when collecting data; it is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.

Glyph maps simultaneously expose data from multiple levels of the hierarchy. As a set of geoms, the glyphs reveal relationships between data points in the higher level, compact data set. As individual plots, each glyph retains information about the data points in the lower level group of data that it summarises. This dual display makes glyph maps particularly useful for certain data analysis tasks. It also provides two different approaches to constructing glyph maps.

Glyph maps can be built from the top down by treating each glyph as a geom within the plot of interest. \texttt{ggplyr} provides new geoms \texttt{geom_star}, \texttt{geom_radar}, \texttt{geom_dart}, and \texttt{geom_plyr} which allow a user to quickly incorporate geoms into an existing plot.

Alternatively, glyph maps can be built from the bottom up by treating each glyph as an existing plot and then combining these plots together with an external coordinate system. This approach parallels the split-apply-combine strategy of data analysis \citep{wickham2011split}. \texttt{ggplyr} provides a set of functions such as \texttt{ggply}, \texttt{dgply}, \texttt{grid}, and \texttt{nest} that help users use the familiar split-apply-combine process to create complex glyph maps. This approach will illustrate that any graph can be turned into a glyph to be used in a higher order graph.

The remainder of this paper will demonstrate how users can use \texttt{ggplyr} tools to build glyph maps from either the bottom up or the top down.

Although we have chosen to describe such plots as glyph maps, they really represent a general class of graphs that have been difficult to construct until now. Because each glyph is equivalent to an independent plot, glyph maps describe any plot that embeds other plots within itself. Embedded plots occur in a variety of contexts, and I have a stack of photocopies to help demonstrate this. Embedded plots also fit into the grammar of graphics, where they represent an extension of facetting. When plots are embedded into two categorical axes, a facetted plot results. When a graph is facetted on continuous axes (which is rarely done), a glyph map results.

(Interesetingly most embedded plots have settled to the stage of embedding plots within a geographic map. This resembles how Wainer, Tufte and others have traced the development of graphs through a concrete stage where cartesian axes seldom refered to anything other than latitude and longitude. Eventually graphs became more sophisticated and axes were used to represent  abstract concepts. Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one order to have a concrete interpretation.)

\section{top down methods}
\texttt{ggplyr}'s top down methods help users fit glyphs into existing methods of visualization. These methods provide new geoms (often based on new grobs) that can be used alongside existing geoms in \texttt{ggplot2}.

Top down methods are difficult to implement in \texttt{ggplot2} for two reasons
\begin{ennumerate}
  \item First, \texttt{ggplot2} calculates aesthetics on the entire data set at once. However, glyphs contain aesthetics that must be keyed to subsets of the data.
  \item Second, the final width and height of individual glyphs often depends on non-position aesthetics, such as angle and length. In the \texttt{ggplot2} pipeline, these aesthetics are scaled right before the plot ranges are trained. As a result, the final widths of glyphs must be computed at draw time and frequently place parts of the glyph outside the plot window.
\end{ennumerate}

\subsection{geom_plyr}
\texttt{ggplyr} provides a way for users to compute aesthetics by

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{set.seed}\hlkeyword{(}\hlnumber{1121}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlkeyword{(}\hlsymbol{x}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{rnorm}\hlkeyword{(}\hlnumber{20}\hlkeyword{)}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
##  [1]  0.14496  0.43832  0.15319  1.08494  1.99954 -0.81188  0.16027  0.58589  0.36009
## [10] -0.02531  0.15088  0.11008  1.35968 -0.32699 -0.71638  1.80977  0.50840 -0.52746
## [19]  0.13272 -0.15594
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.3217
\end{verbatim}
\begin{flushleft}
\ttfamily\noindent
\hlfunctioncall{var}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\begin{verbatim}
## [1] 0.5715
\end{verbatim}
\end{kframe}
\end{knitrout}

The first element of \texttt{x} is 0.145. Boring boxplots
and histograms recorded by the PDF device:

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{flushleft}
\ttfamily\noindent
\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }two{\ }plots{\ }side{\ }by{\ }side{\ }(option{\ }fig.show=\usebox{\hlnormalsizeboxsinglequote}hold\usebox{\hlnormalsizeboxsinglequote})}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{par}\hlkeyword{(}\hlargument{mar}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{cex.lab}{\ }\hlargument{=}{\ }\hlnumber{0.95}\hlkeyword{,}{\ }\hlargument{cex.axis}{\ }\hlargument{=}{\ }\hlnumber{0.9}\hlkeyword{,}{\ }\hlargument{mgp}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{2}\hlkeyword{,}{\ }\hlnumber{0.7}\hlkeyword{,}{\ }\hlnumber{0}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
\hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\hspace*{\fill}\\
\hlstd{}\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
\normalfont
\end{flushleft}
\end{kframe}

{\centering \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots1} \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots2}

}

\end{knitrout}

Do the above chunks work? You should be able to compile the \TeX{}
document and get a PDF file like this one: \url{https://github.com/downloads/yihui/knitr/knitr-minimal.pdf}.
The Rnw source of this document is at \url{https://github.com/yihui/knitr/blob/master/inst/examples/knitr-minimal.Rnw}.

\bibliographystyle{plainnat}
\bibliography{ggplyr}
\end{document}

### 2012-05-30 R/glayer_build.r

@@ -1,0 +1,72 @@
+#' glayer_build prepares an glyph layer (class glayer) for plotting
+glayer_build <- function(layer) {
+
+	if (!("embed" %in% ls(layer))) {
+		stop("layer does not have embedded subplots")
+	}
+	 
+	layer <- layer_clone(layer)
+	
+	minor <- ggplot_build(ggplot() + layer + facet_wrap("GLYPH")) 
+	
+	
+	### combine subplots (minor) into single plot
+	# data
+	data <- unpanel(minor$data[[1]])
+	data <- layer$embed$fun(data)
+	data$PANEL <- 1L
+	
+	# panel
+	xspan <- range(unlist(data[names(data) %in% .x_aes]))
+	yspan <- range(unlist(data[names(data) %in% .y_aes]))
+	panel <- ggplot_build(qplot(xspan, yspan))$panel
+
+	# scales
+	scales <- minor$plot$scales$scales
+	scales[[which_x(scales)]] <- panel$x_scales[[1]]
+	scales[[which_y(scales)]] <- panel$y_scales[[1]]
+	
+	# axis labels
+	if (!is.null(layer$embed$major_aes)) {
+		labels <- labs(layer$embed$major_aes)
+		minor$plot$options$labels[c("x", "y")] <- labels[c("x", "y")]
+	}
+	if (!is.null(layer$embed$ref_aes)) {
+		lab.names <- names(minor$plot$options$labels)
+		new <- setdiff(names(layer$embed$ref_aes), lab.names)
+		minor$plot$options$labels[new] <- layer$embed$ref_aes[new]
+	}
+		
+	# make build
+	minor$data <- list(data)
+	minor$panel <- panel
+	minor$plot$facet <- facet_null()
+	minor$plot$scales$scales <- scales
+	
+	minor
+}
+
+
+unpanel <- function(df) {
+	df$GLYPH <- as.numeric(as.character(df$PANEL))
+	df$PANEL <- NULL
+	df
+}
+
+
+which_x <- function(scales) {
+	vars <-  names_scales(scales)
+	which(vars == "x")
+}
+
+
+which_y <- function(scales) {
+	vars <- names_scales(scales)
+	which(vars == "y")
+}
+
+#' Returns the first aes of a scale, to use as a name
+names_scales <- function(scales) {
+	unlist(lapply(scales, function(s) s$aesthetics[[1]]))
+}	
+	

### 2012-05-30 R/nest.r

@@ -33,7 +33,7 @@ nest <- function(data, major_aes = NULL, x_scale = identity, y_scale = identity,
 		width <- diff(range(globals$x)) / max(globals$.gid) * unclass(width)
 	}
 	if (is.rel(height)) {
-		height <- diff(range(globals$Y)) / max(globals$.gid) * unclass(height)
+		height <- diff(range(globals$y)) / max(globals$.gid) * unclass(height)
 	}
 	
 	# combine overlapping subplots

### 2012-05-30 demo.r

@@ -1,7 +1,7 @@
 # ggplyr features demo - sift for tests
 
 # setwd("/Users/GarrettsiMac/Dropbox/research")
-# setwd("/Users/garrettgrolemund/Dropbox/research")
+# setwd("/Users/garrettgrolemund/Documents/git")
 library(devtools)
 load_all("ggplyr")
 
@@ -34,7 +34,17 @@ qplot(Fertility, Agriculture, data = test.data, color = interaction(lat, long))
 
 # use plyr to create a graph
 # the aesthetics can be calculated groupwise
-ggplot() + geom_plyr(aes(x = Fertility, y = Agriculture, color = rank(Fertility)), test.data, c("lat", "long"), geom = "point", size = 3)
+ggplot() + 
+  geom_plyr(aes(x = long, y = lat, color = rank(Fertility)), 
+    test.data, c("lat", "long"), geom = "jitter", size = 3, 
+    position = position_jitter(width = 0.2, height = 0.2))
+
+ggplot() + 
+  geom_jitter(aes(x = long, y = lat, color = rank(Fertility)), 
+    data = test.data, size = 3, 
+    position = position_jitter(width = 0.2, height = 0.2))
+
+
 
 data <- group_by(test.data, c("lat", "long"))
 lapply(data, fun(qplot, "data", x = Fertility, y = Agriculture, color = rank(Education), size = I(3), xlim = range(test.data$Fertility), ylim = range(test.data$Agriculture)))

### 2012-05-30 geom_dart.r

@@ -1,0 +1,126 @@
+#' @include plyr_aes.r
+NULL
+
+geom_dart <- function (mapping = NULL, data = NULL, glyphs = NULL, stat = "identity", position = "identity", ...) { 
+	GeomDart$new(mapping = mapping, data = data, glyphs = glyphs, 
+		stat = stat, position = position, ...)
+}
+
+
+GeomDart <- proto(ggplot2:::Geom, {
+	objname <- "dart"
+	
+	compute_aesthetics <- plyr_aesthetics_summary
+	
+	reparameterise <- function(., df, params) {
+
+		df$size <- df$size %||% params$size %||% 
+			(resolution(df$y, FALSE) * 0.2)
+		df$width <- df$width %||% params$width %||% 
+			(resolution(df$x, FALSE) * 2)
+		df$angle <- df$angle %||% params$angle %||% 0			
+		df				
+	}
+	
+	# finalize minor axes from non_position aes
+	compute_glyphs <- function(df, layer) {
+	
+		l <- sqrt((df$width + 1)^2 * df$size^2) / 2
+		lambda <- atan(1 / df$width)
+		
+		df$ax <- l * cos(df$angle + pi - lambda)
+		df$ay <- l * sin(df$angle + pi - lambda)
+		
+		df$bx <- l * cos(df$angle + pi + lambda)
+		df$by <- l * sin(df$angle + pi + lambda)
+		
+		df$cx <- df$width * df$size * cos(df$angle) / 2
+		df$cy <- df$width * df$size * sin(df$angle) / 2
+		
+		df
+		
+	}
+	
+	# map minor axes into major axes. 
+	# scale minor x and y to 0 - 1
+	# use width and height to compute actual x, y
+	# center at major x and y
+	# Will need width, height, majors
+	# majors already stored in data frame as x and y
+	map_glyphs <- function(df, width, height) {
+		if (any(table$df > 1)) {
+			message("Multiple subplots mapped to same glyph. do you need to adjust your glyph's group criteria?")
+		}
+		
+		data <- ddply(df, "GLYPH", function(df) {
+			df <- df[c(1,1,1), ]
+			df$x <- c(ax, bx, cx)
+			df$y <- c(ay, by, cy)
+			df[setdiff(names(df), c("ax", "bx", "cx", "ay", "by", 
+				"cy"))]
+		})
+	}
+	
+	
+	draw <- function(., data, scales, coordinates, ...) {
[…]

### 2012-05-30 R/dyply.r
clearing out directory for glyphmaps

@@ -1,24 +1,0 @@
-#' split data.frame, apply function, and return results as a ggplot2 layer object
-#' 
-#' For each subset of a data.frame, apply a function and then combine results into a ggplot2 layer object. This layer can be used to build a ggplot2 plot. Results will be combined with the function provided as the .combine.fun argument. This function should be compatible with the output of the function provided as the .apply.fun argument.
-#'
-#' @param .data data.frame to be processed
-#' @param .split.fun function used to split the data set, or the names of the variables to split on as character strings.
-#' @param .apply.fun function to apply to each subset of the data.frame
-#' @param .combine.fun function to combine output of .apply.fun into a graph. .combine.fun should return a ggplot2 layer object, or a list of ggplot2 layer objects.
-#' @param .progress name of the progress bar to use, see \code{\link{create_progress_bar}}
-#' @param .parallel if TRUE, apply function in parallel using the parallel backend provided by foreach
-#' @return A ggplot2 layer object
-#' @export
-dyply <- function(.data, .split.fun = NULL, .apply.fun = NULL, .combine.fun = fun(as_is, "layers"), .progress = "none", .parallel = FALSE) {
-
-	if (!is.call(.split.fun)) {
-		vars <- .split.fun
-		.split.fun <- function(x) group_by(x, vars = vars)
-	}
-	
-	data <- .split.fun(.data)
-	
-	lyply(data, .apply.fun, .combine.fun, .progress, .parallel)
-	
-}

### 2012-05-30 R/fun.r

@@ -1,22 +1,0 @@
-#' Reorganize a function to take a single argument
-#'
-#' fun reorganizes a function to take a single argument. It sets other specified arguments to known values. fun can be used to change the order of arguments within a function. This is helpful for passing functions to \code{\link{plyr}} style functions which automatically use the first argument of a function's definition.
-#'
-#' @param name name of the function to reorganize as an object or character string
-#' @param .input name of the argument that new function will take
-#' @param ... values to set other arguments of function to. If an argument is not named it will be set to its default value.
-#' @export
-fun <- function(name, .input, ...) {
-
-	if (!is.function(name)) {
-		FUN <- get(name)
-	} else {
-		FUN <- name
-	}
-	
-	args <- c(FUN, .input = as.name("x"), match.call(expand.dots = FALSE)$...)
-	names(args)[names(args) == ".input"] <- .input
-	function(x) eval(as.call(args))
-}
-# ggplot() + fun(geom_point, "data", mapping = aes(x = hwy, y = cty), size = 5)(mpg)
-# fun(median, "x")

### 2012-05-30 R/geom_plyr.r

@@ -1,9 +1,0 @@
-# geom_plyr performs groupwise aesthetic calculation
-geom_plyr <- function(mapping = NULL, data, .vars, geom = "point", ...) {
-	geom_fun <- get(paste("geom", geom, sep = "_"))
-	data$.gid <- id(data[.vars], drop = TRUE)
-	layer <- geom_fun(mapping, data, ...)
-	layer$plyr <- TRUE
-	layer$compute_aesthetics <- compute_plyr_aesthetics
-	layer
-}

### 2012-05-30 R/glayer_build.r

@@ -1,72 +1,0 @@
-#' glayer_build prepares an glyph layer (class glayer) for plotting
-glayer_build <- function(layer) {
-
-	if (!("embed" %in% ls(layer))) {
-		stop("layer does not have embedded subplots")
-	}
-	 
-	layer <- layer_clone(layer)
-	
-	minor <- ggplot_build(ggplot() + layer + facet_wrap("GLYPH")) 
-	
-	
-	### combine subplots (minor) into single plot
-	# data
-	data <- unpanel(minor$data[[1]])
-	data <- layer$embed$fun(data)
-	data$PANEL <- 1L
-	
-	# panel
-	xspan <- range(unlist(data[names(data) %in% .x_aes]))
-	yspan <- range(unlist(data[names(data) %in% .y_aes]))
-	panel <- ggplot_build(qplot(xspan, yspan))$panel
-
-	# scales
-	scales <- minor$plot$scales$scales
-	scales[[which_x(scales)]] <- panel$x_scales[[1]]
-	scales[[which_y(scales)]] <- panel$y_scales[[1]]
-	
-	# axis labels
-	if (!is.null(layer$embed$major_aes)) {
-		labels <- labs(layer$embed$major_aes)
-		minor$plot$options$labels[c("x", "y")] <- labels[c("x", "y")]
-	}
-	if (!is.null(layer$embed$ref_aes)) {
-		lab.names <- names(minor$plot$options$labels)
-		new <- setdiff(names(layer$embed$ref_aes), lab.names)
-		minor$plot$options$labels[new] <- layer$embed$ref_aes[new]
-	}
-		
-	# make build
-	minor$data <- list(data)
-	minor$panel <- panel
-	minor$plot$facet <- facet_null()
-	minor$plot$scales$scales <- scales
-	
-	minor
-}
-
-
-unpanel <- function(df) {
-	df$GLYPH <- as.numeric(as.character(df$PANEL))
-	df$PANEL <- NULL
-	df
-}
-
-
-which_x <- function(scales) {
-	vars <-  names_scales(scales)
-	which(vars == "x")
-}
-
-
-which_y <- function(scales) {
-	vars <- names_scales(scales)
-	which(vars == "y")
-}
-
-#' Returns the first aes of a scale, to use as a name
-names_scales <- function(scales) {
-	unlist(lapply(scales, function(s) s$aesthetics[[1]]))
-}	
-	

### 2012-05-30 R/grid.r

@@ -1,9 +1,0 @@
-#' grid takes a vector of numbers, divides it into n bins, and rounds each point to the nearest center of a bin.
-grid <- function(x, npoints = 10, range. = range(x)) {
-	binwidth <- diff(range.) / npoints
-	x <- x - range.[1] + binwidth / 2
-	a <- x %/% binwidth + c(0,1)[round(x %% binwidth / binwidth) + 1]
-	a[a == 0] <- 1 # include leftmost point
-	x <- range.[1] + binwidth * a
-	x - binwidth/2
-}

### 2012-05-30 R/in_place.r

@@ -1,8 +1,0 @@
-#' Combine subplots without adjusting locations
-#'
-#' @param layers a list of ggplot2 layer objects. These should be \code{\link{proto}} objects that work to create a graph when added to ggplot().
-in_place <- function(layers) {
-	
-	list(fun = function(data) data, major_aes = NULL)
-
-}

### 2012-05-30 R/lgply.r

@@ -1,16 +1,0 @@
-#' split list, apply function, and return results as a ggplot2 graph
-#' 
-#' For each element of a list, apply a function and then combine results into a ggplot2 plot. Results will be combined with the function provided as the .combine.fun argument. This function should be compatible with the output of the function provided as the .apply.fun argument.
-#'
-#' @param .data list to be processed
-#' @param .apply.fun function to apply to each piece of the list
-#' @param .combine.fun function to combine output of .apply.fun into a graph. .combine.fun should return a ggplot2 layer object, or a list of ggplot2 layer objects.
-#' @param .progress name of the progress bar to use, see \code{\link{create_progress_bar}}
-#' @param .parallel if TRUE, apply function in parallel using the parallel backend provided by foreach
-#' @return A ggplot2 object of class ggplot
-#' @export
-lgply <- function(.data, .apply.fun = NULL, .combine.fun = fun(as_is, "layers"), .progress = "none", .parallel = FALSE) {
-	
-	ggplot() + lyply(.data, .apply.fun, .combine.fun, .progress, .parallel)
-
-}

### 2012-05-30 R/merge_overlaps.r

@@ -1,21 +1,0 @@
-merge_overlaps <- function(globals, width, height) {
-	globals <- globals[c(".gid", "x", "y")]
-	globals <- arrange(globals, x)
-	g1 <- globals[1:(nrow(globals) - 1), ]
-	g2 <- globals[-1, ]
-
-	x.overlaps <- abs(g1$x - g2$x) < width 
-	y.overlaps <- abs(g1$y - g2$y) < height
-	overlaps <- x.overlaps & y.overlaps
-	
-	# but what about multiple overlaps?
-	new = g1$.gid[overlaps]
-	old = g2$.gid[overlaps]
-	still.old <- which(new %in% old)
-	while(length(still.old)) {
-		new[still.old[1]] <- new[which(old == new[still.old[1]])]
-		still.old <- which(new %in% old)
-	}
-	names(new) <- old
-	new
-}

### 2012-05-30 R/ref_box.r

@@ -1,28 +1,0 @@
-ref_box <- function(mapping = NULL, data = NULL, ...) {	
-	make_layer <- function(globals, width, height) {
-
-		ref_data <- transform(globals, xmin = x - width, xmax = x + width, 
-			ymin = y - height, ymax = y + height)
-		ref_data$x <- NULL
-		ref_data$y <- NULL
-	
-		ref_map <- lapply(names(ref_data), as.name)
-		names(ref_map) <- names(ref_data)
-		ref_map$.gid <- NULL
-		class(ref_map) <- "uneval"
-		
-		ref_layer <- geom_rect(ref_map, data = ref_data, ...)
-		ref_layer
-	}
-	
-	
-	list(mapping = mapping, make_layer = make_layer)
-	
-}
-
-
-
-just_refs <- function(ref.aes, major.aes) {
-	non_ref <- c(.x_aes, .y_aes, names(major.aes))
-	ref.aes[setdiff(names(ref.aes), non_ref)]
-}

### 2012-05-30 R/ref_vline.r

@@ -1,20 +1,0 @@
-ref_vline <- function(mapping = NULL, ...) {
-	make_layer <- function(globals, width, height) {
-
-		ref_data1 <- transform(globals, y = y - height, group = .gid)
-		ref_data2 <- transform(globals, y = y + height, group = .gid)
-		ref_data <- rbind(ref_data1, ref_data2)
-	
-		ref_map <- lapply(names(ref_data), as.name)
-		names(ref_map) <- names(ref_data)
-		ref_map$.gid <- NULL
-		class(ref_map) <- "uneval"
-		
-		ref_layer <- geom_line(ref_map, data = ref_data, ...)
-		ref_layer
-	}
-	
-	
-	list(mapping = mapping, make_layer = make_layer)
-
-}

### 2012-05-30 R/rel-class.r

@@ -1,8 +1,0 @@
-# Relative dimensions ------------------------------------------------------
-rel <- function(x) {
-  structure(x, class = "rel")
-}
-
-print.rel <- function(x, ...) print(noquote(paste(x, " *", sep = "")))
-
-is.rel <- function(x) inherits(x, "rel")

### 2012-05-30 R/rescalers.r

@@ -1,36 +1,0 @@
-rescale_01 <- free <- function(xvars, xlim=NULL) {
-	xnames <- names(xvars)
-	numberfy <- function(x) {
-		if (is.character(x)) {
-			x <- as.numeric(factor(x))
-		}	
-		if (is.factor(x)) {
-			x <- as.numeric(x)
-		}
-		x
-	}
-	xvars <- lapply(xvars, numberfy)
-  
-	if (is.null(xlim)) {
-		rng <- range(unlist(xvars), na.rm = TRUE)
-	} else {
-		rng <- xlim
-	}
-	
-	scale <- function(x) {
-		if ((rng[2] - rng[1]) == 0) {
-			x - rng[1]
-		} else {	
-			(x - rng[1]) / (rng[2] - rng[1])
-		}
-	}
-	xvars <- lapply(xvars, scale)
-	names(xvars) <- xnames
-	data.frame(xvars)
-}
-
-
-rescale_11 <- function(xvars, xlim=NULL) {
-	2 * rescale_01(xvars, xlim) - 1
-}
-

### 2012-05-30 R/util.r

@@ -1,142 +1,0 @@
-.x_aes <- c("x", "xend", "xmax", "xmin")
-.y_aes <- c("y", "yend", "ymax", "ymin")
-
-#' add_gid intelligently adds the .gid variable to the group slot of an uneval object. If the group slot is NULL, add_gid sets group = .gid. If the group slot already contains a mapping, add_gid adds .gid to this mapping with interaction().
-#'
-#' @keywords internal
-#' @param aes_group the group value of an uneval object
-#' @export
-add_gid <- function(aes_group) {
-	if (is.null(aes_group)) {
-		as.name(".gid")
-	} else {
-		as.call(list(as.name("interaction"), as.name(".gid"), aes_group))
-	}
-}
-
-
-apply_maps <- function(data, mapping, enclos = parent.frame()) {
-
-	map <- null_omit(mapping)
-	vars <- llply(map, eval, envir = data, enclos)
-	
-	n <- nrow(data)
-	lengths <- unlist(lapply(vars, length))
-	wrong <- lengths != 1 & lengths != n
-	if (any(wrong)) {
-        stop(paste("Aesthetics must either be length one, or the same length as the
-        	data", "Problems:", paste(names(wrong)[wrong], collapse = ", ")), 
-            call. = FALSE)
-    }
-    
-    data.frame(vars)
-}
-
-#' the difference between apply maps and apply major is that apply major will only allow one value per gid per variable
-apply_major <- function(data, mapping, enclos = parent.frame()) {
-
-	map <- null_omit(mapping)
-	vars <- llply(map, eval, envir = data, enclos)
-	vars  <- llply(vars, unique)
-	
-	lengths <- unlist(llply(vars, length))
-	wrong <- lengths != 1 
-	if (any(wrong)) {
-        stop(paste("Major aesthetics must return one value per variable per subplot group (.gid).", "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
-    }
-    
-    data.frame(vars)
-}
-
-# combine_aes is used by ggtransform and ref_box
-#' combine_aes combines two uneval objects overwriting values of the second that are defined in the first.
-#'
-#' @keywords internal
-#' @param map1 an uneval object
-#' @param map2 an uneval object
-#' @export
-combine_aes <- function(map1, map2) {
-	structure(c(map1, map2[setdiff(names(map2), names(map1))]),
[…]

### 2012-05-30 R/yyply.r

@@ -1,26 +1,0 @@
-#' split ggplot2 layer, apply function, and return results as a ggplot2 layer object
-#' 
-#' For each subset of data in a ggplot2 layer, apply a function and then combine results into a ggplot2 layer object. This layer can be used to build a ggplot2 plot. Results will be combined with the function provided as the .combine.fun argument. This function should be compatible with the output of the function provided as the .apply.fun argument. Note, yyply can only manipulate data that is stored in the ggplot2 layer. If the layer inherits data from a global \code{\layer{ggplot}} call, yyplot will not have access to the data.
-#'
-#' @param .data ggplot2 layer to be processed
-#' @param .split.fun function used to split the data set that underlies the ggplot2 layer, or the names of the variables to split on as character strings.
-#' @param .apply.fun function to apply to each subset of the ggplot2 layer
-#' @param .combine.fun function to combine output of .apply.fun into a graph. .combine.fun should return a ggplot2 layer object, or a list of ggplot2 layer objects.
-#' @param .progress name of the progress bar to use, see \code{\link{create_progress_bar}}
-#' @param .parallel if TRUE, apply function in parallel using the parallel backend provided by foreach
-#' @return A ggplot2 layer object
-#' @export
-yyply <- function(.data, .split.fun = NULL, .apply.fun = NULL, .combine.fun = fun(as_is, "layers"), .progress = "none", .parallel = FALSE) {
-	
-	if (!is.call(.split.fun)) {
-		.split.fun <- fun(group_by, "data", vars = .split.fun)
-	}
-	
-	data <- .split.fun(.data$data)
-	data <- lapply(data, recreate_geom(.data))
-	
-	data <- llply(data, .fun = .apply.fun, .progress = .progress, 
-		.parallel = .parallel)
-	
-	.combine.fun(data)
-}

### 2012-05-30 R2/embed_layers.r

@@ -1,11 +1,0 @@
-#' embed_layers takes a list of layers and arranges them into an embedded layer that will be plotted as an embedded plot. Each layer must have the same parameters, geom, stat, and aesthetics. Only the data differs.
-#'
-#' note: .fun should return a list(fun = combine function, major_aes = aes(x = , y = )
-embed_layers <- function(layers, .fun) {
-	prototype <- layer_clone(layers[[1]])
-	prototype$data <- individual_data(layers)
-	prototype$embed <- .fun(prototype$data)
-	gglayer(prototype)
-}
-
-

### 2012-05-30 R2/geom_single.r

@@ -1,27 +1,0 @@
-#' geom_single representa a dataframe with a single point
-geom_single <- function(mapping = NULL, data, ...) {
-	data <- lapply(mapping, eval, envir = data, enclos = parent.frame())
-	lengths <- lapply(data, length)
-	wrong <- lengths != 1
-	if (any(wrong)) {
-		stop(paste("geom_single requires all aes to return one value per variable.",
-			"Problems:", paste(names(wrong)[wrong], collapse = ", ")))
-	}
-	
-	data <- data.frame(data)
-	maps <- map_to_names(mapping)
-	labels <- labs(mapping)
-	
-	list(geom_point(maps, data = data, ...), labels)
-}
-
-
-map_to_names <- function(uneval) {
-	vars <- lapply(names(uneval), as.name)
-	names(vars) <- names(uneval)
-	class(vars) <- "uneval"
-	vars
-}
-
-
-compute_single_aesthetic <- 

### 2012-05-30 R2/internal-functions.r

@@ -1,221 +1,0 @@
-.x_aes <- c("x", "xend", "xmax", "xmin")
-.y_aes <- c("y", "yend", "ymax", "ymin")
-
-
-#' add_gid intelligently adds the .gid variable to the group slot of an uneval object. If the group slot is NULL, add_gid sets group = .gid. If the group slot already contains a mapping, add_gid adds .gid to this mapping with interaction().
-#'
-#' @keywords internal
-#' @param aes_group the group value of an uneval object
-#' @export
-add_gid <- function(aes_group) {
-	if (is.null(aes_group)) {
-		as.name(".gid")
-	} else {
-		as.call(list(as.name("interaction"), as.name(".gid"), aes_group))
-	}
-}
-
-
-#' Relocate subplots within main plot
-#'
-#' adjust_position_nest is substituted for a layer's position_adjust function when embedding graphics in a nested design
-adjust_position_nest <- function (., data, plot, new) {
-
-	data <- ddply(data, "PANEL", function(data) {
-		.$position$adjust(data)
-     })
-	
-	if ("subplots" %in% ls(.)) {
-		
-		# locate x and y related variables in data
-		xvar <- get_xs(data)
-		yvar <- get_ys(data)
-		
-		# scale if necessary
-		if (!identical(.$subplots$x_scale, identity) || 
-			!identical(.$subplots$y_scale, identity)) {
-	    		data <- ddply(data, ".gid", 
-	    			function(df) {
-	      				df[xvar] <- lapply(df[xvar], .$subplots$x_scale)
-	      				df[yvar] <- lapply(df[yvar], .$subplots$y_scale)
-	      				df
-	    			}
-	    		)
-	  	}
-	  	
-	  	width <- .$subplots$width
-	  	height <- .$subplots$height
-	  	
-	  	if (is.rel(width)) {
-			width <- diff(range(data$X)) / max(data$.gid) * unclass(width)
-		}
-    
-		if (is.rel(height)) {
-			height <- diff(range(data$Y)) / max(data$.gid) * unclass(height)
-		}
-	  	
-	  	# update x and y related variables
-		fun.x <- function(x) data$X + rescale11(x) * width
-	  	fun.y <- function(y) data$Y + rescale11(y) * height
-	  	data[xvar] <- lapply(data[xvar], fun.x)
-	    data[yvar] <- lapply(data[yvar], fun.y)
-		
-    }
-    
[…]

### 2012-05-30 R2/reference-functions.R

@@ -1,25 +1,0 @@
-ref_box <- function(layers, mapping = NULL, ...) {
-	# make mapping
-	xy <- prototype$mapping[c("x", "y", "X", "Y")]
-	map <- combine_aes(mapping, xy)
-
-		
-	layer <- geom_rect(data = prototype$data, mapping = map, ...)
-	layer$plyr <- prototype$plyr
-	layer$subplots <- prototype$subplots
-	layer$compute_aesthetics <- prototype$compute_aesthetics
-	layer$adjust_position <- prototype$adjust_position
-	gglayer(layer)
-}
-
-ref_hline <- function(layers, mapping = NULL, ...) {
-	
-}
-
-ref_vline <- function(layers, mapping = NULL, ...) {
-
-}
-
-ref_points <- function(layers, mapping = NULL, ...) {
-	
-}

### 2012-05-30 R2/scrap.R

@@ -1,24 +1,0 @@
-
-
-	panel <- major$panel
-	
-	scales <- minor$plot$scales$scales
-	SCALES <- major$plot$scales$scales
-	scales[which_xy(scales)] <- SCALES[which_xy(SCALES)]
-	minor$plot$scales$scales <- scales # necessary?
-	 
-	minor$plot$facet <- facet_null()
-	minor$plot$options$labels[c("x", "y")] <- major$plot$options$labels[c("x", "y")]
-	
-	ggplot2:::reset_scales(panel)
-	scale_x <- function() minor$plot$scales$get_scales("x")
-    scale_y <- function() minor$plot$scales$get_scales("y")
-    panel <- ggplot2:::train_position(panel, list(data), scale_x(), scale_y())
-    data <- ggplot2:::map_position(panel, list(data), scale_x(), scale_y())
-	panel <- ggplot2:::train_ranges(panel, minor$plot$coordinates)
-	
-	minor$data <- data
-	minor$panel <- panel
-	
-	minor
-}

### 2012-05-30 demo.r

@@ -1,312 +1,0 @@
-# ggplyr features demo - sift for tests
-
-# setwd("/Users/GarrettsiMac/Dropbox/research")
-# setwd("/Users/garrettgrolemund/Documents/git")
-library(devtools)
-load_all("ggplyr")
-
-#######################################
-###       transform a graph         ###
-#######################################
-
-# make a graph
-head(test.data)
-(p <- qplot(Fertility, Agriculture, data = test.data, color = Education))
-
-# change mappings
-ggtransform(p, mapping = aes(size = Catholic))
-
-# nullify mappings, change axis
-ggtransform(p, mapping = aes(y = Education, color = NULL, size = Catholic))
-
-# change parameter
-ggtransform(p, mapping = aes(y = Education, color = NULL, size = Catholic), shape = 15)
-
-
-
-
-#########################################################
-###       use plyr to make groupwise mappings         ###
-#########################################################
-
-# Our data is (arbitrarily) grouped into four locations
-qplot(Fertility, Agriculture, data = test.data, color = interaction(lat, long))
-
-# use plyr to create a graph
-# the aesthetics can be calculated groupwise
-ggplot() + 
-  geom_plyr(aes(x = long, y = lat, color = rank(Fertility)), 
-    test.data, c("lat", "long"), geom = "jitter", size = 3, 
-    position = position_jitter(width = 0.2, height = 0.2))
-
-ggplot() + 
-  geom_jitter(aes(x = long, y = lat, color = rank(Fertility)), 
-    data = test.data, size = 3, 
-    position = position_jitter(width = 0.2, height = 0.2))
-
-
-
-data <- group_by(test.data, c("lat", "long"))
-lapply(data, fun(qplot, "data", x = Fertility, y = Agriculture, color = rank(Education), size = I(3), xlim = range(test.data$Fertility), ylim = range(test.data$Agriculture)))
-
-ggplot() + geom_plyr(aes(x = trans, fill = mean(hwy), group = year), mpg, "year", geom = "bar", position = "dodge")
-
-# compare with	
-ggplot() + geom_bar(aes(x = trans, fill = mean(hwy), group = year), data = mpg, position = "dodge")
-
-
-# new tactic for overplotting
-ggplot() + geom_jitter(aes(x = temperature, y = ozone, color = lat), data = nasa)	
[…]

### 2012-05-30 exploration.r

@@ -1,88 +1,0 @@
-# okay this works
-	
-	
-f <- qplot(Fertility, Education, data = test.data, color = rank(Catholic), size = I(3)) + facet_wrap(c("lat", "long"))
-
-f <- qplot(Fertility, data = test.data, size = I(3)) + facet_wrap(c("lat", "long"))
-
-fplot <- ggplot_build(f)
-
-fdata <- fplot$data[[1]]
-
-gdata <- ddply(test.data, c("lat", "long"), summarise, X = mean(Fertility), Y = mean(Education))
-gdata$.gid <- id(gdata[c("lat", "long")], drop = TRUE)
-names(fdata)[names(fdata) == "PANEL"] <- ".gid"
-
-data <- join(fdata, gdata, by = ".gid")
-
-data[c(".gid", "x", "y")] <- ddply(data, ".gid", summarize, x = (x - mean(x)) / var(x) + X, y = (y - mean(y)) / var(y) + Y)
-
-data$PANEL <- 1
-
-g <- qplot(x, y, data = data)
-panelist <- ggplot_build(g)$panel
-
-fplot$data <- list(data)
-fplot$panel <- panelist
-# fplot$plot <- g
-fplot$plot$facet <- facet_null()
-fplot$plot$scales$scales[[2]] <- panelist$x_scales[[1]]
-fplot$plot$scales$scales[[3]] <- panelist$y_scales[[1]]
-fplot$plot$options$labels[c("x", "y")] <- g$options$labels[c("x", "y")]
-gshow(fplot)
-
-gshow <- function(data, newpage = is.null(vp), vp = NULL, ...){
-    if (newpage) 
-        grid.newpage()
-    gtable <- ggplot_gtable(data)
-    if (is.null(vp)) {
-        grid.draw(gtable)
-    }
-    else {
-        if (is.character(vp)) 
-            seekViewport(vp)
-        else pushViewport(vp)
-        grid.draw(gtable)
-        upViewport()
-    }
-    invisible(data)
-}
-test.data2 <- test.data
-test.data2$.gid <- id(test.data2[c("lat", "long")], drop = TRUE)
-layer <- geom_histogram(aes(x = Fertility, fill = long), data = test.data2, size = 3)
-
-
-
[…]

### 2012-05-30 geom_dart.r

@@ -1,126 +1,0 @@
-#' @include plyr_aes.r
-NULL
-
-geom_dart <- function (mapping = NULL, data = NULL, glyphs = NULL, stat = "identity", position = "identity", ...) { 
-	GeomDart$new(mapping = mapping, data = data, glyphs = glyphs, 
-		stat = stat, position = position, ...)
-}
-
-
-GeomDart <- proto(ggplot2:::Geom, {
-	objname <- "dart"
-	
-	compute_aesthetics <- plyr_aesthetics_summary
-	
-	reparameterise <- function(., df, params) {
-
-		df$size <- df$size %||% params$size %||% 
-			(resolution(df$y, FALSE) * 0.2)
-		df$width <- df$width %||% params$width %||% 
-			(resolution(df$x, FALSE) * 2)
-		df$angle <- df$angle %||% params$angle %||% 0			
-		df				
-	}
-	
-	# finalize minor axes from non_position aes
-	compute_glyphs <- function(df, layer) {
-	
-		l <- sqrt((df$width + 1)^2 * df$size^2) / 2
-		lambda <- atan(1 / df$width)
-		
-		df$ax <- l * cos(df$angle + pi - lambda)
-		df$ay <- l * sin(df$angle + pi - lambda)
-		
-		df$bx <- l * cos(df$angle + pi + lambda)
-		df$by <- l * sin(df$angle + pi + lambda)
-		
-		df$cx <- df$width * df$size * cos(df$angle) / 2
-		df$cy <- df$width * df$size * sin(df$angle) / 2
-		
-		df
-		
-	}
-	
-	# map minor axes into major axes. 
-	# scale minor x and y to 0 - 1
-	# use width and height to compute actual x, y
-	# center at major x and y
-	# Will need width, height, majors
-	# majors already stored in data frame as x and y
-	map_glyphs <- function(df, width, height) {
-		if (any(table$df > 1)) {
-			message("Multiple subplots mapped to same glyph. do you need to adjust your glyph's group criteria?")
-		}
-		
-		data <- ddply(df, "GLYPH", function(df) {
-			df <- df[c(1,1,1), ]
-			df$x <- c(ax, bx, cx)
-			df$y <- c(ay, by, cy)
-			df[setdiff(names(df), c("ax", "bx", "cx", "ay", "by", 
-				"cy"))]
-		})
-	}
-	
-	
-	draw <- function(., data, scales, coordinates, ...) {
[…]

### 2012-05-30 paper/demo.Rnw
moved over glyphmaps files

@@
 
-\title{A Minimal Demo of glyphmaps}
+\title{A Minimal Demo of ggplyr/glyphmaps/modelglyphs}
 
@@
 \maketitle
-This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change).
-\texttt{glyphmaps} was written to create plots like the ones below.
-Each was made with \texttt{glyphmaps}.
+This document demonstrates the basic features of the \texttt{ggplyr} package (name may change).
+\texttt{ggplyr} was written to create plots like the ones below.
+Each plot was made with \texttt{ggplyr}.
 The first two plots recreate plots from \citet{wickham2011split}.
@@
 
-To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with
-
-<<>>=
-library(devtools)
-load_all("glyphmaps")
-@
-
-I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \Sexpr{install_github()}.
-
 \section{Glyphs}
-The above graphs are similar in that they are all built around glyphs.
-I define a glyph as a geom that visually summarises information about a subset of data.
-Glyphs can be quite simple, or very complex.
-Often glyphs contain their own internal coordinate systems.
+The common feature of these graphs is that they are built around glyphs.
+Glyphs are geometric objects (i.e, geoms) designed to display information within the geom.
+In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system.
+In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
 The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
@@
 @
-
-\section{glyphmaps geoms}
-\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs.
-These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots.
-\texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.
-
-<<geom_star>>=
-@
-
-<<geom_radar>>=
-@
-
-<<geom_scatterplots>>=
-@
-
-\subsection{new arguments}
-\label{sec:args}
-Each \texttt{glyphmaps} geom takes two new arguments, \Sexpr{glyphs} and \Sexpr{reference}.
-\Sexpr{glyphs} controls how the data will be divided into subsets to be paired with individual glyphs.
-\Sexpr{reference} lets the user draw reference features with the glyphs to facilitate comparisons.
-
-\Sexpr{glyphs} takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs.
-The \Sexpr{glyphs} argument works just like a \Sexpr{.vars} argument in a \Sexpr{ddply} call.
-(In fact, the \Sexpr{glyphs} argument in a \texttt{glyphmaps} geom will be passed into a \Sexpr{ddply} call as the \Sexpr{.vars} argument.)
-
-\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points.
-These objects help viewers make comparisons between data inside different glyphs.
-
-<<ref-type-examples>>=
-@
-
-<<comparison-with-boxes>>=
-@
-
-\section{Generalized glyphs}
-Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \Sexpr{glyph} and \Sexpr{plyr} functions.
-
-<<bar-chart-glyph>>=
-@
-
-<<point-overplotting-ex>>=
-@
-
-The plyr function also allows users to key aesthetics to subgroups of data.
-
-<<groupwise-aesthetics>>=
-@
-
-\section{Discussion - glyphs, embedded graphics, hierarchies}
-Glyphs are geometric objects (i.e, geoms) designed to display information within the geom.
-In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system.
-In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
 
@@
 Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one order to have a concrete interpretation.)
+
+To run the examples contained below, you'll have to download the \texttt{ggplyr} development package from \url{https://github.com/garrettgman/ggplyr} and install it with
+
+<<boring-random>>=
+library(devtools)
+load_all("ggplyr")
+@
+
+I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \Sexpr{install_github()}.
 

### 2012-05-30 paper/demo.tex
moved over glyphmaps files

@@
 
-To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{load\usebox{\hlnormalsizeboxunderscore}all}\hlkeyword{(}\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-## Error: Can't find package glyphmaps
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \begin{verbatim}
-
-Error in paste(repo, branch, sep = "/", collapse = ", ") : 
-  argument "repo" is missing, with no default
-
-\end{verbatim}
-.
-
 \section{Glyphs}
-The above graphs are similar in that they are all built around glyphs.
-I define a glyph as a geom that visually summarises information about a subset of data.
-Glyphs can be quite simple, or very complex.
-Often glyphs contain their own internal coordinate systems.
+The common feature of these graphs is that they are built around glyphs.
+Glyphs are geometric objects (i.e, geoms) designed to display information within each geom.
+In other words, a glyph can display information even if it ids drawn by itself, without references to an external coordinate system.
+In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
 The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
 
-\section{glyphmaps geoms}
-\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs.
-These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots.
-\texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.
+Glyphs reveal a heirarchical structure to graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot.
+Sometimes these subplots are not very interesting, as in the subplot created by a single point geom.
+Sometimes, they are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.
 
-\subsection{new arguments}
-\label{sec:args}
-Each \texttt{glyphmaps} geom takes two new arguments, \begin{verbatim}
+Graphs inherit this heirarchical structure from the data they describe.
+Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data.
+Humans innately perform this process when collecting data.
+It is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.
 
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- and \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'reference' not found
-
-\end{verbatim}
-. \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- controls how the data will be divided into subsets to be paired with individual glyphs. \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'reference' not found
-
-\end{verbatim}
- lets the user draw reference features with the glyphs to facilitate comparisons.
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs. The \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- argument works just like a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object '.vars' not found
-
-\end{verbatim}
- argument in a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'ddply' not found
-
-\end{verbatim}
- call. (In fact, the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- argument in a \texttt{glyphmaps} geom will be passed into a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'ddply' not found
-
-\end{verbatim}
- call as the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object '.vars' not found
-
-\end{verbatim}
- argument.)
-
-\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points.
-These objects help viewers make comparisons between data inside different glyphs.
-
-\section{Generalized glyphs}
-Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyph' not found
-
-\end{verbatim}
- and \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'plyr' not found
-
-\end{verbatim}
- functions.
-
-The plyr function also allows users to key aesthetics to subgroups of data.
-
-\section{Discussion - glyphs, embedded graphics, hierarchies}
-Glyphs are geometric objects (i.e, geoms) designed to display information within the geom.
-In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system.
-In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
-
-Glyphs reveal a hierarchical structure of graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot.
-Sometimes these subplots are not very interesting, as in the subplot created by a single point geom.
-At other times, thee subplots are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.
-
-Graphs inherit this hierarchical structure from the data they describe.
-Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data.
-Humans innately perform this process when collecting data; it is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.
-
-Glyph maps simultaneously expose data from multiple levels of the hierarchy.
+Glyph maps simultaneously expose data from multiple levels of this process.
 As a set of geoms, the glyphs reveal relationships between data points in the higher level, compact data set.
@@
 Because each glyph is equivalent to an independent plot, glyph maps describe any plot that embeds other plots within itself.
-Embedded plots occur in a variety of contexts, and I have a stack of photocopies to help demonstrate this.
+Embedded plots occur in a variety of contexts and I have a stack of photocopies to help demonstrate this.
 Embedded plots also fit into the grammar of graphics, where they represent an extension of facetting.
@@
 
-(Interesetingly most embedded plots have settled to the stage of embedding plots within a geographic map.
+(Interesetingly most embedded plots have settled to or are stuck in the stage of embedding plots within a geographic map.
 This resembles how Wainer, Tufte and others have traced the development of graphs through a concrete stage where cartesian axes seldom refered to anything other than latitude and longitude.
 Eventually graphs became more sophisticated and axes were used to represent  abstract concepts.
-Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one order to have a concrete interpretation.)
+Infovis developed from SciVis in the same way.
+Scivis focused on visualizing physical dimensions and eventually spawned InfoVis, which focuses on visualizing abstract dimensions.
+Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one of them to have a concrete interpretation.)
 
@@
 \end{ennumerate}
-
-\subsection{geom_plyr}
-\texttt{ggplyr} provides a way for users to compute aesthetics by
 

### 2012-05-30 R/glayer-class.r

@@ -1,0 +1,119 @@
+#' @include glyphs-class.r
+NULL
+
+check_glayer <- function(object) {
+	errors <- character()
+	if (!is.proto(object@layer)) {
+		msg <- "glayer must be a proto object."
+		errors <- c(errors, msg)
+	}
+	if (length(errors) == 0) 
+		TRUE
+	else
+		errors
+}
+
+setOldClass(c("proto", "environment"))
+
+#' glayer class
+#'
+#' glayers are layers made with ggplyr methods. They are equivalent to the layers made by ggplot2 functions in all ways except that they contain an extra grouping variable (to denote subplot membership) and a plyr_function slot, which correctly locates subplots within the graph when plotting.
+#'
+#' @name glayer-class
+#' @rdname glayer-class
+#' @exportClass glayer
+#' @aliases show,glayer-method
+#' @aliases c,glayer-method
+#' @aliases rep,glayer-method
+#' @aliases ls,glayer-method
+#' @aliases [,glayer-method
+#' @aliases [<-,glayer-method
+#' @aliases $,glayer-method
+#' @aliases $<-,glayer-method
+#' @aliases +,ggplot,glayer-method
+#' @aliases +,glyphs,glayer-method
+#' @aliases ggtransform,glayer-method
+setClass("glayer", representation(layer = "proto"), validity = check_glayer)
+
+#' @export
+setMethod("show", signature(object = "glayer"), function(object) {
+	print(object@layer) 
+})
+
+#' @export
+setMethod("c", signature(x = "glayer"), function(x, ...){
+	# c(get_layer(x), unlist(lapply(list(...), get_layer)))
+	stop("object of type 'glayer' is not subsettable")
+})
+
+#' @export
+setMethod("rep", signature(x = "glayer"), function(x, ...){
+	stop("object of type 'glayer' is not subsettable")
+})
+
+#' @export
+setMethod("[", signature(x = "glayer"), 
+	function(x, i, j, ..., drop = TRUE) {
+    	new("glayer", layer = x@layer[i])
+	}
+)
+
+#' @export
+setMethod("[<-", signature(x = "glayer"), function(x, i, j, ..., value) {
+  	x@layer[i] <- value
+	x
+})
+
+
+#' @export
+setMethod("$", signature(x = "glayer"), function(x, name) {
+	slot(x, "layer")[[name]]
+})
+
+#' @export
[…]

### 2012-05-30 R/glayer_build.r

@@ -1,0 +1,66 @@
+#' glayer_build prepares an glyph layer (class glayer) for plotting
+glayer_build <- function(layer) {
+  if (!("embed" %in% ls(layer))) {
+    stop("layer does not have embedded subplots")
+  }
+	 
+  layer <- layer_clone(layer)
+  layer$data <- layer$assign_glyphs(layer$data)
+  minor <- ggplot_build(ggplot() + layer + facet_wrap("GLYPH")) 
+	
+	
+  ### combine subplots (minor) into single plot
+  # data
+  data <- unpanel(minor$data[[1]])
+  data <- layer$combine_glyphs(data)
+  data$PANEL <- 1L
+	
+  # panel
+  xspan <- range(unlist(data[names(data) %in% .x_aes]))
+  yspan <- range(unlist(data[names(data) %in% .y_aes]))
+  panel <- ggplot_build(qplot(xspan, yspan))$panel
+
+  # scales
+  scales <- minor$plot$scales$scales
+  scales[[which_x(scales)]] <- panel$x_scales[[1]]
+  scales[[which_y(scales)]] <- panel$y_scales[[1]]
+	
+  # axis labels
+  if (!is.null(layer$embed$major.aes)) {
+    labels <- labs(layer$embed$major.aes)
+    minor$plot$options$labels[c("x", "y")] <- labels[c("x", "y")]
+  }
+		
+  # make build
+  minor$data <- list(data)
+  minor$panel <- panel
+  minor$plot$facet <- facet_null()
+  minor$plot$scales$scales <- scales
+	
+  minor
+}
+
+
+unpanel <- function(df) {
+  df$GLYPH <- as.numeric(as.character(df$PANEL))
+  df$PANEL <- NULL
+  df
+}
+
+
+which_x <- function(scales) {
+  vars <-  names_scales(scales)
+  which(vars == "x")
+}
+
+
+which_y <- function(scales) {
+  vars <- names_scales(scales)
+  which(vars == "y")
+}
+
+#' Returns the first aes of a scale, to use as a name
+names_scales <- function(scales) {
+  unlist(lapply(scales, function(s) s$aesthetics[[1]]))
+}	
+	

### 2012-05-30 R/glyph.r

@@ -1,0 +1,118 @@
+#' glyph turns an ordinary layer into a set of glyphs. To do this, it must:
+#' 1. ensure that the x and y aesthetics are mapped to a single value
+#' 2. add reference boxes
+#' 3. handle merging
+#' 4. split up the data
+glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95), 
+  height = rel(0.95), x_scale = identity, y_scale = identity, 
+  merge.overlaps = FALSE, reference = NULL, .ref = FALSE) {
+  
+  missing <- c(is.null(major.aes$x), is.null(major.aes$y))
+  if (any(missing)) {
+    stop(paste("Missing required aesthetic in major.aes:", 
+      paste(c("x", "y")[missing], collapse = ", ")))
+  }
+  
+  layer <- layer_clone(layer)
+  layer$embed <- list(width = width, height = height, 
+    x_scale = x_scale, y_scale = y_scale, merge.overlaps = merge.overlaps,
+    major.aes = major.aes[c("x", "y")], glyph.by = glyph.by)
+  layer$assign_glyphs <- assign_glyphs
+  layer$combine_glyphs <- combine_glyphs
+  if (.ref) layer$combine_glyphs <- combine_refs
+  layer$compute_aesthetics <- plyr_aesthetics
+    
+  if (is.null(reference)) {
+  	glayer(layer)
+  } else {
+  	ref.layer <- reference(layer, major.aes, glyph.by, width, height, 
+  	  merge.overlaps)
+  	list(ref.layer, glayer(layer))
+  }
+} 
+  
+assign_glyphs <- function(., data) {
+  # major x and y
+  data$GLYPH <- id(data[embed$glyph.by], drop = TRUE)
+  globals <- ddply(data, "GLYPH", apply_glyphs, embed$major.aes)
+    
+  # parse width, height
+  width <- embed$width
+  height <- embed$height
+
+  if (is.rel(width)) {
+    .$embed$width <- (diff(range(vet(globals$x))) + unclass(width)) / 
+      length(unique(globals$x)) * unclass(width)
+  }
+  if (is.rel(height)) {
+    .$embed$height <- (diff(range(vet(globals$y))) + unclass(height)) / 
+      length(unique(globals$y)) * unclass(height)
+  }
+
+  if (embed$merge) {
[…]

### 2012-05-30 R/glyph_aesthetics.r

@@ -1,0 +1,50 @@
+glyph_aesthetics <- function (., data, plot) {
+  aesthetics <- .$layer_mapping(plot$mapping)
+  if (!is.null(.$subset)) {
+    include <- data.frame(eval.quoted(.$subset, data, plot$env))
+    data <- data[rowSums(include, na.rm = TRUE) == ncol(include), ]
+  }
+  if (!is.null(.$geom_params$group)) {
+    aesthetics["group"] <- .$geom_params$group
+  }
+  scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
+  compact(eval.glyph(aesthetics, data, c("GLYPH", "PANEL"), plot$plot_env))
+}
+
+eval.glyph <- function (exprs, data = NULL, by = NULL, enclos = NULL, try = FALSE) {
+  if (is.numeric(exprs)) 
+    return(envir[exprs])
+  qenv <- if (is.quoted(exprs)) 
+    attr(exprs, "env")
+  else parent.frame()
+  if (is.null(data)) 
+    data <- qenv
+  if (is.data.frame(data) && is.null(enclos)) 
+    enclos <- qenv
+  if (try) {
+    results <- failwith(NULL, ddply, quiet = TRUE) (data, by, apply_glyphs, exprs, qenv)    
+  } else {
+    results <- ddply(data, by, apply_glyphs, exprs, qenv)    
+  }
+  results
+}
+
+apply_glyphs <- function(data, mapping, enclos = parent.frame()) {
+  map <- null_omit(mapping)
+  vars <- llply(map, eval, envir = data, enclos)
+  n <- nrow(data)
+  lengths <- unlist(lapply(vars, length))
+  wrong <- lengths != 1 & lengths != n
+  if (any(wrong)) {
+    stop(paste("Aesthetics must either be length one, or the same length as the data", "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
+  }
+  xy <- lengths[c("x", "y")]
+  wrong <- xy != 1
+  if (any(wrong)) {
+    vars[["x"]] <- vars[["x"]][1]
+    vars[["y"]] <- vars[["y"]][1]
+    message(paste("glyph mapped to multiple " , paste(names(wrong)[wrong], 
+      collapse = " and "), ". Using first value.", sep = ""))
+  }
+  data.frame(vars)
+}

### 2012-05-30 R/glyph_build.r

@@ -1,0 +1,115 @@
+#' Build a glyph object for rendering
+#'
+#' This function takes the plot object, and performs all steps necessary to produce an object that can be rendered.
+glyph_build <- function(plot){
+  if (length(plot$layers) == 0) stop("No layers in plot", call.=FALSE)
+  if (!identical(plot$facet, facet_null())) {
+  	stop("glyphs do not support facetting", call. = FALSE)
+  }
+	
+  plot <- ggplot2:::plot_clone(plot)
+  layers <- plot$layers
+  layers <- propogate_data(layers, plot$data)
+	
+  # separate into glayers and normal layers
+  gls <- unlist(lapply(layers, is.glayer))
+  if (all(!gls)) return(ggplot_build(plot))
+  if (all(gls) && sum(gls) == 1) return(glayer_build(layers[[gls]]))
+  glayers <- layers[gls]
+  plot$layers <- layers[!gls]
+  gl.order <- seq_along(layers)[gls]
+  nl.order <- seq_along(layers)[!gls]
+	
+  # build normal layers
+  normal <- NULL
+  if (length(plot$layers) > 0) {
+	normal <- ggplot_build(plot)
+  }
+	
+  # build glyph layers (embedded plots)
+  embedded <- list()
+  for (i in seq_along(glayers)) {
+    embedded[[i]] <- glayer_build(glayers[[i]])
+  }
+	
+	
+  ### combine the builds
+	
+  # plot
+  build <- embedded[[1]]
+	
+  # data
+  # take care to order
+  edata <- lapply(embedded, function(bd) bd$data[[1]])
+  data <- list()
+  data[gl.order] <- edata
+  data[nl.order] <- normal$data
+	
+  # panel
+  xspan <- range(unlist(lapply(data, function(df) df[names(df) %in% .x_aes])))
+  yspan <- range(unlist(lapply(data, function(df) df[names(df) %in% .y_aes])))
+  panel <- ggplot_build(qplot(xspan, yspan))$panel
+	
+  # scales 
+  # collect all unique scales
+  scales <- build$plot$scales$scales
+  scales[[which_x(scales)]] <- panel$x_scales[[1]]
+  scales[[which_y(scales)]] <- panel$y_scales[[1]]
+  scale.names <- names_scales(scales)
+  for (i in seq_along(embedded[-1])) {
+    escales <- embedded[[i + 1]]$plot$scales$scales
[…]

### 2012-05-30 R/glyphs-class.r

@@ -1,0 +1,77 @@
+#' @include ggtransform.r
+NULL
+
+check_glyphs <- function(object) {
+	errors <- character()
+	if (!is(object@.Data, "ggplot")) {
+		msg <- "glyphs must be a ggplot object."
+		errors <- c(errors, msg)
+	}
+	if (length(errors) == 0) 
+		TRUE
+	else
+		errors
+}
+
+setOldClass(c("ggplot", "list"))
+
+#' glyphs class
+#'
+#' a glyphs object is a ggplot object that has been extended to include methods for applying plyr when plotting. 
+#'
+#' @name glyphs-class
+#' @rdname glyphs-class
+#' @exportClass glyphs
+#' @aliases show,glyphs-method
+#' @aliases ggtransform,glyphs-method
+setClass("glyphs", contains = c("ggplot"), validity = check_glyphs)
+
+#' @export
+setMethod("show", signature(object = "glyphs"), function(object){
+	print(object)
+})
+
+#' @S3method print ggplyr
+print.glyphs <- function(x, newpage = is.null(vp), vp = NULL, ...) {
+    ggplot2:::set_last_plot(x)
+    if (newpage) 
+        grid.newpage()
+    data <- glyph_build(x)
+    gtable <- ggplot_gtable(data)
+    if (is.null(vp)) {
+        grid.draw(gtable)
+    }
+    else {
+        if (is.character(vp)) 
+            seekViewport(vp)
+        else pushViewport(vp)
+        grid.draw(gtable)
+        upViewport()
+    }
+    invisible(data)
+}
+
+#' @export
+setGeneric("ggtransform")
+
+#' @export
+setMethod("ggtransform", signature(ggobject = "glyphs"), 
+	function(ggobject, mapping, ...){
+		ggplot <- ggtransform(ggobject@.Data, mapping, ...)
+		new("glyphs", ggplot)
+	}
+)
+
+
+
+
+#' Create a glyphs object
+#' 
+#' glyph_plot gives a ggplot object the S4 class ggplyr, see \code{\link{glyphs-class}}. ggplyr denotes ggplot objects that contain extra information to be used to apply plyr functions when plotting.
+#' 
+#' @export glyph_plot
+#' @param ggplot a ggplot object
+glyph_plot <- function(ggplot) {
+	new("glyphs", ggplot)
+}
+

### 2012-05-30 R/map_glyph.r

@@ -1,0 +1,6 @@
+map_glyph <- function(data, layer){
+  if("map_glyphs" %in% ls(layer)) {
+    data <- layer$map_glyphs(data)
+  }
+  data
+}

### 2012-05-30 R/merge_overlaps.r

@@ -1,0 +1,20 @@
+merge_overlaps <- function(globals, width, height) {
+	globals <- arrange(globals, x)
+	g1 <- globals[1:(nrow(globals) - 1), ]
+	g2 <- globals[-1, ]
+
+	x.overlaps <- abs(g1$x - g2$x) < width 
+	y.overlaps <- abs(g1$y - g2$y) < height
+	overlaps <- x.overlaps & y.overlaps
+	
+	# fix multiple overlaps
+	new = g1$GLYPH[overlaps]
+	old = g2$GLYPH[overlaps]
+	still.old <- which(new %in% old)
+	while(length(still.old)) {
+		new[still.old[1]] <- new[which(old == new[still.old[1]])]
+		still.old <- which(new %in% old)
+	}
+	names(new) <- old
+	new
+}

### 2012-05-30 R/ply_aes.r

@@ -1,0 +1,20 @@
+#' ply_aes causes the aesthetics of a layer to be computed groupwise.
+ply_aes <- function(layer, .vars) {
+  if (is.glayer(layer)) return(layer)	
+  layer <- layer_clone(layer)
+  layer$ply.by <- .vars
+  layer$compute_aesthetics <- function (., data, plot) {
+    data$.gid <- id(data[ply.by], drop = TRUE)	
+    aesthetics <- .$layer_mapping(plot$mapping)
+    if (!is.null(.$subset)) {
+      include <- data.frame(eval.quoted(.$subset, data, plot$env))
+      data <- data[rowSums(include, na.rm = TRUE) == ncol(include), ]
+    }
+    if (!is.null(.$geom_params$group)) {
+      aesthetics["group"] <- .$geom_params$group
+    }
+    scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
+    compact(eval.plyr(aesthetics, data, c(".gid", "PANEL"), plot$plot_env))
+  }
+  layer
+} 

### 2012-05-30 R/plyr_aes.r

@@ -1,0 +1,44 @@
+plyr_aesthetics <- function (., data, plot) {
+    aesthetics <- .$layer_mapping(plot$mapping)
+    if (!is.null(.$subset)) {
+        include <- data.frame(eval.quoted(.$subset, data, plot$env))
+        data <- data[rowSums(include, na.rm = TRUE) == ncol(include), ]
+    }
+    if (!is.null(.$geom_params$group)) {
+        aesthetics["group"] <- .$geom_params$group
+    }
+    scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
+    compact(eval.plyr(aesthetics, data, c("GLYPH", "PANEL"), plot$plot_env))
+}
+
+eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, try = FALSE) {
+    if (is.numeric(exprs)) 
+        return(envir[exprs])
+    qenv <- if (is.quoted(exprs)) 
+        attr(exprs, "env")
+    else parent.frame()
+    if (is.null(data)) 
+        data <- qenv
+    if (is.data.frame(data) && is.null(enclos)) 
+        enclos <- qenv
+    if (try) {
+        results <- failwith(NULL, ddply, quiet = TRUE) (data, by, apply_maps, 
+          exprs, qenv)    
+    } else {
+        results <- ddply(data, by, apply_maps, exprs, qenv)    
+    }
+    results
+}
+
+
+apply_maps <- function(data, mapping, enclos = parent.frame()) {
+	map <- null_omit(mapping)
+	vars <- llply(map, eval, envir = data, enclos)
+	n <- nrow(data)
+	lengths <- unlist(lapply(vars, length))
+	wrong <- lengths != 1 & lengths != n
+	if (any(wrong)) {
+        stop(paste("Aesthetics must either be length one, or the same length as the data", "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
+    }
+  data.frame(vars)
+}

### 2012-05-30 R/ref_box.r

@@ -1,0 +1,73 @@
+ref_box <- function(mapping = NULL, fill = "grey80", ...) {	
+  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
+  height = rel(1), merge.overlaps = FALSE) {
+  	
+  	def_aes <- aes(xmin = -1, xmax = 1, ymin = -1, ymax = 1)
+  	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
+  	class(mapping) <- "uneval"
+  	
+  	rlayer <- geom_rect(mapping = mapping, ...)
+  	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
+  	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
+  	
+  	glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, width = width,
+  	  height = height, merge = merge.overlaps, ref = NULL, .ref = TRUE)
+  }
+} 
+
+ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {	
+  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
+  height = rel(1), merge.overlaps = FALSE) {
+
+  	def_aes <- list(xmin = -1, xmax = 1, ymin = -width./2, ymax = width./2)
+  	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
+  	class(mapping) <- "uneval"
+  	
+  	rlayer <- geom_rect(mapping = mapping, ...)
+  	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
+  	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
+  	
+  	glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, width = width,
+  	  height = height, merge = merge.overlaps, ref = NULL, .ref = TRUE)
+  }
+} 
+
+ref_vline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {	
+  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
+  height = rel(1), merge.overlaps = FALSE) {
+  	
+  	def_aes <- list(xmin = -width./2, xmax = width./2, ymin = -1, ymax = 1)
+  	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
[…]

### 2012-05-30 R/rescalers.r

@@ -1,0 +1,36 @@
+rescale_01 <- free <- function(xvars, xlim=NULL) {
+	xnames <- names(xvars)
+	numberfy <- function(x) {
+		if (is.character(x)) {
+			x <- as.numeric(factor(x))
+		}	
+		if (is.factor(x)) {
+			x <- as.numeric(x)
+		}
+		x
+	}
+	xvars <- lapply(xvars, numberfy)
+  
+	if (is.null(xlim)) {
+		rng <- range(unlist(xvars), na.rm = TRUE)
+	} else {
+		rng <- xlim
+	}
+	
+	scale <- function(x) {
+		if ((rng[2] - rng[1]) == 0) {
+			x - rng[1]
+		} else {	
+			(x - rng[1]) / (rng[2] - rng[1])
+		}
+	}
+	xvars <- lapply(xvars, scale)
+	names(xvars) <- xnames
+	data.frame(xvars)
+}
+
+
+rescale_11 <- function(xvars, xlim=NULL) {
+	2 * rescale_01(xvars, xlim) - 1
+}
+

### 2012-05-30 R/util.r

@@ -1,0 +1,37 @@
+.x_aes <- c("x", "xend", "xmax", "xmin")
+.y_aes <- c("y", "yend", "ymax", "ymin")
+
+#' get_xs retrieves all mappings that need to be altered for plotting on a new x axis
+#'
+#' @keywords internal
+#' @param data a data.frame
+#' @export
+get_xs <- function(data) {
+  names(data)[names(data) %in% .x_aes]
+}
+
+#' get_ys retrieves all mappings that need to be altered for plotting on a new y axis
+#'
+#' @keywords internal
+#' @param data a data.frame
+#' @export
+get_ys <- function(data) {
+  names(data)[names(data) %in% .y_aes]
+}
+
+
+layer_clone <- function(layer) {
+  ggplot2:::plot_clone(ggplot() + layer)$layers[[1]]
+}
+
+#' null.omit removes the NULL elements from a list and returns the remaining objects as a more concise list.
+#' 
+#' @keywords internal
+#' @param lst a list
+#' @export
+null_omit <- function(lst) {
+  if (is.null(lst)) {
+    return(NULL)
+  }
+  lst[!(unlist(lapply(lst, is.null)))]
+}

### 2012-05-30 demo.r

@@ -1,0 +1,100 @@
+###################################################
+###             geom_scatterplots               ###
+###################################################
+
+
+###########################################
+###              working                ###
+###########################################
+ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature), 
+  size = 1/5), glyph.by = c("lat", "long"), width = 3, height = 3, 
+  major = aes(x = long[1], y = lat[1]))
+
+# embed points (continuous - no stats)
+ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
+  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
+  major = aes(Fertility, Education))
+
+# embed bars (categorical with stats)
+ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
+  aes(x = mean(displ), y = mean(cty)), c("year"), y_scale = free, 
+  width = 1/4, height = 1/4)
+  
+# reference boxes
+# boxes
+ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
+  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
+  major = aes(Fertility, Education), ref = ref_box(aes(fill = 
+  mean(Catholic))))
+
+# hlines
+ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
+  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
+  major = aes(mean(Fertility), mean(Education)), ref = ref_hline(aes(fill = 
+  mean(Catholic))))
+
+# vlines
+ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
+  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
+  major = aes(mean(Fertility), mean(Education)), ref = ref_vline(aes(fill = 
+  mean(Catholic))))
+  
+# points
+ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
+  aes(x = mean(displ), y = mean(cty)), c("year"), y_scale = free, 
+  width = 1/3, height = 1/3, reference = ref_points(aes(fill = mean(hwy)), 
+  size = 1)) 
+
+# merging overlaps
[…]

### 2012-05-30 scrap-code/glyph-1.r

@@ -1,0 +1,121 @@
+#' glyph turns an ordinary layer into a set of glyphs. To do this, it must:
+#' 1. ensure that the x and y aesthetics are mapped to a single value
+#' 2. add reference boxes
+#' 3. handle merging
+#' 4. split up the data
+glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(1), height = rel(1), 
+  x_scale = identity, y_scale = identity, merge.overlaps = FALSE, reference = NULL) {
+  
+  missing <- c(is.null(major.aes$x), is.null(major.aes$y))
+  if (any(missing)) {
+    stop(paste("Missing required aesthetic in major.aes:", 
+      paste(c("x", "y")[missing], collapse = ", ")))
+  }
+  
+  layer <- layer_clone(layer)
+  
+  # update aesthetics, save glyph params
+  mnames <- names(layer$mapping)
+  mnames[mnames == "x"] <- "minor.x"
+  mnames[mnames == "y"] <- "minor.y"
+  names(layer$mapping) <- mnames
+  layer$mapping$x <- major.aes$x
+  layer$mapping$y <- major.aes$y
+  layer$width <- width
+  layer$height <- height
+  layer$glyph.by <- glyph.by
+  layer$x_scale <- x_scale
+  layer$y_scale <- y_scale
+  layer$merge.overlaps <- merge.overlaps
+  
+  layer$assign_glyphs <- function(., data){
+    data$GLYPH <- id(data[glyph.by], drop = TRUE)
+    # parse width, height
+    if (is.rel(width) || is.rel(height) || merge.overlaps) {
+      global_aes <- list(x = mapping$x, y = mapping$y)
+      globals <- ddply(data, "GLYPH", apply_glyphs, global_aes)
+    }
+    if (is.rel(width)) {
+      .$width <- (max(diff(range(globals$x)), 1) + unclass(width)) / 
+        length(unique(globals$x)) * unclass(width)
+    }
+    if (is.rel(height)) {
+      .$height <- (max(diff(range(globals$y)), 1) + unclass(height)) / 
+        length(unique(globals$y)) * unclass(height)
+    }
+  
+    # split data
+    if (merge.overlaps) {
+      merges <- merge_overlaps(globals, width, height)
+      data$GLYPH <- update_GLYPH(data$GLYPH, merges)
+    }
+    data
+  }
+  
+  
+  # update build time functions
[…]

### 2012-05-30 scrap-code/glyph_build-1.r

@@ -1,0 +1,81 @@
+#' this might work for normal layers too? yes. confirmed!
+glyph_build <- function (plot) {
+    if (length(plot$layers) == 0) 
+        stop("No layers in plot", call. = FALSE)
+    plot <- ggplot2:::plot_clone(plot)
+    layers <- plot$layers
+    layer_data <- lapply(layers, function(y) y$data)
+    scales <- plot$scales
+    dlapply <- function(f) {
+        out <- vector("list", length(data))
+        for (i in seq_along(data)) {
+            out[[i]] <- f(d = data[[i]], p = layers[[i]])
+        }
+        out
+    
+        }
+    panel <- ggplot2:::new_panel()
+    panel <- ggplot2:::train_layout(panel, plot$facet, layer_data, plot$data)
+    data <- ggplot2:::map_layout(panel, plot$facet, layer_data, plot$data)
+    data <- dlapply(function(d, p) add_glyph(d, p))
+    
+    data <- dlapply(function(d, p) p$compute_aesthetics(d, plot))
+    
+    data <- lapply(data, ggplot2:::add_group)
+    data <- lapply(data, ggplot2:::scales_transform_df, 
+    	scales = scales)
+    scale_x <- function() scales$get_scales("x")
+    scale_y <- function() scales$get_scales("y")
+    panel <- ggplot2:::train_position(panel, data, scale_x(), scale_y())
+    data <- ggplot2:::map_position(panel, data, scale_x(), scale_y())
+    
+    data <- dlapply(function(d, p) name_majors(d, p))
+    data <- plyr_stats(panel, data, layers) #slow. can we skip?
+    data <- dlapply(function(d, p) p$map_statistic(d, plot))
+    data <- dlapply(function(d, p) name_minors(d, p))
+    data <- lapply(data, ggplot2:::order_groups)
+    
+    npscales <- scales$non_position_scales()
+    if (npscales$n() > 0) {
+        lapply(data, ggplot2:::scales_train_df, scales = npscales)
+        data <- lapply(data, ggplot2:::scales_map_df, 
+        	scales = npscales)
+    }
+    
+    data <- dlapply(function(d, p) p$reparameterise(d))
+    data <- dlapply(function(d, p) p$adjust_position(d))
+    data <- dlapply(function(d, p) map_glyph(d, p))
+    
+    ggplot2:::reset_scales(panel)
[…]

### 2012-05-30 R/aesply.r
rewrote how plyr handles aesthetics and global variables, created grid and grid_by and aesply

@@ -1,0 +1,31 @@
+aesply <- function(data, .var, aesthetics) {
+  no.ply <- unlist(lapply(aesthetics, first_name)) == "I"
+  if (!any(no.ply)) {
+    return(compact(eval.plyr(aesthetics, data, .var)))
+  }
+  
+  all.aes <- lapply(aesthetics[no.ply], remove_I)
+  evaled <- compact(eval.quoted(all.aes, data))
+  lengths <- vapply(evaled, length, integer(1))
+  n <- if (length(lengths) > 0) max(lengths) else 0
+  wrong <- lengths != 1 & lengths != n
+  if (any(wrong)) {
+    stop("Aesthetics must either be length one, or the same length as the data", 
+         "Problems:", paste(aesthetics[wrong], collapse = ", "), 
+         call. = FALSE)
+  }
+  data <- cbind(data, evaled)
+  all.data <- data.frame(evaled) 
+  aesthetics[no.ply] <- lapply(names(aesthetics)[no.ply], as.name)
+  compact(eval.plyr(aesthetics, data, .var))
+}
+
+
+remove_I <- function(expr) {
+  Identity <- function(x) {
+    if (x == "I") x <- quote(identity)
+    x
+  }
+  as.call(lapply(expr, Identity))
+}
+

### 2012-05-30 R/grid_by.r

@@ -1,0 +1,31 @@
+#' grid_by is a second order function: it returns a function that can be used on 
+#' a data set to provide a grouping variable.
+grid_by <- function(x, x.npoints = 10, y, y.npoints = 10, x.range. = NULL, 
+  y.range. = NULL) {
+  
+  if(inherits(try(x, silent = TRUE), "try-error")) {
+    x <- as.character(substitute(x))
+  }
+  if(inherits(try(y, silent = TRUE), "try-error")) {
+    y <- as.character(substitute(y))
+  }
+  
+  function(df) {
+    missing <- !(c(x, y) %in% names(df))
+    if (any(missing)) {
+      stop(paste(paste(c(x, y)[missing], collapse = ", "), 
+        "variables not found in data.frame"), call. = FALSE)
+    }
+    if (is.null(x.range.)) {
+      df$.X_GRID <- grid(df[[x]], x.npoints)
+    } else {
+      df$.X_GRID <- grid(df[[x]], x.npoints, x.range.)
+    }
+    if (is.null(y.range.)) {
+      df$.Y_GRID <- grid(df[[y]], y.npoints)
+    } else {
+      df$.Y_GRID <- grid(df[[y]], y.npoints, y.range.)
+    }
+    id(df[c(".X_GRID", ".Y_GRID")], drop = TRUE)
+  }
+}

### 2012-05-30 R/group_by.r

@@ -1,0 +1,15 @@
+#' group_by is a second order function: it returns a function that can be used on a data set to provide a grouping variable.
+group_by <- function(vars) {
+  if (!is.character(vars)) {
+    stop("vars should be of class 'character'", call. = FALSE)
+  }
+  
+  function(df) {
+    missing <- !(vars %in% names(df))
+    if (any(missing)) {
+      stop(paste(paste(vars[missing], collapse = ", "), 
+        "variables not found in data.frame"), call. = FALSE)
+    }
+    id(df[vars], drop = TRUE)
+  }
+}

### 2012-05-30 R/plyr_aes.r

@@ -1,15 +1,23 @@
+#' what replaces compute_aesthetics
 plyr_aesthetics <- function (., data, plot) {
-    aesthetics <- .$layer_mapping(plot$mapping)
-    if (!is.null(.$subset)) {
-        include <- data.frame(eval.quoted(.$subset, data, plot$env))
-        data <- data[rowSums(include, na.rm = TRUE) == ncol(include), ]
+  aesthetics <- .$layer_mapping(plot$mapping)
+  if (!is.null(.$subset)) {
+    include <- data.frame(eval.quoted(.$subset, data, plot$env))
+    data <- data[rowSums(include, na.rm = TRUE) == ncol(include), ]
+  }
+  if (!is.null(.$geom_params$group)) {
+    aesthetics["group"] <- .$geom_params$group
+  }
+  scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
+  
+  if (is.null(data$GLYPH)) {
+    if ("plyr" %in% ls(.)) {
+      data$GLYPH <- plyr$ply.by(data)
     }
-    if (!is.null(.$geom_params$group)) {
-        aesthetics["group"] <- .$geom_params$group
-    }
-    scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
-    compact(eval.plyr(aesthetics, data, c("GLYPH", "PANEL"), plot$plot_env))
+  }
+  aesply(data, c("GLYPH", "PANEL"), aesthetics)
 }
+
 
 eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, try = FALSE) {
     if (is.numeric(exprs)) 
@@ -38,7 +46,9 @@ apply_maps <- function(data, mapping, enclos = parent.frame()) {
 	lengths <- unlist(lapply(vars, length))
 	wrong <- lengths != 1 & lengths != n
 	if (any(wrong)) {
-        stop(paste("Aesthetics must either be length one, or the same length as the data", "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
+    stop(paste(
+      "Aesthetics must either be length one, or the same length as the data", 
+      "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
     }
   data.frame(vars)
 }

### 2012-05-30 R/ref_box.r

@@ -6,7 +6,7 @@ ref_box <- function(mapping = NULL, fill = "grey80", ...) {
   	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
   	class(mapping) <- "uneval"
   	
-  	rlayer <- geom_rect(mapping = mapping, ...)
+    rlayer <- ply_aes(geom_rect(mapping = mapping, ...))
   	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
   	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
   	
@@ -15,7 +15,9 @@ ref_box <- function(mapping = NULL, fill = "grey80", ...) {
   }
 } 
 
-ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {	
+ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ..., 
+  ply = TRUE) {	
+  
   function(layer, major.aes, glyph.by = NULL, width = rel(1), 
   height = rel(1), merge.overlaps = FALSE) {
 
@@ -23,7 +25,11 @@ ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {
   	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
   	class(mapping) <- "uneval"
   	
-  	rlayer <- geom_rect(mapping = mapping, ...)
+  	if (ply) {
+  	  rlayer <- ply_aes(geom_rect(mapping = mapping, ...), glyph.by)
+  	} else {
+  	  rlayer <- geom_rect(mapping = mapping, ...)
+  	}
   	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
   	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
   	
@@ -32,7 +38,9 @@ ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {
   }
 } 
 
-ref_vline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {	
+ref_vline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ..., 
+  ply = TRUE) {	
+  
   function(layer, major.aes, glyph.by = NULL, width = rel(1), 
   height = rel(1), merge.overlaps = FALSE) {
   	
[…]

### 2012-05-30 R/util.r

@@ -1,6 +1,17 @@
 .x_aes <- c("x", "xend", "xmax", "xmin")
 .y_aes <- c("y", "yend", "ymax", "ymin")
 
+#' retrieves the first name used in an expression. To be used with screening 
+#' mappings for plyr or regular computation.
+first_name <- function(expr) {
+  names <- all.names(expr)
+  names <- names[names != "["]
+  firsts <- names[1]
+  firsts[is.na(firsts)] <- ""
+  firsts
+}
+
+
 #' get_xs retrieves all mappings that need to be altered for plotting on a new x axis
 #'
 #' @keywords internal

### 2012-05-30 demo.r

@@ -1,19 +1,28 @@
 ###################################################
 ###             geom_scatterplots               ###
 ###################################################
-
-
+library(devtools)
+load_all("../ggplyr")
 ###########################################
 ###              working                ###
 ###########################################
-ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature), 
+# imac 10.274 seconds
+# system.time(print(qplot(surftemp, temperature, data = nasa) + facet_grid(long~lat)))
+system.time(print(ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature), 
   size = 1/5), glyph.by = c("lat", "long"), width = 3, height = 3, 
-  major = aes(x = long[1], y = lat[1]))
+  major = aes(x = long[1], y = lat[1]))))
+# compare to 21.736 seconds
+# system.time(print(qplot(surftemp, temperature, data = nasa, size = I(1/5)) + 
+#  facet_grid(long~lat)))
 
 # embed points (continuous - no stats)
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
-  major = aes(Fertility, Education))
+  major = aes(mean(Fertility), mean(Education)))
+
+ggplot(test.data) + glyph(ply_aes(geom_point(aes(Fertility, Agriculture, 
+  color = rank(Catholic)), size = 3), .vars = c("lat", "long")), 
+  glyph.by = c("lat", "long"), major = aes(mean(Fertility), mean(Education)))
 
 # embed bars (categorical with stats)
 ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
@@ -24,7 +33,7 @@ ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)),
 # boxes
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
-  major = aes(Fertility, Education), ref = ref_box(aes(fill = 
+  major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
   mean(Catholic))))
 
 # hlines
@@ -89,9 +98,21 @@ ggplot(mpg) + ply_aes(geom_point(aes(hwy, cty, color = rank(displ))),
[…]

### 2012-06-01 paper/demo.Rnw
Changed merge overlaps method. Added seasonal data from glyph-maps paper.

@@
 
-\title{A Minimal Demo of ggplyr/glyphmaps/modelglyphs}
+\title{A Minimal Demo of \texttt{glyphmaps}}
 
@@
 \maketitle
-This document demonstrates the basic features of the \texttt{ggplyr} package (name may change).
-\texttt{ggplyr} was written to create plots like the ones below.
-Each plot was made with \texttt{ggplyr}.
-The first two plots recreate plots from \citet{wickham2011split}.
-The last two recreate plots from \citet{wickham2012glyph}.
+This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change).
+\texttt{glyphmaps} was written to create plots like the ones below.
+Each plot is made with \texttt{glyphmaps}.
+They recreate Figures 8 and 10a from \citet{wickham2012glyph}.
+To run the examples in this document please download the \texttt{glyphmaps} package with the code below.
 
-<<plyr star glyphs>>=
+<<install>>=
+library(devtools)
+install_github("ggplyr", "garrettgman", "glyphmaps")
+library(glyphmaps)
 @
 
-<<plyr filled star glyphs>>=
-@
-
-<<glyph maps grided>>=
+<<glyphmaps-fig-8>>=
+p <- ggplot(seasons) + 
+  map_us +
+  glyph(
+    geom_line(aes(x = time, y = pred)), 
+    major = aes(lon[1], lat[1]), glyph.by = "stn", 
+    ref = ref_box(aes(fill = avg)))
 @

### 2012-06-01 R/glayer_build.r

@@ -7,7 +7,7 @@ glayer_build <- function(layer) {
   layer <- layer_clone(layer)
   layer$data <- layer$assign_glyphs(layer$data)
   minor <- ggplot_build(ggplot() + layer + facet_wrap("GLYPH")) 
-	
+
   ### combine subplots (minor) into single plot
   # data
   data <- unpanel(minor$data[[1]])
@@ -41,6 +41,9 @@ glayer_build <- function(layer) {
 
 
 unpanel <- function(df) {
+  if (!is.null(df$group)) {
+    df$group <- interaction(df$group, df$PANEL)
+  } 
   df$GLYPH <- as.numeric(as.character(df$PANEL))
   df$PANEL <- NULL
   df

### 2012-06-01 R/glyph.r

@@ -55,21 +55,29 @@ assign_glyphs <- function(., data) {
   # parse width, height
   width <- embed$width
   height <- embed$height
-
   if (is.rel(width)) {
-    .$embed$width <- (diff(range(vet(globals$x))) + unclass(width)) / 
-      length(unique(globals$x)) * unclass(width)
+    .$embed$width <- width <- (diff(range(vet(globals$x))) + unclass(width)) / 
+      sqrt(length(unique(globals$x))) * unclass(width)
   }
   if (is.rel(height)) {
-    .$embed$height <- (diff(range(vet(globals$y))) + unclass(height)) / 
-      length(unique(globals$y)) * unclass(height)
+    .$embed$height <- height <- (diff(range(vet(globals$y))) + unclass(height)) / 
+      sqrt(length(unique(globals$y))) * unclass(height)
   }
 
   if (embed$merge) {
     # search for overlapping glyphs, combine
-    merges <- merge_overlaps(globals, embed$width, embed$height)
-    data <- update_GLYPH(data, merges)
-    embed$globals <- ddply(data, "GLYPH", apply_glyphs, embed$major.aes)
+    data$.gid <- data$GLYPH
+    data$GLYPH <- merge_overlaps(globals, embed$width, embed$height)
+    globals <- aesply(data, "GLYPH", embed$major.aes)
+    .$mapping <- add_gid(mapping)
+    
+    too.many <- c(length(unique(globals$x)) > length(unique(globals$GLYPH)), 
+                  length(unique(globals$y)) > length(unique(globals$GLYPH)))
+    if (any(too.many)) {
+      message(paste("Major", paste(c("x", "y")[too.many], collapse = " and "), 
+                    "return more than one value per glyph. Only using first."))
+      globals <- unique(ddply(globals, "GLYPH", transform, x = x[1], y = y[1]))
+    }
   }
   .$embed$globals <- globals
   data
@@ -147,13 +155,17 @@ globalize <- function(obj){
 	obj
 }
 
-update_GLYPH <- function(df, updates) {
-	if (!("GLYPH" %in% names(df))) {
-		stop("Cannot update glyphs: data.frame does not have GLYPH variable", 
-			call. = FALSE)
-	}
-	
-	obsolete <- df$GLYPH %in% names(updates)
-	df$GLYPH[obsolete] <- updates[as.character(df$GLYPH[obsolete])]
-	df
+
[…]

### 2012-06-01 R/merge_overlaps.r

@@ -1,20 +1,14 @@
 merge_overlaps <- function(globals, width, height) {
-	globals <- arrange(globals, x)
-	g1 <- globals[1:(nrow(globals) - 1), ]
-	g2 <- globals[-1, ]
+  x.overlaps <- abs(outer(globals$x, globals$x, "-")) < width
+  y.overlaps <- abs(outer(globals$y, globals$y, "-")) < height
+  overlaps <- data.frame(x.overlaps & y.overlaps)
+  names(overlaps) <- as.character(globals$GLYPH)
 
-	x.overlaps <- abs(g1$x - g2$x) < width 
-	y.overlaps <- abs(g1$y - g2$y) < height
-	overlaps <- x.overlaps & y.overlaps
-	
-	# fix multiple overlaps
-	new = g1$GLYPH[overlaps]
-	old = g2$GLYPH[overlaps]
-	still.old <- which(new %in% old)
-	while(length(still.old)) {
-		new[still.old[1]] <- new[which(old == new[still.old[1]])]
-		still.old <- which(new %in% old)
-	}
-	names(new) <- old
-	new
+  for (i in seq_along(overlaps)) {
+    names(overlaps)[overlaps[[i]]] <- names(overlaps)[i]
+  }
+  vec <- as.numeric(factor(names(overlaps)))
+  names(vec) <- globals$GLYPH
+  vec
 }
+  

### 2012-06-01 R/ref_box.r

@@ -11,7 +11,7 @@ ref_box <- function(mapping = NULL, fill = "grey80", ...) {
   	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
   	
   	glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, width = width,
-  	  height = height, merge = merge.overlaps, ref = NULL, .ref = TRUE)
+  	  height = height, merge = merge.overlaps, ref = NULL, .ref = TRUE) 
   }
 } 
 

### 2012-06-01 demo.r

@@ -8,9 +8,9 @@ load_all("../ggplyr")
 ###########################################
 # imac 10.274 seconds
 # system.time(print(qplot(surftemp, temperature, data = nasa) + facet_grid(long~lat)))
-ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature), 
+system.time(print(ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature), 
   size = 1/5), glyph.by = c("lat", "long"), width = 3, height = 3, 
-  major = aes(x = long[1], y = lat[1]))
+  major = aes(x = long[1], y = lat[1]))))
 # compare to 21.736 seconds
 # system.time(print(qplot(surftemp, temperature, data = nasa, size = I(1/5)) + 
 #  facet_grid(long~lat)))
@@ -151,3 +151,9 @@ p <- ggplot(mpg) +
 mpg$lat <- sample(1:4, nrow(mpg), replace = TRUE)
 ggplot(mpg) + GeomCoxcomb$new(mapping = aes(r = trans, fill = lat, group = lat)) + facet_wrap(~cyl)
 
+ggplot(seasons) + 
+  glyph(
+    geom_line(aes(x = time, y = pred)), 
+    major = aes(lon[1], lat[1]), glyph.by = "stn", 
+    height = rel(4.375), width = rel(1.9),
+    ref = ref_box(aes(fill = avg)), merge = TRUE)

### 2012-06-03 paper/demo.Rnw
wrote new grid() function that works like glyph() to turn a layer into a gridded layer.

@@
     geom_line(aes(x = time, y = pred)), 
-    major = aes(lon[1], lat[1]), glyph.by = "stn", 
+    major = aes(I(grid(lon)),I(grid(lat))), glyph.by = grid_by(lon, lat))
+, 
     ref = ref_box(aes(fill = avg)))

### 2012-06-03 R/glyph.r

@@ -33,7 +33,7 @@ glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95),
   if (is.null(reference)) {
   	glayer(layer)
   } else {
-  	ref.layer <- reference(layer, major.aes, glyph.by, width, height, 
+  	ref.layer <- reference(layer, "glyph", major.aes, glyph.by, width, height, 
   	  merge.overlaps)
   	list(ref.layer, glayer(layer))
   }
@@ -156,7 +156,10 @@ globalize <- function(obj){
 }
 
 
-#' add_gid intelligently adds the .gid variable to the group slot of an uneval object. If the group slot is NULL, add_gid sets group = .gid. If the group slot already contains a mapping, add_gid adds .gid to this mapping with interaction().
+#' add_gid intelligently adds the .gid variable to the group slot of an uneval 
+#' object. If the group slot is NULL, add_gid sets group = .gid. If the group 
+#' slot already contains a mapping, add_gid adds .gid to this mapping with 
+#' interaction().
 #'
 #' @keywords internal
 #' @param aes_group the group value of an uneval object
@@ -165,7 +168,8 @@ add_gid <- function(maps) {
   if (is.null(maps$group)) {
     maps$group <- as.name(".gid")
   } else {
-    maps$group <- as.call(list(as.name("interaction"), as.name(".gid"), aes_group))
+    maps$group <- as.call(list(as.name("interaction"), as.name(".gid"), 
+      aes_group))
   }
   maps
 }

### 2012-06-03 R/grid.r

@@ -1,10 +1,74 @@
-#' grid takes a vector of numbers, divides it into n bins, and rounds each point to the nearest center of a bin.
-grid <- function(x, npoints = 10, range. = range(x)) {
-  if (diff(range.) == 0) return(x)
-  binwidth <- diff(range.) / npoints
-  x <- x - range.[1] + binwidth / 2
-  a <- x %/% binwidth + c(0,1)[round(x %% binwidth / binwidth) + 1]
-  a[a == 0] <- 1 # include leftmost point
-  x <- range.[1] + binwidth * a
-  x - binwidth/2
+#' grid turns an ordinary layer into a gridded version of itself. 
+grid <- function(layer, grid.aes = aes(), x.nbin = 10, y.nbin = 10, 
+  x_scale = identity, y_scale = identity, width.adjust = 0.95, 
+  height.adjust = 0.95, reference = NULL, .ref = FALSE) {
+  
+  if (is.glayer(layer) || is.list(layer)) {
+    stop("Cannot grid glayer", call. = TRUE)
+  }
+  if ("plyr" %in% ls(layer)) {
+    stop("grid and ply_aes used in wrong order.", call. = FALSE)
+  }
+  
+  # aes
+  if(is.null(grid.aes$x)) {
+    if (is.null(layer$mapping$x)) {
+      stop("Grid requires an x aesthetic to be specified")
+    }
+    grid.aes$x <- layer$mapping$x
+  }
+  if(is.null(grid.aes$y)) {
+    if (is.null(layer$mapping$y)) {
+      stop("Grid requires a y aesthetic to be specified")
+    }
+    grid.aes$y <- layer$mapping$y
+  }
+    
+  
+  layer <- layer_clone(layer)
+  layer$embed <- list(x_scale = x_scale, y_scale = y_scale,
+    major.aes = grid.aes[c("x", "y")], grid_env = parent.frame(),
+    x.nbin = x.nbin, y.nbin = y.nbin, w.adjust = width.adjust, 
+    h.adjust = height.adjust)
+  layer$assign_glyphs <- assign_grid
+  layer$combine_glyphs <- combine_glyphs
+  if (.ref) layer$combine_glyphs <- combine_refs
+  #layer$compute_aesthetics <- plyr_aesthetics
+  
+  if (is.null(reference)) {
+    glayer(layer)
+  } else {
[…]

### 2012-06-03 R/grid_by.r

@@ -1,6 +1,6 @@
 #' grid_by is a second order function: it returns a function that can be used on 
 #' a data set to provide a grouping variable.
-grid_by <- function(x, x.npoints = 10, y, y.npoints = 10, x.range. = NULL, 
+grid_by <- function(x, y, x.npoints = 10, y.npoints = 10, x.range. = NULL, 
   y.range. = NULL) {
   
   if(inherits(try(x, silent = TRUE), "try-error")) {

### 2012-06-03 R/ref_box.r

@@ -1,6 +1,6 @@
-ref_box <- function(mapping = NULL, fill = "grey80", ...) {	
-  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
-  height = rel(1), merge.overlaps = FALSE) {
+ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {	
+  function(layer, type, major.aes, glyph.by = NULL, width = rel(1), 
+  height = rel(1), merge.overlaps = FALSE, x.nbin = 10, y.nbin = 10) {
   	
   	def_aes <- aes(xmin = -1, xmax = 1, ymin = -1, ymax = 1)
   	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
@@ -8,14 +8,22 @@ ref_box <- function(mapping = NULL, fill = "grey80", ...) {
   	
     rlayer <- ply_aes(geom_ref_rect(mapping = mapping, ...))
   	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
+  	if (is.null(mapping$colour)) rlayer$geom_params$colour <- color
   	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
   	
-  	glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, width = width,
-  	  height = height, merge = merge.overlaps, ref = NULL, .ref = TRUE) 
+    switch(type,
+      glyph = glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, 
+        width = width, height = height, merge = merge.overlaps, ref = NULL, 
+        .ref = TRUE), 
+      grid = grid(rlayer, grid.aes = major.aes, x.nbin = x.nbin, 
+        y.nbin = y.nbin, width.adjust = 1, height.adjust = 1, .ref = TRUE)
+    )
   }
 } 
 
-ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {	
+
+
+ref_hline <- function(mapping = NULL, width. = 0.2, fill = "white", ...) {	
   function(layer, major.aes, glyph.by = NULL, width = rel(1), 
   height = rel(1), merge.overlaps = FALSE) {
 
@@ -27,12 +35,17 @@ ref_hline <- function(mapping = NULL, width. = 0.2, fill = "grey80", ...) {
   	if (is.null(mapping$fill)) rlayer$geom_params$fill <- fill
[…]

### 2012-06-03 demo.r

@@ -105,16 +105,6 @@ ggplot() + ply_aes(geom_point(aes(x = mean(temperature), y = mean(ozone),
 ggplot() + geom_point(aes(x = mean(temperature), y = mean(ozone), 
   color = lat[1]), data = nasa)  
 
-# splitting by grids
-ggplot(test.data) + geom_point(aes(Fertility, Agriculture, color = Catholic))
-ggplot(test.data) + geom_point(aes(grid(Fertility, 3), grid(Agriculture, 3), 
-  color = Catholic))
-ggplot(test.data) + geom_point(aes(grid(Fertility, 3), grid(Agriculture, 3), 
-  color = Catholic), position = position_jitter(width = 1, height = 1))
-ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
-  color = Catholic)), aes(I(grid(Fertility, 3)), I(grid(Agriculture, 3))), 
-  grid_by(Fertility, 3, Agriculture, 3))
-
 # geom_scatterplots
 ggplot(nasa) + geom_scatterplots(mapping = aes(x = long[1], y = lat[1], 
   minor.x = surftemp, minor.y = temperature), glyph.by = c("long", "lat"), 
@@ -126,7 +116,7 @@ ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature),
 
 # trying out geom_star
 # without glyphing
-ggplot(mpg) + GeomStar$new(mapping = aes(r = hwy, angle = cty)) + 
+ggplot(mpg) + GeomStar$new(mapping = aes(r = hwy, angle = cty, group = cyl)) + 
   facet_wrap(~cyl)
 
 ggplot(test.data) + geom_star(mapping = aes(x = long[1], y = lat[1], 
@@ -139,7 +129,18 @@ ggplot(mpg) + ply_aes(geom_star(mapping = aes(x = cyl[1], y = 1,
 ggplot(nasa) + geom_star(aes(r = ozone, angle = date, x = long[1], y = lat[1]), 
   glyph.by = c("long", "lat"))
 
+# new gridding
+ggplot(test.data) + geom_point(aes(Fertility, Education))
+ggplot(test.data) + 
+  grid(geom_point(aes(Fertility, Education)), x.nbin = 10, y.nbin = 10, 
+       ref= ref_box(aes(color = mean(Catholic))))
 
+cheap.diamonds <- subset(diamonds, price <= 5000 & price >= 600)
[…]

### 2012-06-03 scrap-code/grid-1.r

@@ -1,0 +1,12 @@
+#' grid takes a vector of numbers, divides it into n bins, and rounds each point to the nearest center of a bin.
+grid <- function(x, binwidth = NULL, nbins = 10, range. = range(x)) {
+  if (diff(range.) == 0) return(x)
+  if (is.null(binwidth)) {
+    binwidth <- diff(range.) / npoints
+  }
+  x <- x - range.[1] + binwidth / 2
+  a <- x %/% binwidth + c(0,1)[round(x %% binwidth / binwidth) + 1]
+  a[a == 0] <- 1 # include leftmost point
+  x <- range.[1] + binwidth * a
+  x - binwidth/2
+}

### 2012-06-03 R/geom_star.r
Rewrote geom_star to behave like geom_boxplot, with no automatic glyphing. Improved way plyr_aes handles groups

@@ -1,18 +1,8 @@
-geom_star <- function(mapping = NULL, glyph.by = NULL, data = NULL, 
-  stat = "identity", position = "identity", width = rel(0.95), 
-  height = rel(0.95), x_scale = identity, y_scale = identity, 
-  merge.overlaps = FALSE, reference = NULL, ...) { 
-  
-  major <- mapping[c("x", "y")]
-  mapping$x <- NULL
-  mapping$y <- NULL
-  
-  glyph(
+geom_star <- function(mapping = NULL, data = NULL, stat = "identity", 
+  position = "identity", ...) { 
+
     GeomStar$new(mapping = mapping, data = data, stat = stat, 
-      position = position, ...), 
-    major.aes = major, glyph.by = glyph.by, 
-      width = width, height = height, x_scale = x_scale, y_scale = y_scale, 
-      merge.overlaps = merge.overlaps, reference = reference)
+      position = position, ...)
 }
 
 
@@ -20,24 +10,32 @@ GeomStar <- proto(ggplot2:::Geom, {
   objname <- "star"
   
   # turn cartesian coordinates polar
-  reparameterise <- function(., df, params) {
+  reparameterise <- function(., df, params) {    
     # scale x to be between 0 and 2*pi
-    df$theta <- unlist(rescale_2pi(df["x"]))
-    df$r <- df$y
-    df$x <- df$r * cos(df$theta)
-    df$y <- df$r * sin(df$theta)	
+    df$theta <- unlist(rescale_2pi(df["angle"]))
+    df$r <- unlist(rescale_01(df["r"]))
     
     include_origin <- function(data) {
       data <- data[order(data$theta, data$r), ]
       if (data$theta[1] > 0.01) {
         first <- data[1, ]
-        first$x <- 0
-        first$y <- 0
+        first$theta <- 0
+        first$r <- 0
         data <- rbind(first, data)
+      }
+      if (data$theta[length(data$theta)] < 6.27) {
+        last <- data[length(data$theta), ]
+        last$theta <- 6.28
+        last$r <- 0
+        data <- rbind(data, last)
       }
       data
     }
-    ddply(df, "PANEL", include_origin)
+    df <- ddply(df, c("group", "PANEL"), include_origin)
+    
+    df$x <- df$r * cos(df$theta) + df$x
+    df$y <- df$r * sin(df$theta) + df$y
+    df
   }
   
[…]

### 2012-06-03 R/glyph.r

@@ -56,12 +56,17 @@ assign_glyphs <- function(., data) {
   width <- embed$width
   height <- embed$height
   if (is.rel(width)) {
-    .$embed$width <- width <- (diff(range(vet(globals$x))) + unclass(width)) / 
-      sqrt(length(unique(globals$x))) * unclass(width)
+    .$embed$width <- width <- resolution(globals$x, zero = FALSE) * 
+      unclass(width)
+      
+      #(diff(range(vet(globals$x))) + unclass(width)) / 
+      #sqrt(length(unique(globals$x))) * unclass(width)
   }
   if (is.rel(height)) {
-    .$embed$height <- height <- (diff(range(vet(globals$y))) + unclass(height)) / 
-      sqrt(length(unique(globals$y))) * unclass(height)
+    .$embed$height <- height <- resolution(globals$y, zero = FALSE) * 
+      unclass(height)
+      #(diff(range(vet(globals$y))) + unclass(height)) / 
+      #sqrt(length(unique(globals$y))) * unclass(height)
   }
 
   if (embed$merge) {

### 2012-06-03 R/ply_aes.r

@@ -14,10 +14,9 @@ ply_aes.glayer <- function(layer, .vars = NULL) {
   
 ply_aes.proto <- function(layer, .vars = NULL) {
   if (!is.null(.vars)) {
-    if (!is.function(.vars)) {
-      .vars <- group_by(.vars)
-    }
     layer$plyr <- list(ply.by = .vars)
+  } else {
+    layer$plyr <- list()
   }
   layer$compute_aesthetics <- plyr_aesthetics
   layer

### 2012-06-03 R/plyr_aes.r

@@ -9,46 +9,22 @@ plyr_aesthetics <- function (., data, plot) {
     aesthetics["group"] <- .$geom_params$group
   }
   scales_add_defaults(plot$scales, data, aesthetics, plot$plot_env)
-  
-  if (is.null(data$GLYPH)) {
-    if ("plyr" %in% ls(.)) {
-      data$GLYPH <- plyr$ply.by(data)
-    }
+
+  if (!is.null(aesthetics$group)) {
+    data$group <- unlist(eval.quoted(aesthetics$group, envir = data, 
+      enclos = plot$plot_env))
+    aesthetics$group <- quote(group)
   }
-  aesply(data, c("GLYPH", "PANEL"), aesthetics)
-}
-
-
-eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, try = FALSE) {
-    if (is.numeric(exprs)) 
-        return(envir[exprs])
-    qenv <- if (is.quoted(exprs)) 
-        attr(exprs, "env")
-    else parent.frame()
-    if (is.null(data)) 
-        data <- qenv
-    if (is.data.frame(data) && is.null(enclos)) 
-        enclos <- qenv
-    if (try) {
-        results <- failwith(NULL, ddply, quiet = TRUE) (data, by, apply_maps, 
-          exprs, qenv)    
-    } else {
-        results <- ddply(data, by, apply_maps, exprs, qenv)    
-    }
-    results
-}
-
-
-apply_maps <- function(data, mapping, enclos = parent.frame()) {
-	map <- null_omit(mapping)
-	vars <- llply(map, eval, envir = data, enclos)
-	n <- nrow(data)
-	lengths <- unlist(lapply(vars, length))
-	wrong <- lengths != 1 & lengths != n
-	if (any(wrong)) {
-    stop(paste(
-      "Aesthetics must either be length one, or the same length as the data", 
-      "Problems:", paste(names(wrong)[wrong], collapse = ", ")), call. = FALSE)
-    }
-  data.frame(vars)
+  if ("GLYPH" %in% names(data)) {
+    aesthetics$GLYPH <- quote(GLYPH)
+  }
+  aesthetics$PANEL <- quote(PANEL)
+  
+  criteria <- c("group", "GLYPH", "PANEL", .$plyr$ply.by)
+  criteria <- criteria[criteria %in% names(data)]  
+  data$ply.by <- id(data[criteria], drop = TRUE)
+  
[…]

### 2012-06-03 R/rescalers.r

@@ -7,6 +7,9 @@ rescale_01 <- free <- function(xvars, xlim=NULL) {
 		if (is.factor(x)) {
 			x <- as.numeric(x)
 		}
+    if (inherits(x, "POSIXt")) {
+      x <- as.numeric(x)
+    }
 		x
 	}
 	xvars <- lapply(xvars, numberfy)

### 2012-06-03 demo.r

@@ -34,8 +34,9 @@ ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)),
 # boxes
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
+  width = 10, height = 10,
   major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
-  mean(Catholic)), alpha = 0.1))
+  mean(Catholic))))
 
 # hlines
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
@@ -116,18 +117,19 @@ ggplot(nasa) + glyph(geom_point(aes(x = surftemp, y = temperature),
 
 # trying out geom_star
 # without glyphing
-ggplot(mpg) + GeomStar$new(mapping = aes(r = hwy, angle = cty, group = cyl)) + 
-  facet_wrap(~cyl)
+ggplot(mpg) + GeomStar$new(mapping = aes(r = hwy, angle = cty, x = 0, y = 0, 
+  group = cyl)) + facet_wrap(~cyl)
 
-ggplot(test.data) + geom_star(mapping = aes(x = long[1], y = lat[1], 
-  r = Catholic, angle = Fertility, fill = mean(Education)), 
-  glyph.by = c("long", "lat"))
+ggplot(test.data) + geom_star(mapping = aes(x = 0, y = 0, 
+  r = Catholic, angle = Fertility, group = lat)) + facet_wrap(~lat)
+
+ggplot(nasa) + ply_aes(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, 
+  fill = mean(temperature))), c("lat")) + facet_wrap(~ lat)
 
 ggplot(mpg) + ply_aes(geom_star(mapping = aes(x = cyl[1], y = 1, 
-  r = hwy, angle = cty, fill = mean(hwy)), glyph.by = c("cyl")))
+  r = hwy, angle = cty, fill = mean(hwy), group = cyl)))
 
-ggplot(nasa) + geom_star(aes(r = ozone, angle = date, x = long[1], y = lat[1]), 
-  glyph.by = c("long", "lat"))
+ggplot(nasa) + geom_star(aes(r = ozone, angle = date, x = 0, y = 0))
 
 # new gridding
 ggplot(test.data) + geom_point(aes(Fertility, Education))
@@ -141,6 +143,8 @@ ggplot(cheap.diamonds) +
[…]

### 2012-06-03 R/grid.r
fixed bug in grid/ply_aes interaction

@@ -6,9 +6,9 @@ grid <- function(layer, grid.aes = aes(), x.nbin = 10, y.nbin = 10,
   if (is.glayer(layer) || is.list(layer)) {
     stop("Cannot grid glayer", call. = TRUE)
   }
-  if ("plyr" %in% ls(layer)) {
-    stop("grid and ply_aes used in wrong order.", call. = FALSE)
-  }
+ # if ("plyr" %in% ls(layer)) {
+  #  stop("grid and ply_aes used in wrong order.", call. = FALSE)
+  #}
   
   # aes
   if(is.null(grid.aes$x)) {
@@ -29,7 +29,7 @@ grid <- function(layer, grid.aes = aes(), x.nbin = 10, y.nbin = 10,
   layer$embed <- list(x_scale = x_scale, y_scale = y_scale,
     major.aes = grid.aes[c("x", "y")], grid_env = parent.frame(),
     x.nbin = x.nbin, y.nbin = y.nbin, w.adjust = width.adjust, 
-    h.adjust = height.adjust)
+    h.adjust = height.adjust, ref = .ref)
   layer$assign_glyphs <- assign_grid
   layer$combine_glyphs <- combine_glyphs
   if (.ref) layer$combine_glyphs <- combine_refs
@@ -51,14 +51,15 @@ assign_grid <- function(., data) {
   x.range <- range(x)
   x.breaks <- seq(x.range[1], x.range[2], length = embed$x.nbin + 1)
   width <- diff(x.range) / embed$x.nbin * embed$w.adjust
-  x.val <- (x.breaks + width/2)[-(embed$x.nbin + 1)]
+  # positions in bottom left corner, not on ref box line
+  x.val <- (x.breaks + (1 + 0.05 * !embed$ref) * width/2)[-(embed$x.nbin + 1)]
   x.glyph <- cut(x, x.breaks, labels = x.val, include.lowest = TRUE)
   
   y <- eval(embed$major.aes$y, envir = data, enclos = embed$grid_env)
   y.range <- range(y)
   y.breaks <- seq(y.range[1], y.range[2], length = embed$y.nbin + 1)
   height <- diff(y.range) / embed$y.nbin * embed$h.adjust
-  y.val <- (y.breaks + height/2)[-(embed$y.nbin + 1)]
+  y.val <- (y.breaks + (1 + 0.05 * !embed$ref) * height/2)[-(embed$y.nbin + 1)]
[…]

### 2012-06-03 R/geom_coxcomb.r
altered geom_coxcomb to not consume the group aes

@@ -1,23 +1,8 @@
-geom_coxcomb <- function(mapping = NULL, glyph.by = NULL, data = NULL, 
-  stat = "bin", position = "identity", width = rel(0.95), height = rel(0.95), 
-  x_scale = identity, y_scale = identity, merge.overlaps = FALSE, 
-  reference = NULL, ...) { 
+geom_coxcomb <- function(mapping = NULL, data = NULL, stat = "bin", 
+  position = "identity", ...) { 
   
-  major <- mapping[c("x", "y")]
-  mapping$x <- NULL
-  mapping$y <- NULL
-  
-  glyph(
     GeomCoxcomb$new(mapping = mapping, data = data, stat = stat, 
-      position = position, ...), 
-    major.aes = major, glyph.by = glyph.by, width = width, height = height, 
-      x_scale = x_scale, y_scale = y_scale, merge.overlaps = merge.overlaps, 
-      reference = reference)
-}
-
-
-geom_coxcomb <- function (mapping = NULL, data = NULL, stat = "bin", position = "stack", ...) {
-  GeomBar$new(mapping = mapping, data = data, stat = stat, position = position, ...)
+      position = position, ...)
 }
 
 GeomCoxcomb <- proto(ggplot2:::Geom, {
@@ -68,7 +53,7 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
       df <- cbind(df, x, y)
       rbind(df, df[1, ])
     }
-    ddply(df, c("section", "PANEL"), poly_curve, params$npoints)
+    ddply(df, c("section", "group", "PANEL"), poly_curve, params$npoints)
   }
   
   draw <- draw_groups <- function(., data, scales, coordinates, ...) {
@@ -104,6 +89,7 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
       stop("Missing required aesthetic: r", call. = FALSE)
     }
     names(mapping)[names(mapping) == "r"] <- "x"
+    mapping$section <- coxcomb_sections(mapping)
     
     lyr <- do.call("layer", list(mapping = mapping, data = data, stat = stat, geom = ., 
       position = position, ...))  
@@ -112,3 +98,11 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
   }
   
 })
+
+coxcomb_sections <- function(mapping) {
+  sections <- mapping[c("alpha", "fill", "colour")]
+  sections <- sections[!unlist(lapply(sections, is.null))]
[…]

### 2012-06-04 R/aesply.r
moved eval.plyr into aesply.r

@@ -29,3 +29,39 @@ remove_I <- function(expr) {
   as.call(lapply(expr, Identity))
 }
 
+eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, 
+  try = FALSE) {
+  if (is.numeric(exprs)) 
+    return(envir[exprs])
+  qenv <- if (is.quoted(exprs)) 
+    attr(exprs, "env")
+  else parent.frame()
+  if (is.null(data)) 
+    data <- qenv
+  if (is.data.frame(data) && is.null(enclos)) 
+    enclos <- qenv
+  if (try) {
+    results <- failwith(NULL, ddply, quiet = TRUE) (data, by, apply_maps, 
+                                                    exprs, qenv)    
+  } else {
+    results <- ddply(data, by, apply_maps, exprs, qenv)    
+  }
+  results
+}
+
+
+apply_maps <- function(data, mapping, enclos = parent.frame()) {
+  map <- null_omit(mapping)
+  vars <- llply(map, eval, envir = data, enclos)
+  n <- nrow(data)
+  lengths <- unlist(lapply(vars, length))
+  wrong <- lengths != 1 & lengths != n
+  if (any(wrong)) {
+    stop(paste(
+      "Aesthetics must either be length one, or the same length as the data", 
+      "Problems:", paste(names(wrong)[wrong], collapse = ", ")), 
+      call. = FALSE)
+  }
+  data.frame(vars)
+}
+

### 2012-06-04 R/ref_box.r
fixed bug in ref functions

@@ -13,8 +13,8 @@ ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {
   	
     switch(type,
       glyph = glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, 
-        width = width, height = height, merge = merge.overlaps, ref = NULL, 
-        .ref = TRUE), 
+        width = width, height = height, merge.overlaps = merge.overlaps, 
+        ref = NULL, .ref = TRUE), 
       grid = grid(rlayer, grid.aes = major.aes, x.nbin = x.nbin, 
         y.nbin = y.nbin, width.adjust = 1, height.adjust = 1, .ref = TRUE)
     )
@@ -24,7 +24,7 @@ ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {
 
 
 ref_hline <- function(mapping = NULL, width. = 0.2, fill = "white", ...) {	
-  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
+  function(layer, type, major.aes, glyph.by = NULL, width = rel(1), 
   height = rel(1), merge.overlaps = FALSE) {
 
   	def_aes <- list(xmin = -1, xmax = 1, ymin = -width./2, ymax = width./2)
@@ -37,7 +37,7 @@ ref_hline <- function(mapping = NULL, width. = 0.2, fill = "white", ...) {
   	
   	switch(type,
   	  glyph = glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, 
-  	    width = width, height = height, merge = merge.overlaps, ref = NULL, 
+  	    width = width, height = height, merge.overlaps = merge.overlaps, ref = NULL, 
   	    .ref = TRUE), 
   	  grid = grid(rlayer, grid.aes = major.aes, x.nbin = x.nbin, 
   	    y.nbin = y.nbin, width.adjust = 1, height.adjust = 1, .ref = TRUE)
@@ -46,7 +46,7 @@ ref_hline <- function(mapping = NULL, width. = 0.2, fill = "white", ...) {
 } 
 
 ref_vline <- function(mapping = NULL, width. = 0.2, fill = "white", ...) {	
-  function(layer, major.aes, glyph.by = NULL, width = rel(1), 
+  function(layer, type, major.aes, glyph.by = NULL, width = rel(1), 
   height = rel(1), merge.overlaps = FALSE) {
   	
[…]

### 2012-06-04 R/glyph.r
added ply.aes argument to glyph and grid. Made the functions ply  aesthetics by default

@@ -5,7 +5,7 @@
 #' 4. split up the data
 glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95), 
   height = rel(0.95), x_scale = identity, y_scale = identity, 
-  merge.overlaps = FALSE, reference = NULL, .ref = FALSE) {
+  merge.overlaps = FALSE, reference = NULL, ply.aes = TRUE, .ref = FALSE) {
   
   missing <- c(is.null(major.aes$x), is.null(major.aes$y))
   if (any(missing)) {
@@ -31,11 +31,19 @@ glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95),
   #layer$compute_aesthetics <- plyr_aesthetics
     
   if (is.null(reference)) {
-  	glayer(layer)
+    if (ply.aes) {
+  	  ply_aes(glayer(layer))
+    } else {
+      glayer(layer)
+    }
   } else {
   	ref.layer <- reference(layer, "glyph", major.aes, glyph.by, width, height, 
   	  merge.overlaps)
-  	list(ref.layer, glayer(layer))
+    if (ply.aes) {
+  	  list(ref.layer, ply_aes(glayer(layer)))
+    } else {
+      list(ref.layer, glayer(layer))
+    }
   }
 }
 
@@ -56,17 +64,14 @@ assign_glyphs <- function(., data) {
   width <- embed$width
   height <- embed$height
   if (is.rel(width)) {
-    .$embed$width <- width <- resolution(globals$x, zero = FALSE) * 
-      unclass(width)
-      
-      #(diff(range(vet(globals$x))) + unclass(width)) / 
-      #sqrt(length(unique(globals$x))) * unclass(width)
+    .$embed$width <- width <- max(resolution(globals$x, zero = FALSE) * 
+      unclass(width), (diff(range(vet(globals$x))) + unclass(width)) / 
+      length(unique(globals$x)) * unclass(width))
   }
   if (is.rel(height)) {
-    .$embed$height <- height <- resolution(globals$y, zero = FALSE) * 
-      unclass(height)
-      #(diff(range(vet(globals$y))) + unclass(height)) / 
-      #sqrt(length(unique(globals$y))) * unclass(height)
+    .$embed$height <- height <- max(resolution(globals$y, zero = FALSE) * 
+      unclass(height), (diff(range(vet(globals$y))) + unclass(height)) / 
+      length(unique(globals$y)) * unclass(height))
   }
 
   if (embed$merge) {

### 2012-06-04 R/grid.r

@@ -1,7 +1,7 @@
 #' grid turns an ordinary layer into a gridded version of itself. 
 grid <- function(layer, grid.aes = aes(), x.nbin = 10, y.nbin = 10, 
   x_scale = identity, y_scale = identity, width.adjust = 0.95, 
-  height.adjust = 0.95, reference = NULL, .ref = FALSE) {
+  height.adjust = 0.95, reference = NULL, ply.aes = TRUE, .ref = FALSE) {
   
   if (is.glayer(layer) || is.list(layer)) {
     stop("Cannot grid glayer", call. = TRUE)
@@ -36,13 +36,22 @@ grid <- function(layer, grid.aes = aes(), x.nbin = 10, y.nbin = 10,
   #layer$compute_aesthetics <- plyr_aesthetics
   
   if (is.null(reference)) {
-    glayer(layer)
+    if (ply.aes) {
+      ply_aes(glayer(layer))
+    } else {
+      glayer(layer)
+    }
   } else {
     ref.layer <- reference(layer, "grid", grid.aes, x.nbin = x.nbin, 
       y.nbin = y.nbin)
-    list(ref.layer, glayer(layer))
+    if (ply.aes) {
+      list(ref.layer, ply_aes(glayer(layer)))
+    } else {
+      list(ref.layer, glayer(layer))
+    }
   }
 }
+
 
 
 assign_grid <- function(., data) {

### 2012-06-04 R/ply_aes.r

@@ -8,6 +8,11 @@ ply_aes.list <- function(layer, .vars = NULL) {
 }
 
 ply_aes.glayer <- function(layer, .vars = NULL) {
+  if (!is.null(.vars)) {
+    layer$plyr <- list(ply.by = .vars)
+  } else {
+    layer$plyr <- list()
+  }
   layer$compute_aesthetics <- plyr_aesthetics
   layer
 }

### 2012-06-04 demo.r

@@ -106,7 +106,10 @@ ggplot() + ply_aes(geom_point(aes(x = mean(temperature), y = mean(ozone),
 ggplot() + geom_point(aes(x = mean(temperature), y = mean(ozone), 
   color = lat[1]), data = nasa)  
 
-# geom_scatterplots
+# geom_scatterplots - are redundant
+ggplot(nasa) + glyph(geom_point(aes(surftemp, temperature), size = 1/5), 
+  aes(long[1], lat[1]), c("lat", "long"), reference = ref_box())
+
 ggplot(nasa) + geom_scatterplots(mapping = aes(x = long[1], y = lat[1], 
   minor.x = surftemp, minor.y = temperature), glyph.by = c("long", "lat"), 
   size = 1/5, reference = ref_box())
@@ -126,6 +129,10 @@ ggplot(test.data) + geom_star(mapping = aes(x = 0, y = 0,
 ggplot(nasa) + ply_aes(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, 
   fill = mean(temperature))), c("lat")) + facet_wrap(~ lat)
 
+ggplot(nasa) + map_nasa +
+  ply_aes(glyph(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, 
+  fill = mean(temperature))), aes(long[1], lat[1]), c("long", "lat")))
+
 ggplot(mpg) + ply_aes(geom_star(mapping = aes(x = cyl[1], y = 1, 
   r = hwy, angle = cty, fill = mean(hwy), group = cyl)))
 
@@ -135,20 +142,15 @@ ggplot(nasa) + geom_star(aes(r = ozone, angle = date, x = 0, y = 0))
 ggplot(test.data) + geom_point(aes(Fertility, Education))
 ggplot(test.data) + 
   grid(geom_point(aes(Fertility, Education)), x.nbin = 10, y.nbin = 10, 
-       ref= ref_box(aes(color = mean(Catholic))))
+       ref= ref_box(aes(fill = mean(Catholic))))
 
 cheap.diamonds <- subset(diamonds, price <= 5000 & price >= 600)
 ggplot(cheap.diamonds) +
   grid(geom_bar(aes(x = color, fill = color), position = "dodge"),
     grid.aes = aes(x = carat, y = price), x.nbin = 10, y.nbin = 14,
     y_scale = free, height.adjust = 0.5, width.adjust = 0.5,
-    ref = ref_box())
+    ref = ref_box(aes(color = mean(as.numeric(color)))))
 
-
[…]

### 2012-06-04 R/glyph.r
changed width and height parsing to handle factors

@@ -64,12 +64,12 @@ assign_glyphs <- function(., data) {
   width <- embed$width
   height <- embed$height
   if (is.rel(width)) {
-    .$embed$width <- width <- max(resolution(globals$x, zero = FALSE) * 
+    .$embed$width <- width <- max(resolution(vet(globals$x), zero = FALSE) * 
       unclass(width), (diff(range(vet(globals$x))) + unclass(width)) / 
       length(unique(globals$x)) * unclass(width))
   }
   if (is.rel(height)) {
-    .$embed$height <- height <- max(resolution(globals$y, zero = FALSE) * 
+    .$embed$height <- height <- max(resolution(vet(globals$y), zero = FALSE) * 
       unclass(height), (diff(range(vet(globals$y))) + unclass(height)) / 
       length(unique(globals$y)) * unclass(height))
   }

### 2012-06-04 R/aesply.r
changed aesply to return summarised dataframes when appropriate

@@ -54,6 +54,7 @@ apply_maps <- function(data, mapping, enclos = parent.frame()) {
   map <- null_omit(mapping)
   vars <- llply(map, eval, envir = data, enclos)
   n <- nrow(data)
+  vars <- lapply(vars, condense)
   lengths <- unlist(lapply(vars, length))
   wrong <- lengths != 1 & lengths != n
   if (any(wrong)) {
@@ -65,3 +66,9 @@ apply_maps <- function(data, mapping, enclos = parent.frame()) {
   data.frame(vars)
 }
 
+condense <- function(var) {
+  if (length(unique(var)) == 1) {
+    return(unique(var))
+  } 
+  var
+}

### 2012-06-04 paper/demo.Rnw
tweaked geom_coxcomb. Wrote paper

--- 2012-06-01 paper/demo.Rnw
@@
 This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change).
-\texttt{glyphmaps} was written to create plots like the ones below.
+\texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}.
 Each plot is made with \texttt{glyphmaps}.
-They recreate Figures 8 and 10a from \citet{wickham2012glyph}.
 To run the examples in this document please download the \texttt{glyphmaps} package with the code below.
+(Note: Install won't work until I finish documenting the namespace.
+But cloning and load_all() DOES work.)
 
@@
 <<glyphmaps-fig-8>>=
-p <- ggplot(seasons) + 
-  map_us +
-  glyph(
-    geom_line(aes(x = time, y = pred)), 
-    major = aes(lon[1], lat[1]), glyph.by = "stn", 
-    ref = ref_box(aes(fill = avg)))
+ggplot(seasons) + map_us + 
+  grid(geom_line(aes(x = time, y = pred)), 
+    grid.aes = aes(lon, lat), x.nbin = 58, y.nbin = 25,
+    ref = ref_box(aes(fill = avg), alpha = 0.2, color = NA))
+ggplot(nasa) +  map_nasa +
+  glyph(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, fill = mean(temperature))), 
+  major.aes = aes(long[1], lat[1]), glyph.by = c("long", "lat"))
 @
 
-<<glyph maps merged>>=
+<<glyph-maps-merged>>=
 @
@@
 
-<<boring-random>>=
-library(devtools)
-load_all("ggplyr")
-@
-
 I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \Sexpr{install_github()}.
@@
 \texttt{ggplyr} provides a way for users to compute aesthetics by
-
-<<boring-random>>=
-set.seed(1121)
-(x=rnorm(20))
-mean(x);var(x)
-@
 

### 2012-06-04 paper/paper.Rnw
tweaked geom_coxcomb. Wrote paper

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass{article}
\usepackage[sc]{mathpazo}
\usepackage[T1]{fontenc}
\usepackage{geometry}
\usepackage[round]{natbib}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\setcounter{secnumdepth}{2}
\setcounter{tocdepth}{2}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=2,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{
 pdfstartview={XYZ null null 1}}
\usepackage{breakurl}
\begin{document}

<<setup,echo=FALSE,results=hide,message=FALSE>>=
# this is equivalent to \SweaveOpts{...}
opts_chunk$set(fig.path='figure/minimal-', fig.align='center', fig.show='hold', fig.lp='fig:')
opts_knit$set(cache = TRUE)
opts_knit$set(tidy = FALSE)
opts_knit$set(replace.assign=TRUE, width=90)
setwd("/Users/GarrettsiMac/Documents/git/ggplyr")
library(devtools)
load_all("../ggplyr")
@

\title{A Minimal Demo of \texttt{glyphmaps}}

\author{Garrett Grolemund}

\maketitle
This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change). \texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}. Both plots were made with \texttt{glyphmaps}. To run the examples in this document please download the \texttt{glyphmaps} package with the code below.

<<install, eval = FALSE>>=
library(devtools)
install_github("ggplyr", "garrettgman", "glyphmaps")
library(glyphmaps)
@

<<glyphmaps-fig-11, fig.width = 6, fig.height = 3, out.width = 1.05\textwidth, cache = TRUE, tidy = FALSE>>=
ggplot(seasons) + map_us + 
  grid(geom_line(aes(x = time, y = pred)), 
    grid.aes = aes(lon, lat), x.nbin = 58, y.nbin = 25,
    ref = ref_box(aes(fill = avg), alpha = 0.2, color = NA))
@

<<split-fig, fig.width = 7, fig.height = 5, out.width = 0.8\textwidth, cache = TRUE, tidy = FALSE>>=
ggplot(nasa) +  map_nasa +
  glyph(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, fill = mean(temperature))), 
  major.aes = aes(long[1], lat[1]), glyph.by = c("long", "lat"))
@

\texttt{glyphmaps} helps users create embedded graphics: plots that have subplots built into them. The package is built around three main functions: \texttt{grid}, \texttt{glyph}, and \texttt{ply$\_$aes}. This demo will demonstrate these three functions, their options, and the various uses of embedded graphics.

\section{Glyphs}
Glyphs are geometric objects (i.e, geoms) that display information within the geom. In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system. In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems. The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.

<<star-glyphs, fig.width=3,fig.height=3, out.width = 0.49\textwidth, echo = FALSE, cache = TRUE, fig.cap = 'Individual star glyphs comprise a complete polar plot with an internal coordinate system (r = temperature and angle = date). Multiple star glyphs can be organized with an external coordinate sytem to reveal second order effects.'>>=
one <- subset(nasa, id == "3-1")
# single glyph
ggplot(one) + geom_line(aes(x = date, y = temperature)) + coord_polar()
# embedded
ggplot(nasa) + map_nasa + glyph(geom_star(aes(r = temperature, angle = date, x = 0, y = 0)), major.aes = aes(long[1], lat[1]), glyph.by = c("long", "lat"))
@

Glyphs provide one connection between the grammar and graphics and embedded plots. Every glyph is an embedded plot and each embedded plot is a type of glyph. Hence, we can treat embedded plots as a type of geom. Embedded plots can also be thought of as facetting on a continuous scale. However, since embedded plots are often used with other non-facetted layers, it is easier to implement embedded plots as a type of geom.

\subsection{glyph}
\texttt{glyph} provides a general method for creating glyph geoms in \texttt{ggplot2}. \texttt{glyph} works in conjunction with \texttt{ggplot2} layer objects in the following way. Each layer specifies a type of graph through a combination of geom, stat, aesthetics, parameters, and a position adjustment. \texttt{glyph} uses this specification to build the individual glyphs (i.e, \texttt{glyph} builds glyphs that match the plot specified by the layer). \texttt{glyph} will plot a separate glyph for different subsets of the layer's data. Users specify how to split the data into subsets with the \texttt{glyph.by} argument. This argument behaves similarly to the \texttt{.variables} argument in the \texttt{plyr} package \citep{wickham2011split}. Finally, \texttt{glyph} plots each glyph on the x and y axis specified by the \texttt{major.aes} argument. These ``major axiis'' are mapped to the data just as other \texttt{ggplot2} aesthetics are mapped to the data. Hence they should be specified with the \texttt{aes} function.

To summarize, a layer can be turned into a set of glyphs with the glyph function and three arguments:
<<glyph-example, fig.width = 4, fig.height = 4, out.width = 0.49\textwidth, cache = TRUE, fig.caption = 'Any layer can be made into a collection of glyphs with \texttt{glyph}. The major.aes argument separates glyph from regular grouping (which it may resemble); each glyph can be plotted on new axes independent of the axes used to build its contents.'>>= 
points <- geom_point(aes(x = surftemp, y = temperature), size = 1/2)
## A normal plot
ggplot(nasa) + points
## Glyphed plot
ggplot(nasa) + glyph(layer = points, major = aes(x = long[1], y = lat[1]), glyph.by = c("lat", "long"))
@

\texttt{glyph} gives users incredible freedom to design their own types of glyphs. Any graph that can be specified with a single layer in \texttt{ggplot2} can be turned into a glyph. \texttt{glyphmaps} also provides two new geoms that can be used to create frequently seen glyphs. These are \texttt{geom$\_$star} and \texttt{geom$\_$coxcomb}, see Figure~\ref{fig:star}. Although coxcomb glyphs don't seem much better than Venn pie-agrams (which I suppose we could also do with glyphmaps).
<<star, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, fig.caption = 'Two new geoms to be used with \texttt{glyphmaps}.'>>=
ggplot(nasa) + geom_star(aes(r = ozone, angle = date, x = 0, y = 0, fill = mean(ozone)), alpha = 0.5)
ggplot(seasons) + geom_coxcomb(aes(angle = state, fill = lat))
@

\texttt{glyph} also has a variety of optional arguments that allow users to customize glyph plots. These allow the user to control the dimensions of each glyph, to control the scale within each glyph, to add reference objects to each glyph, and to merge glyphs that overlap into a single glyph.

\subsubsection{width, height}
The \texttt{width} and \texttt{height} arguments of \texttt{glyph} control the width and height of each glyph in the plot.
<<dims, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'Control glyph dimensions with width and height.'>>=
ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
  aes(x = mean(displ), y = mean(cty)), c("year"))
ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
  aes(x = mean(displ), y = mean(cty)), c("year"), 
  width = 1/4, height = 1/4)
@

\subsubsection{x$\_$scale, y$\_$scale}
The \texttt{x$\_$scale} and \texttt{y$\_$scale} arguments of \texttt{glyph} control the scaling within each glyph. As with facets, glyphs share the same scaling by default. The x and y scales can also be set to `free'.
<<scales, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'Control glyph scales with x$\_$scale and y$\_$scale.'>>=
## glyphs share y scale
ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
  aes(x = mean(displ), y = mean(cty)), c("year"), 
  width = 1/4, height = 1/4)
## independent y scales
ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
  aes(x = mean(displ), y = mean(cty)), c("year"),
  width = 1/4, height = 1/4, y_scale = free)
@

\subsubsection{reference objects}
Users can set the \texttt{reference} argument to include reference objects with each glyph. These objects provide a frame of reference which facilitates comparing between glyphs. The \texttt{reference} argument should be set to one of \texttt{ref$\_$box}, \texttt{ref$\_$hline}, \texttt{ref$\_$vline}, \texttt{ref$\_$points}, or \texttt{NULL}. Each of these functions creates a different type of reference object, see Figure~\ref{fig:refs}. Each also accepts a mapping and parameters to customize the appearance of the reference objects.
<<refs, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'Four types of reference objects that can be used with glyphs.'>>=
# boxes
ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
  color = rank(Catholic))), glyph.by = c("lat", "long"), 
  major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
  mean(Catholic))))

# hlines
ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
  color = rank(Catholic))), glyph.by = c("lat", "long"), 
  major = aes(mean(Fertility), mean(Education)), ref = ref_hline(aes(fill = 
  mean(Catholic))))

# vlines
ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
  color = rank(Catholic))), glyph.by = c("lat", "long"), 
  major = aes(mean(Fertility), mean(Education)), ref = ref_vline(aes(fill = 
  mean(Catholic))))

# points
ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
  aes(x = mean(displ), y = mean(cty)), c("year"), y_scale = free, 
  width = 1/3, height = 1/3, reference = ref_points(aes(fill = mean(hwy))))
@

\subsection{merge overlaps}
Often two or more glyphs will overlap each other when plotted. It is possible to combine such overlapping glyphs into a single glyph by setting the \texttt{merge.overlaps} argument to \texttt{TRUE}. \texttt{glyphmaps} will then screen the glyph output for sets of overlapping graphs. Each set will be combined into one glyph and plotted at the at the average location of the combined glyphs, see Figure~\ref{fig:overlaps}.
Caution should be used when setting \texttt{merge.overlaps} to \texttt{TRUE}. Merge overlaps currently works \emph{exactly} as described, which means that for very dense graphs, \texttt{merge.overlaps} is likely to thin out the plot a bit more than the user expects.
<<overlaps, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'Merging overlaps can be a barrel of laughs ...to program.'>>=
## no merging 
ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
  major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
  mean(Catholic))), width = rel(1), height = rel(1))
## merging
ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
  color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
  major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
  mean(Catholic))), merge = TRUE, width = rel(1), height = rel(1))

## compare with dense situations
ggplot(seasons) + map_us + 
  glyph(geom_line(aes(x = time, y = pred)), 
    aes(lon[1], lat[1]), c("stn"), height = 1, width = 1.5,
       ref = ref_box(aes(fill = avg), color = NA))
## vs
ggplot(seasons) + map_us + 
  glyph(geom_line(aes(x = time, y = pred)), 
    aes(lon[1], lat[1]), c("stn"), height = 1, width = 1.5,
       ref = ref_box(aes(fill = avg), color = NA), merge = TRUE)
@

\subsection{grid}
Gridding is the simplest way to embed subplots in a graph. \texttt{grid} basically bins a layer into a 2d grid of subplots and plots a glyph at every bin location. The syntax of \texttt{grid} is very similar to that of \texttt{glyph}. However, \texttt{grid} does not take \texttt{width}, \texttt{height}, or \texttt{merge.overlaps} arguments. the dimension of the glyphs (and the lack of overlaps) is implicit in the intention to grid a layer. Instead, \texttt{grid} takes two new arguments" \texttt{x.nbin} and \texttt{y.nbin}. These specify the number of bins to use on the x and y axiis. They are set to ten by default. 
Like \texttt{glyph}, \texttt{grid} can take reference and scale arguments.

<<grids, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'An example of gridding. The graph on the front page of this article was also constructed with \texttt{grid}.'>>=
## without gridding
ggplot(test.data) + geom_point(aes(Fertility, Education))
## with gridding
ggplot(test.data) + 
  grid(geom_point(aes(Fertility, Education)), x.nbin = 10, y.nbin = 10, 
       ref= ref_box(aes(fill = mean(Catholic))))
@

Gridding provides a new way to approach the problem of overplotting. A grid of glyphs can work like a bin2d but provides more information. Figure~\ref{fig:overplotting} demonstrates this approach by recreating an overplotting graph that has garnered recent interest on the ggplot2 mailing list.

<<overplotting, fig.width = 6, fig.height = 5, out.width = \textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'A recreation of a plot from Hadley's webinar. The glyphmaps differs from the original in that the stroke of the reference box is not mapped to the count within the box. Glyphmaps currently only calculates stats at the subplot levels.'>>=
## without gridding
cheap.diamonds <- subset(diamonds, price <= 5000 & price >= 600)
ggplot(cheap.diamonds) +
  grid(geom_bar(aes(x = color, fill = color), position = "dodge"),
    grid.aes = aes(x = carat, y = price), x.nbin = 10, y.nbin = 14,
    y_scale = free, height.adjust = 0.5, width.adjust = 0.5,
    ref = ref_box(aes(color = mean(as.numeric(color)))))
@

\subsection{ply$\_$aes}

\texttt{ply$\_$aes} wraps a layer like \texttt{grid} or \texttt{glyph} and causes the aesthetics to be calculated in a groupwise manner. Users can specify the groups to use, or \texttt{ply$\_$aes} will default to any groupings provided by glyph.by or group.by parameters, or a group aesthetic.

<<groupwise1, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = '\texttt{ply$\_$aes} forces aesthetics to be calculated within groups.'>>=
## normal \texttt{ggplot2} behavior
ggplot(mpg) + geom_point(aes(hwy, cty, color = rank(displ))) + 
  facet_wrap(~cyl)
## groupwise aesthetics
ggplot(mpg) + ply_aes(geom_point(aes(hwy, cty, color = rank(displ))), 
  c("cyl")) + facet_wrap(~cyl)
@

\texttt{ply$\_$aes} doesn't actually create embedded plots, but is closely related to the glyphing process. By default, \texttt{ggplot2} does not calculate statistics in a groupwise manner. However, such calculations can be useful in certain situations. For example, \texttt{ply$\_$aes} is built into \texttt{glyph} and \texttt{grid} by default; embedded plots quickly become boring when aesthetics do not vary from subplot to subplot. Groupwise aesthetics can be interesting in other situations as well. For example, Figure~\ref{fig:groupwise2} demonstrates how groupwise aesthetics can be used to avoid overplotting.
<<groupwise2, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = '\texttt{ply$\_$aes} provides another way to deal with overplotting.'>>=
## overplotted
ggplot() + geom_point(aes(x = temperature, y = ozone, color = lat), data = nasa)  
# using ply_aes to plot just summaries
ggplot() + ply_aes(geom_point(aes(x = mean(temperature), y = mean(ozone), 
  color = lat[1]), data = nasa), c("lat", "long"))
@

\section{speed considerations}
Embedding plots is computationally expensive. Such plots require much of both the computational engine of R and the graphics device. Hence, it is not unreasonable to expect \texttt{glyphmaps} to be slow, which it can be. However, relative to similar visual tasks \texttt{glyphmaps} is surprisingly fast. The final code chunk compares the speed of plotting forty thousand points in an embedded plot to plotting the same points in a facetted plot. \texttt{glyphmaps} performs quite well in comparison.

<<<<speed-test, fig.width = 4, fig.height = 3, out.width = 0.49\textwidth, cache = TRUE, tidy = FALSE, fig.caption = 'Plot speeds for an embedded plot vs. a facetted plot. Originally run on my iMac desktop.'>>=
## embedded plot
system.time(print(ggplot(nasa) + glyph(geom_point(aes(x = surftemp, 
  y = temperature), size = 1/2), aes(x = long[1], y = lat[1]), c("lat", "long"))))

## facetted plot
system.time(print(ggplot(nasa) + glyph(geom_point(aes(x = surftemp, 
  y = temperature), size = 1/2) + facet_grid(long~lat))))
@

\bibliographystyle{plainnat}
\bibliography{ggplyr}
\end{document}

### 2012-06-04 paper/paper.tex
tweaked geom_coxcomb. Wrote paper

--- 2012-05-30 paper/demo.tex
@@
 \maketitle
-This document demonstrates the basic features of the \texttt{ggplyr} package (name may change).
-\texttt{ggplyr} was written to create plots like the ones below.
-Each plot was made with \texttt{ggplyr}.
-The first two plots recreate plots from \citet{wickham2011split}.
-The last two recreate plots from \citet{wickham2012glyph}.
-
-\section{Glyphs}
-The common feature of these graphs is that they are built around glyphs.
-Glyphs are geometric objects (i.e, geoms) designed to display information within each geom.
-In other words, a glyph can display information even if it ids drawn by itself, without references to an external coordinate system.
-In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
-The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
-
-Glyphs reveal a heirarchical structure to graphics: every plot is a collection of geoms, each of which can be thought of as its own self contained plot.
-Sometimes these subplots are not very interesting, as in the subplot created by a single point geom.
-Sometimes, they are quite complex, as in the star glyph of Figure~\ref{fig:star-glyphs}.
-
-Graphs inherit this heirarchical structure from the data they describe.
-Data is produced through an iterative process of collecting observations, grouping observations and summarizing groups of observations to create more compact, information dense sets of data.
-Humans innately perform this process when collecting data.
-It is a cognitive pattern of the human brain which I will write about in the second cognitive chapter of my thesis.
-
-Glyph maps simultaneously expose data from multiple levels of this process.
-As a set of geoms, the glyphs reveal relationships between data points in the higher level, compact data set.
-As individual plots, each glyph retains information about the data points in the lower level group of data that it summarises.
-This dual display makes glyph maps particularly useful for certain data analysis tasks.
-It also provides two different approaches to constructing glyph maps.
-
-Glyph maps can be built from the top down by treating each glyph as a geom within the plot of interest.
-\texttt{ggplyr} provides new geoms \texttt{geom_star}, \texttt{geom_radar}, \texttt{geom_dart}, and \texttt{geom_plyr} which allow a user to quickly incorporate geoms into an existing plot.
-
-Alternatively, glyph maps can be built from the bottom up by treating each glyph as an existing plot and then combining these plots together with an external coordinate system.
-This approach parallels the split-apply-combine strategy of data analysis \citep{wickham2011split}.
-\texttt{ggplyr} provides a set of functions such as \texttt{ggply}, \texttt{dgply}, \texttt{grid}, and \texttt{nest} that help users use the familiar split-apply-combine process to create complex glyph maps.
-This approach will illustrate that any graph can be turned into a glyph to be used in a higher order graph.
-
-The remainder of this paper will demonstrate how users can use \texttt{ggplyr} tools to build glyph maps from either the bottom up or the top down.
-
-Although we have chosen to describe such plots as glyph maps, they really represent a general class of graphs that have been difficult to construct until now.
-Because each glyph is equivalent to an independent plot, glyph maps describe any plot that embeds other plots within itself.
-Embedded plots occur in a variety of contexts and I have a stack of photocopies to help demonstrate this.
-Embedded plots also fit into the grammar of graphics, where they represent an extension of facetting.
-When plots are embedded into two categorical axes, a facetted plot results.
-When a graph is facetted on continuous axes (which is rarely done), a glyph map results.
-
-(Interesetingly most embedded plots have settled to or are stuck in the stage of embedding plots within a geographic map.
-This resembles how Wainer, Tufte and others have traced the development of graphs through a concrete stage where cartesian axes seldom refered to anything other than latitude and longitude.
-Eventually graphs became more sophisticated and axes were used to represent  abstract concepts.
-Infovis developed from SciVis in the same way.
-Scivis focused on visualizing physical dimensions and eventually spawned InfoVis, which focuses on visualizing abstract dimensions.
-Perhaps embedded plots are developing through the same process, or perhaps the complexity of visualizing two orders of data at once requires one of them to have a concrete interpretation.)
-
-\section{top down methods}
-\texttt{ggplyr}'s top down methods help users fit glyphs into existing methods of visualization.
-These methods provide new geoms (often based on new grobs) that can be used alongside existing geoms in \texttt{ggplot2}.
-
-Top down methods are difficult to implement in \texttt{ggplot2} for two reasons
-\begin{ennumerate}
-  \item First, \texttt{ggplot2} calculates aesthetics on the entire data set at once. However, glyphs contain aesthetics that must be keyed to subsets of the data.
-  \item Second, the final width and height of individual glyphs often depends on non-position aesthetics, such as angle and length.
-In the \texttt{ggplot2} pipeline, these aesthetics are scaled right before the plot ranges are trained.
-As a result, the final widths of glyphs must be computed at draw time and frequently place parts of the glyph outside the plot window.
-\end{ennumerate}
+This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change).
+\texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}.
+Both plots were made with \texttt{glyphmaps}.
+To run the examples in this document please download the \texttt{glyphmaps} package with the code below.
 
@@
 \ttfamily\noindent
-\hlfunctioncall{set.seed}\hlkeyword{(}\hlnumber{1121}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlkeyword{(}\hlsymbol{x}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{rnorm}\hlkeyword{(}\hlnumber{20}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{install\usebox{\hlnormalsizeboxunderscore}github}\hlkeyword{(}\hlstring{"{}ggplyr"{}}\hlkeyword{,}{\ }\hlstring{"{}garrettgman"{}}\hlkeyword{,}{\ }\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{glyphmaps}\hlkeyword{)}\mbox{}
 \normalfont
 \end{flushleft}
-\begin{verbatim}
-##  [1]  0.14496  0.43832  0.15319  1.08494  1.99954 -0.81188  0.16027  0.58589  0.36009
-## [10] -0.02531  0.15088  0.11008  1.35968 -0.32699 -0.71638  1.80977  0.50840 -0.52746
-## [19]  0.13272 -0.15594
-\end{verbatim}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-## [1] 0.3217
-\end{verbatim}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{var}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-## [1] 0.5715
-\end{verbatim}
 \end{kframe}
 \end{knitrout}
-
-The first element of \texttt{x} is 0.145. Boring boxplots
-and histograms recorded by the PDF device:
 
@@
 \ttfamily\noindent
-\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }two{\ }plots{\ }side{\ }by{\ }side{\ }(option{\ }fig.show=\usebox{\hlnormalsizeboxsinglequote}hold\usebox{\hlnormalsizeboxsinglequote})}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{par}\hlkeyword{(}\hlargument{mar}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{4}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{,}{\ }\hlnumber{0.1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{cex.lab}{\ }\hlargument{=}{\ }\hlnumber{0.95}\hlkeyword{,}{\ }\hlargument{cex.axis}{\ }\hlargument{=}{\ }\hlnumber{0.9}\hlkeyword{,}{\ }\hlargument{mgp}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlnumber{2}\hlkeyword{,}{\ }\hlnumber{0.7}\hlkeyword{,}{\ }\hlnumber{0}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
-\hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{grid.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{58}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{25}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.2}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{)}\mbox{}
 \normalfont
@@
 
-{\centering \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots1} \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots2}
+{\centering \includegraphics[width=1.05\textwidth]{figure/minimal-glyphmaps-fig-11}
 
@@
 
-Do the above chunks work? You should be able to compile the \TeX{}
-document and get a PDF file like this one: \url{https://github.com/downloads/yihui/knitr/knitr-minimal.pdf}.
-The Rnw source of this document is at \url{https://github.com/yihui/knitr/blob/master/inst/examples/knitr-minimal.Rnw}.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}nasa}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}long"{}}\hlkeyword{,}{\ }\hlstring{"{}lat"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.8\textwidth]{figure/minimal-split-fig}
+
+}
+
+\end{knitrout}
+
+\texttt{glyphmaps} helps users create embedded graphics: plots that have subplots built into them.
+The package is built around three main functions: \texttt{grid}, \texttt{glyph}, and \texttt{ply$\_$aes}.
+This demo will demonstrate these three functions, their options, and the various uses of embedded graphics.
+
+\section{Glyphs}
+Glyphs are geometric objects (i.e, geoms) that display information within the geom.
+In other words, a glyph can display information even if it is drawn by itself, without references to an external coordinate system.
+In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
+The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[]
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-star-glyphs1} \includegraphics[width=0.49\textwidth]{figure/minimal-star-glyphs2}
+
+}
+
+\caption['Individual star glyphs comprise a complete polar plot with an internal coordinate system (r = temperature and angle = date)]{'Individual star glyphs comprise a complete polar plot with an internal coordinate system (r = temperature and angle = date).
+Multiple star glyphs can be organized with an external coordinate sytem to reveal second order effects.'\label{fig:star-glyphs}}
+\end{figure}
+
+\end{knitrout}
+
+Glyphs provide one connection between the grammar and graphics and embedded plots.
+Every glyph is an embedded plot and each embedded plot is a type of glyph.
+Hence, we can treat embedded plots as a type of geom.
+Embedded plots can also be thought of as facetting on a continuous scale.
+However, since embedded plots are often used with other non-facetted layers, it is easier to implement embedded plots as a type of geom.
+
+\subsection{glyph}
+\texttt{glyph} provides a general method for creating glyph geoms in \texttt{ggplot2}.
+\texttt{glyph} works in conjunction with \texttt{ggplot2} layer objects in the following way.
+Each layer specifies a type of graph through a combination of geom, stat, aesthetics, parameters, and a position adjustment.
+\texttt{glyph} uses this specification to build the individual glyphs (i.e, \texttt{glyph} builds glyphs that match the plot specified by the layer).
+\texttt{glyph} will plot a separate glyph for different subsets of the layer's data.
+Users specify how to split the data into subsets with the \texttt{glyph.by} argument.
+This argument behaves similarly to the \texttt{.variables} argument in the \texttt{plyr} package \citep{wickham2011split}.
+Finally, \texttt{glyph} plots each glyph on the x and y axis specified by the \texttt{major.aes} argument.
+These ``major axiis'' are mapped to the data just as other \texttt{ggplot2} aesthetics are mapped to the data.
+Hence they should be specified with the \texttt{aes} function.
+
+To summarize, a layer can be turned into a set of glyphs with the glyph function and three arguments:
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlsymbol{points}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }A{\ }normal{\ }plot}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{points}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }Glyphed{\ }plot}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlargument{layer}{\ }\hlargument{=}{\ }\hlsymbol{points}\hlkeyword{,}{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-glyph-example1} \includegraphics[width=0.49\textwidth]{figure/minimal-glyph-example2}
+
+}
+
+\end{knitrout}
+
+\texttt{glyph} gives users incredible freedom to design their own types of glyphs.
+Any graph that can be specified with a single layer in \texttt{ggplot2} can be turned into a glyph.
+\texttt{glyphmaps} also provides two new geoms that can be used to create frequently seen glyphs.
+These are \texttt{geom$\_$star} and \texttt{geom$\_$coxcomb}, see Figure~\ref{fig:star}.
+Although coxcomb glyphs don't seem much better than Venn pie-agrams (which I suppose we could also do with glyphmaps).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{ozone}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.5}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}coxcomb}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{state}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-star1} \includegraphics[width=0.49\textwidth]{figure/minimal-star2}
+
+}
+
+\end{knitrout}
+
+\texttt{glyph} also has a variety of optional arguments that allow users to customize glyph plots.
+These allow the user to control the dimensions of each glyph, to control the scale within each glyph, to add reference objects to each glyph, and to merge glyphs that overlap into a single glyph.
+
+\subsubsection{width, height}
+The \texttt{width} and \texttt{height} arguments of \texttt{glyph} control the width and height of each glyph in the plot.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-dims1} \includegraphics[width=0.49\textwidth]{figure/minimal-dims2}
+
+}
+
+\end{knitrout}
+
+\subsubsection{x$\_$scale, y$\_$scale}
+The \texttt{x$\_$scale} and \texttt{y$\_$scale} arguments of \texttt{glyph} control the scaling within each glyph.
+As with facets, glyphs share the same scaling by default.
+The x and y scales can also be set to `free'.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }glyphs{\ }share{\ }y{\ }scale}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }independent{\ }y{\ }scales}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{4}\hlkeyword{,}{\ }\hlargument{y\usebox{\hlnormalsizeboxunderscore}scale}{\ }\hlargument{=}{\ }\hlsymbol{free}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-scales1} \includegraphics[width=0.49\textwidth]{figure/minimal-scales2}
+
+}
+
+\end{knitrout}
+
+\subsubsection{reference objects}
+Users can set the \texttt{reference} argument to include reference objects with each glyph.
+These objects provide a frame of reference which facilitates comparing between glyphs.
+The \texttt{reference} argument should be set to one of \texttt{ref$\_$box}, \texttt{ref$\_$hline}, \texttt{ref$\_$vline}, \texttt{ref$\_$points}, or \texttt{NULL}.
+Each of these functions creates a different type of reference object, see Figure~\ref{fig:refs}.
+Each also accepts a mapping and parameters to customize the appearance of the reference objects.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }boxes}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }hlines}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}hline}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }vlines}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}vline}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }points}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{trans}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{year}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{cty}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}year"{}}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y\usebox{\hlnormalsizeboxunderscore}scale}{\ }\hlargument{=}{\ }\hlsymbol{free}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{3}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{3}\hlkeyword{,}{\ }\hlargument{reference}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}points}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{hwy}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-refs1} \includegraphics[width=0.49\textwidth]{figure/minimal-refs2} \includegraphics[width=0.49\textwidth]{figure/minimal-refs3} \includegraphics[width=0.49\textwidth]{figure/minimal-refs4}
+
+}
+
+\end{knitrout}
+
+\subsection{merge overlaps}
+Often two or more glyphs will overlap each other when plotted.
+It is possible to combine such overlapping glyphs into a single glyph by setting the \texttt{merge.overlaps} argument to \texttt{TRUE}.
+\texttt{glyphmaps} will then screen the glyph output for sets of overlapping graphs.
+Each set will be combined into one glyph and plotted at the at the average location of the combined glyphs, see Figure~\ref{fig:overlaps}.
+Caution should be used when setting \texttt{merge.overlaps} to \texttt{TRUE}.
+Merge overlaps currently works \emph{exactly} as described, which means that for very dense graphs, \texttt{merge.overlaps} is likely to thin out the plot a bit more than the user expects.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }no{\ }merging{\ }}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{3}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlfunctioncall{rel}\hlkeyword{(}\hlnumber{1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlfunctioncall{rel}\hlkeyword{(}\hlnumber{1}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }merging}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Agriculture}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{3}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{major}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{merge}{\ }\hlargument{=}{\ }\hlnumber{TRUE}\hlkeyword{,}{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlfunctioncall{rel}\hlkeyword{(}\hlnumber{1}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlfunctioncall{rel}\hlkeyword{(}\hlnumber{1}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }compare{\ }with{\ }dense{\ }situations}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}stn"{}}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{,}{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1.5}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }{\ }{\ }{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }vs}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}stn"{}}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{height}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{,}{\ }\hlargument{width}{\ }\hlargument{=}{\ }\hlnumber{1.5}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }{\ }{\ }{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{merge}{\ }\hlargument{=}{\ }\hlnumber{TRUE}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-overlaps1} \includegraphics[width=0.49\textwidth]{figure/minimal-overlaps2} \includegraphics[width=0.49\textwidth]{figure/minimal-overlaps3} \includegraphics[width=0.49\textwidth]{figure/minimal-overlaps4}
+
+}
+
+\end{knitrout}
+
+\subsection{grid}
+Gridding is the simplest way to embed subplots in a graph.
+\texttt{grid} basically bins a layer into a 2d grid of subplots and plots a glyph at every bin location.
+The syntax of \texttt{grid} is very similar to that of \texttt{glyph}.
+However, \texttt{grid} does not take \texttt{width}, \texttt{height}, or \texttt{merge.overlaps} arguments.
+the dimension of the glyphs (and the lack of overlaps) is implicit in the intention to grid a layer.
+Instead, \texttt{grid} takes two new arguments" \texttt{x.nbin} and \texttt{y.nbin}.
+These specify the number of bins to use on the x and y axiis.
+They are set to ten by default.
+Like \texttt{glyph}, \texttt{grid} can take reference and scale arguments.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }without{\ }gridding}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }with{\ }gridding}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{test.data}\hlkeyword{)}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{Fertility}\hlkeyword{,}{\ }\hlsymbol{Education}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{10}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{10}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }{\ }{\ }{\ }\hlargument{ref}\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{Catholic}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-grids1} \includegraphics[width=0.49\textwidth]{figure/minimal-grids2}
+
+}
+
+\end{knitrout}
+
+Gridding provides a new way to approach the problem of overplotting.
+A grid of glyphs can work like a bin2d but provides more information.
+Figure~\ref{fig:overplotting} demonstrates this approach by recreating an overplotting graph that has garnered recent interest on the ggplot2 mailing list.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }without{\ }gridding}\hspace*{\fill}\\
+\hlstd{}\hlsymbol{cheap.diamonds}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{subset}\hlkeyword{(}\hlsymbol{diamonds}\hlkeyword{,}{\ }\hlsymbol{price}{\ }\usebox{\hlnormalsizeboxlessthan}={\ }\hlnumber{5000}{\ }\hlkeyword{\usebox{\hlnormalsizeboxand}}{\ }\hlsymbol{price}{\ }\hlkeyword{\usebox{\hlnormalsizeboxgreaterthan}=}{\ }\hlnumber{600}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{cheap.diamonds}\hlkeyword{)}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}bar}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{color}\hlkeyword{,}{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{color}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{position}{\ }\hlargument{=}{\ }\hlstring{"{}dodge"{}}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{grid.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{carat}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{price}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{10}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{14}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{y\usebox{\hlnormalsizeboxunderscore}scale}{\ }\hlargument{=}{\ }\hlsymbol{free}\hlkeyword{,}{\ }\hlargument{height.adjust}{\ }\hlargument{=}{\ }\hlnumber{0.5}\hlkeyword{,}{\ }\hlargument{width.adjust}{\ }\hlargument{=}{\ }\hlnumber{0.5}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlfunctioncall{as.numeric}\hlkeyword{(}\hlsymbol{color}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=\textwidth]{figure/minimal-overplotting}
+
+}
+
+\end{knitrout}
+
+\subsection{ply$\_$aes}
 

### 2012-06-04 R/geom_coxcomb.r

@@ -84,11 +84,11 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
   new <- function(., mapping = NULL, data = NULL, stat = NULL, 
     position = NULL, npoints = 10, ...) {
     
-    missing <- !(c("r") %in% names(mapping))
+    missing <- !(c("angle") %in% names(mapping))
     if (any(missing)) {
-      stop("Missing required aesthetic: r", call. = FALSE)
+      stop("Missing required aesthetic: angle", call. = FALSE)
     }
-    names(mapping)[names(mapping) == "r"] <- "x"
+    names(mapping)[names(mapping) == "angle"] <- "x"
     mapping$section <- coxcomb_sections(mapping)
     
     lyr <- do.call("layer", list(mapping = mapping, data = data, stat = stat, geom = ., 

### 2012-06-04 R/aesply.r
added documentation to aesply.r

@@ -1,3 +1,18 @@
+#' Evaluate aesthetics by group
+#' 
+#' aesply splits a data frame into pieces, evaluates a list of aesthetics within 
+#' each piece, and combines the results into a new data frame. Each aesthetic in 
+#' the list must return either a single value per piece or a single value per 
+#' row in the piece.
+#' 
+#' @keywords internal
+#' @param data a data frame
+#' @param .vars a vector of variables to split data by. Each element should be 
+#' the name of a variable in data saved as a character string.
+#' @param aesthetics an object of class uneval, usually the output of 
+#' \code{\link[ggplot2]{aes}}
+#' @return a data frame
+#' @export
 aesply <- function(data, .var, aesthetics) {
   no.ply <- unlist(lapply(aesthetics, first_name)) == "I"
   if (!any(no.ply)) {
@@ -20,7 +35,12 @@ aesply <- function(data, .var, aesthetics) {
   compact(eval.plyr(aesthetics, data, .var))
 }
 
-
+#' replace I() with identity
+#' 
+#' remove_I searches through an expression for the \code{\link{I}} function and 
+#' replaces it with \code{\link{identity}}. remove_I is used when an aesthetic 
+#' has been surrounded with I() to prevent groupwise calculation.
+#' @param expr an expression
 remove_I <- function(expr) {
   Identity <- function(x) {
     if (x == "I") x <- quote(identity)
@@ -29,6 +49,21 @@ remove_I <- function(expr) {
   as.call(lapply(expr, Identity))
 }
 
+#' Evaluate a list of expressions by group
+#' 
+#' Evaluates qoted variables by group in a data frame. Based on 
+#' \code{\link[plyr]{eval.quoted}}.
+#' @keywords internal
+#' @param exprs a list of expressions
+#' @param data a data frame
+#' @param by a vector of character strings that specify variable names in data. 
+#' data will be split into groups based on the unique combinations of the values 
+#' of these variables within the data frame. exprs will be evaluated separately 
+#' for each group.
+#' @param enclos an environment
+#' @return a data frame formed by combining the results of evaluating exprs in 
+#' each group of data
+#' @export
 eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, 
   try = FALSE) {
   if (is.numeric(exprs)) 
@@ -49,7 +84,21 @@ eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL,
   results
 }
 
-
[…]

### 2012-06-04 all.r
removed add_glyphs which no longer appears to be used

@@ -1,0 +1,1365 @@
+aesply <- function(data, .var, aesthetics) {
+  no.ply <- unlist(lapply(aesthetics, first_name)) == "I"
+  if (!any(no.ply)) {
+    return(compact(eval.plyr(aesthetics, data, .var)))
+  }
+  
+  all.aes <- lapply(aesthetics[no.ply], remove_I)
+  evaled <- compact(eval.quoted(all.aes, data))
+  lengths <- vapply(evaled, length, integer(1))
+  n <- if (length(lengths) > 0) max(lengths) else 0
+  wrong <- lengths != 1 & lengths != n
+  if (any(wrong)) {
+    stop("Aesthetics must either be length one, or the same length as the data", 
+         "Problems:", paste(aesthetics[wrong], collapse = ", "), 
+         call. = FALSE)
+  }
+  data <- cbind(data, evaled)
+  all.data <- data.frame(evaled) 
+  aesthetics[no.ply] <- lapply(names(aesthetics)[no.ply], as.name)
+  compact(eval.plyr(aesthetics, data, .var))
+}
+
+
+remove_I <- function(expr) {
+  Identity <- function(x) {
+    if (x == "I") x <- quote(identity)
+    x
+  }
+  as.call(lapply(expr, Identity))
+}
+
+eval.plyr <- function (exprs, data = NULL, by = NULL, enclos = NULL, 
+  try = FALSE) {
+  if (is.numeric(exprs)) 
+    return(envir[exprs])
+  qenv <- if (is.quoted(exprs)) 
+    attr(exprs, "env")
+  else parent.frame()
+  if (is.null(data)) 
+    data <- qenv
+  if (is.data.frame(data) && is.null(enclos)) 
+    enclos <- qenv
+  if (try) {
+    results <- failwith(NULL, ddply, quiet = TRUE) (data, by, apply_maps, 
+                                                    exprs, qenv)    
+  } else {
+    results <- ddply(data, by, apply_maps, exprs, qenv)    
+  }
+  results
+}
+
+
+apply_maps <- function(data, mapping, enclos = parent.frame()) {
+  map <- null_omit(mapping)
+  vars <- llply(map, eval, envir = data, enclos)
+  n <- nrow(data)
+  vars <- lapply(vars, condense)
+  lengths <- unlist(lapply(vars, length))
+  wrong <- lengths != 1 & lengths != n
+  if (any(wrong)) {
+    stop(paste(
+      "Aesthetics must either be length one, or the same length as the data", 
[…]

### 2012-06-04 all.r
removed add-glyph from all.r

@@ -1,3 +1,5 @@
+
+
 aesply <- function(data, .var, aesthetics) {
   no.ply <- unlist(lapply(aesthetics, first_name)) == "I"
   if (!any(no.ply)) {

### 2012-06-04 paper/paper.Rnw
editing paper
paper changes

@@
 ## overplotted
-ggplot() + geom_point(aes(x = temperature, y = ozone, color = lat), data = nasa)  
+ggplot() + geom_point(aes(x = temperature, y = ozone, color = lat), data = nasa, size = 1/2)  
 # using ply_aes to plot just summaries
 ggplot() + ply_aes(geom_point(aes(x = mean(temperature), y = mean(ozone), 
-  color = lat[1]), data = nasa), c("lat", "long"))
+  color = lat[1]), data = nasa, size = 1/2), c("lat", "long"))
 @
@@
 ## facetted plot
-system.time(print(ggplot(nasa) + glyph(geom_point(aes(x = surftemp, 
+system.time(print(ggplot(nasa) + geom_point(aes(x = surftemp, 
   y = temperature), size = 1/2) + facet_grid(long~lat))))

### 2012-06-04 R/geom_coxcomb.r
Documented geom_coxcomb. Added na.rm argument

@@ -1,8 +1,34 @@
+#' Coxcomb glyphs
+#' 
+#' geom_coxcomb draws the type of glyph commonly called a coxcomb plot or polar 
+#' area plot, popularized by Florence Nightingale.
+#' 
+#' @param mapping The aesthetic mapping, usually constructed with 
+#' \code{\link[ggplot2]{aes}}. Only needs to be set at the layer level if you 
+#' are overriding the plot defaults.
+#' @param data A layer specific dataset - only needed if you want to override 
+#' the plot defaults
+#' @param stat The statistical transformation to use for this layer.
+#' @param position The position adjustment to use for overlapping points in this 
+#' layer
+#' @param npoints the number of points to use when drawing the arcs with line 
+#' segments. Defaults to 10.
+#' @param na.rm If FALSE (the default), removes missing values with a warning. 
+#' If TRUE, silently removes missing variables.
+#' @param ... other arguments passed on to \code{\link[ggplot2]{layer}}. This 
+#' can include aesthetics whose values you want to set, not map. See 
+#' \code{\link[ggplot2]{layer}} for more details.
+#' 
+#' @section Aesthetics
+#' geom_coxcomb understands the following aesthetics: x, y, colour, fill, size, 
+#' linetype, weight, and alpha.
+#' 
+#' @export
 geom_coxcomb <- function(mapping = NULL, data = NULL, stat = "bin", 
-  position = "identity", ...) { 
+  position = "identity", npoints = 10, na.rm = FALSE, ...) { 
   
     GeomCoxcomb$new(mapping = mapping, data = data, stat = stat, 
-      position = position, ...)
+      position = position, npoints = npoints, na.rm = na.rm, ...)
 }
 
 GeomCoxcomb <- proto(ggplot2:::Geom, {
@@ -82,7 +108,7 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
   }
   
   new <- function(., mapping = NULL, data = NULL, stat = NULL, 
-    position = NULL, npoints = 10, ...) {
+    position = NULL, npoints = 10, na.rm = FALSE, ...) {
     
     missing <- !(c("angle") %in% names(mapping))
     if (any(missing)) {
@@ -90,6 +116,9 @@ GeomCoxcomb <- proto(ggplot2:::Geom, {
     }
     names(mapping)[names(mapping) == "angle"] <- "x"
     mapping$section <- coxcomb_sections(mapping)
+    
+    if (na.rm) {
[…]

### 2012-06-07 paper/paper.Rnw
Changed options on paper. Trying to get cache to work.

--- 2012-06-04 paper/paper.Rnw
@@
 # this is equivalent to \SweaveOpts{...}
-opts_chunk$set(fig.path='figure/minimal-', fig.align='center', fig.show='hold', fig.lp='fig:')
-opts_knit$set(cache = TRUE)
-opts_knit$set(tidy = FALSE)
+opts_chunk$set(fig.path='figure/', fig.align='center', fig.show='hold', fig.lp='fig:', cache.path = 'cache/', cache = TRUE, tidy = FALSE)
 opts_knit$set(replace.assign=TRUE, width=90)
@@
 
-\subsection{ply$\_$aes}
+\subsection{ply\_aes}
 
-\texttt{ply$\_$aes} wraps a layer like \texttt{grid} or \texttt{glyph} and causes the aesthetics to be calculated in a groupwise manner.
+\texttt{ply\_aes} wraps a layer like \texttt{grid} or \texttt{glyph} and causes the aesthetics to be calculated in a groupwise manner.
 Users can specify the groups to use, or \texttt{ply$\_$aes} will default to any groupings provided by glyph.by or group.by parameters, or a group aesthetic.

### 2012-06-21 demo.r
all demos work

@@ -137,7 +137,7 @@ ggplot(test.data) + geom_star(mapping = aes(x = 0, y = 0,
 ggplot(nasa) + ply_aes(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, 
   fill = mean(temperature))), c("lat")) + facet_wrap(~ lat)
 
-ggplot(nasa) + #map_nasa +
+ggplot(nasa) + map_nasa +
   glyph(geom_star(aes(r = ozone, angle = date, x = 0, y = 0, 
   fill = mean(temperature))), aes(long[1], lat[1]), c("long", "lat"))
 
@@ -170,11 +170,7 @@ mpg$lat <- sample(1:4, nrow(mpg), replace = TRUE)
 ggplot(mpg) + GeomCoxcomb$new(mapping = aes(angle = trans, fill = lat, 
   group = lat)) + facet_wrap(~cyl)
 ggplot(mpg) + geom_coxcomb(aes(angle = trans, fill = lat))
-        , group = lat))
-###########################################
-###          not yet working            ###
-###########################################
-load_all("../ggplyr")
+
 
 ggplot(seasons) + 
   glyph(
@@ -183,18 +179,7 @@ ggplot(seasons) +
     height = 1, width = 2,
     ref = ref_box(aes(fill = avg)), merge = TRUE)
 
-
-
 ###########################################
-###             obsolete                ###
+###          not yet working            ###
 ###########################################
 load_all("../ggplyr")
-# splitting by grids - obsolete
-ggplot(test.data) + geom_point(aes(Fertility, Agriculture, color = Catholic))
-ggplot(test.data) + geom_point(aes(grid(Fertility, 3), grid(Agriculture, 3), 
-  color = Catholic))
-ggplot(test.data) + geom_point(aes(grid(Fertility, 3), grid(Agriculture, 3), 
-  color = Catholic), position = position_jitter(width = 1, height = 1))
-ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
-  color = Catholic)), aes(I(grid(Fertility, 3)), I(grid(Agriculture, 3))), 
-  grid_by(Fertility, 3, Agriculture, 3))

### 2012-06-21 paper/demo.tex

--- 2012-05-30 paper/demo.tex
@@
 
-\title{A Minimal Demo of glyphmaps}
+\title{A Minimal Demo of \texttt{glyphmaps}}
 
@@
 This document demonstrates the basic features of the \texttt{glyphmaps} package (name may change).
-\texttt{glyphmaps} was written to create plots like the ones below.
-Each was made with \texttt{glyphmaps}.
-The first two plots recreate plots from \citet{wickham2011split}.
-The last two recreate plots from \citet{wickham2012glyph}.
-
-To run the examples contained in this paper, you'll have to download the \texttt{glyphmaps} development package from \url{https://github.com/garrettgman/ggplyr} and install it with
+\texttt{glyphmaps} was written to create plots like the ones below, which recreate plots similar to those in \citet{wickham2011split} and \citet{wickham2012glyph}.
+Each plot is made with \texttt{glyphmaps}.
+To run the examples in this document please download the \texttt{glyphmaps} package with the code below.
+(Note: Install won't work until I finish documenting the namespace.
+But cloning and load_all() DOES work.)
 
@@
 \hlfunctioncall{library}\hlkeyword{(}\hlsymbol{devtools}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{load\usebox{\hlnormalsizeboxunderscore}all}\hlkeyword{(}\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
+\hlstd{}\hlfunctioncall{install\usebox{\hlnormalsizeboxunderscore}github}\hlkeyword{(}\hlstring{"{}ggplyr"{}}\hlkeyword{,}{\ }\hlstring{"{}garrettgman"{}}\hlkeyword{,}{\ }\hlstring{"{}glyphmaps"{}}\hlkeyword{)}\mbox{}
 \normalfont
@@
 \begin{verbatim}
-## Error: Can't find package glyphmaps
+## Installing github repo(s) ggplyr/glyphmaps from garrettgman
+## Installing ggplyr.zip from https://github.com/garrettgman/ggplyr/zipball
+## Installing glyphmaps
+## Error: Command failed (1)
+\end{verbatim}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{library}\hlkeyword{(}\hlsymbol{glyphmaps}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+## Error: there is no package called 'glyphmaps'
 \end{verbatim}
@@
 
-I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \begin{verbatim}
-
-Error in paste(repo, branch, sep = "/", collapse = ", ") : 
-  argument "repo" is missing, with no default
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{seasons}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}us}{\ }\hlkeyword{+}{\ }\hlfunctioncall{grid}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}line}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{time}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{pred}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{grid.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{lon}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{x.nbin}{\ }\hlargument{=}{\ }\hlnumber{58}\hlkeyword{,}{\ }\hlargument{y.nbin}{\ }\hlargument{=}{\ }\hlnumber{25}\hlkeyword{,}{\ }\hlargument{ref}{\ }\hlargument{=}{\ }\hlfunctioncall{ref\usebox{\hlnormalsizeboxunderscore}box}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{fill}{\ }\hlargument{=}{\ }\hlsymbol{avg}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{alpha}{\ }\hlargument{=}{\ }\hlnumber{0.2}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlnumber{NA}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+## Error: could not find function "ggplot"
 \end{verbatim}
-.
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlsymbol{map\usebox{\hlnormalsizeboxunderscore}nasa}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}star}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{r}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{angle}{\ }\hlargument{=}{\ }\hlsymbol{date}\hlkeyword{,}{\ }\hlargument{x}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlnumber{0}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }{\ }{\ }\hlargument{fill}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{major.aes}{\ }\hlargument{=}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{glyph.by}{\ }\hlargument{=}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}long"{}}\hlkeyword{,}{\ }\hlstring{"{}lat"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+## Error: could not find function "ggplot"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 \section{Glyphs}
-The above graphs are similar in that they are all built around glyphs.
-I define a glyph as a geom that visually summarises information about a subset of data.
-Glyphs can be quite simple, or very complex.
-Often glyphs contain their own internal coordinate systems.
-The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
-
-\section{glyphmaps geoms}
-\texttt{glyphmaps} introduces a number of new geoms that make it easy to create glyphs.
-These geoms can be used in conjunction with \texttt{ggplot2} functions to create layered plots.
-\texttt{glyphmaps} geoms behave just like \texttt{ggplot2} geoms, but take two additional arguments, described in Section~\ref{sec:args}.
-
-\subsection{new arguments}
-\label{sec:args}
-Each \texttt{glyphmaps} geom takes two new arguments, \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- and \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'reference' not found
-
-\end{verbatim}
-. \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- controls how the data will be divided into subsets to be paired with individual glyphs. \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'reference' not found
-
-\end{verbatim}
- lets the user draw reference features with the glyphs to facilitate comparisons.
-
-\begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- takes the name of the variable(s) to be used to divide the data into the subsets that will correspond to different glyphs. The \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- argument works just like a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object '.vars' not found
-
-\end{verbatim}
- argument in a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'ddply' not found
-
-\end{verbatim}
- call. (In fact, the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyphs' not found
-
-\end{verbatim}
- argument in a \texttt{glyphmaps} geom will be passed into a \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'ddply' not found
-
-\end{verbatim}
- call as the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object '.vars' not found
-
-\end{verbatim}
- argument.)
-
-\texttt{glyphmaps} provides methods for drawing four types of reference objects behind glyphs: boxes, vertical lines, horizontal lines, and corner points.
-These objects help viewers make comparisons between data inside different glyphs.
-
-\section{Generalized glyphs}
-Users are not limited to the pre-packaged glyph geoms. Any single layer plot can be turned into a glyph using the \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'glyph' not found
-
-\end{verbatim}
- and \begin{verbatim}
-
-Error in eval(expr, envir, enclos) : object 'plyr' not found
-
-\end{verbatim}
- functions.
-
-The plyr function also allows users to key aesthetics to subgroups of data.
-
-\section{Discussion - glyphs, embedded graphics, hierarchies}
+The common feature of these graphs is that they are built around glyphs.
 Glyphs are geometric objects (i.e, geoms) designed to display information within the geom.
@@
 In reality, all geoms are a type of glyph, but the term glyph is usually reserved for complicated geoms, such as those that contain their own internal coordinate systems.
+The star glyphs in Figure~\ref{fig:star-glyphs} illustrate how glyphs can contain an internal (minor) coordinate system and can still be plotted in an external (major) coordinate system.
 
@@
 
+To run the examples contained below, you'll have to download the \texttt{ggplyr} development package from \url{https://github.com/garrettgman/ggplyr} and install it with
+
+I haven't yet built the namespace or completed the documentation files, because the exact syntax / function membership hasn't been settled. As a result, the package will fail \begin{verbatim}
+
+Error in paste(repo, branch, sep = "/", collapse = ", ") : 
+  argument "repo" is missing, with no default
+
+\end{verbatim}
+.
+
 \section{top down methods}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{set.seed}\hlkeyword{(}\hlnumber{1121}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlkeyword{(}\hlsymbol{x}{\ }\hlassignement{\usebox{\hlnormalsizeboxlessthan}-}{\ }\hlfunctioncall{rnorm}\hlkeyword{(}\hlnumber{20}\hlkeyword{)}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-##  [1]  0.14496  0.43832  0.15319  1.08494  1.99954 -0.81188  0.16027  0.58589  0.36009
-## [10] -0.02531  0.15088  0.11008  1.35968 -0.32699 -0.71638  1.80977  0.50840 -0.52746
-## [19]  0.13272 -0.15594
+The first element of \texttt{x} is \begin{verbatim}
+
+Error in eval(expr, envir, enclos) : object 'x' not found
+
 \end{verbatim}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-## [1] 0.3217
-\end{verbatim}
-\begin{flushleft}
-\ttfamily\noindent
-\hlfunctioncall{var}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
-\normalfont
-\end{flushleft}
-\begin{verbatim}
-## [1] 0.5715
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-The first element of \texttt{x} is 0.145. Boring boxplots
+. Boring boxplots
 and histograms recorded by the PDF device:
@@
 \hlstd{}{\ }{\ }{\ }{\ }\hlargument{tcl}{\ }\hlargument{=}{\ }\hlkeyword{-}\hlnumber{0.3}\hlkeyword{,}{\ }\hlargument{las}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\hspace*{\fill}\\
-\hlstd{}\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
+\hlstd{}\hlfunctioncall{boxplot}\hlkeyword{(}\hlsymbol{x}\hlkeyword{)}\mbox{}
 \normalfont
 \end{flushleft}
+\begin{verbatim}
+## Error: object 'x' not found
+\end{verbatim}
+\begin{flushleft}
+\ttfamily\noindent
+\hlfunctioncall{hist}\hlkeyword{(}\hlsymbol{x}\hlkeyword{,}{\ }\hlargument{main}{\ }\hlargument{=}{\ }\hlstring{"{}"{}}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+## Error: object 'x' not found
+\end{verbatim}
 \end{kframe}
-
-{\centering \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots1} \includegraphics[width=.4\linewidth]{figure/minimal-boring-plots2}
-
-}
-
 \end{knitrout}

### 2012-06-21 paper/paper.tex

@@
 
+\texttt{ply$\_$aes} wraps a layer like \texttt{grid} or \texttt{glyph} and causes the aesthetics to be calculated in a groupwise manner.
+Users can specify the groups to use, or \texttt{ply$\_$aes} will default to any groupings provided by glyph.by or group.by parameters, or a group aesthetic.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }normal{\ }\usebox{\hlnormalsizeboxbackslash}texttt\usebox{\hlnormalsizeboxopenbrace}ggplot2\usebox{\hlnormalsizeboxclosebrace}{\ }behavior}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{hwy}\hlkeyword{,}{\ }\hlsymbol{cty}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}{\ }\hlkeyword{+}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}wrap}\hlkeyword{(}\hlkeyword{\urltilda{}}\hlsymbol{cyl}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }groupwise{\ }aesthetics}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{mpg}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{ply\usebox{\hlnormalsizeboxunderscore}aes}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlsymbol{hwy}\hlkeyword{,}{\ }\hlsymbol{cty}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlfunctioncall{rank}\hlkeyword{(}\hlsymbol{displ}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}cyl"{}}\hlkeyword{)}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}wrap}\hlkeyword{(}\hlkeyword{\urltilda{}}\hlsymbol{cyl}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise11} \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise12}
+
+}
+
+\end{knitrout}
+
+\texttt{ply$\_$aes} doesn't actually create embedded plots, but is closely related to the glyphing process.
+By default, \texttt{ggplot2} does not calculate statistics in a groupwise manner.
+However, such calculations can be useful in certain situations.
+For example, \texttt{ply$\_$aes} is built into \texttt{glyph} and \texttt{grid} by default; embedded plots quickly become boring when aesthetics do not vary from subplot to subplot.
+Groupwise aesthetics can be interesting in other situations as well.
+For example, Figure~\ref{fig:groupwise2} demonstrates how groupwise aesthetics can be used to avoid overplotting.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }overplotted}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{ozone}\hlkeyword{,}{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{data}{\ }\hlargument{=}{\ }\hlsymbol{nasa}\hlkeyword{)}\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}{\ }using{\ }ply\usebox{\hlnormalsizeboxunderscore}aes{\ }to{\ }plot{\ }just{\ }summaries}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{ggplot}\hlkeyword{(}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{ply\usebox{\hlnormalsizeboxunderscore}aes}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlfunctioncall{mean}\hlkeyword{(}\hlsymbol{ozone}\hlkeyword{)}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{color}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{data}{\ }\hlargument{=}{\ }\hlsymbol{nasa}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise21} \includegraphics[width=0.49\textwidth]{figure/minimal-groupwise22}
+
+}
+
+\end{knitrout}
+
+\section{speed considerations}
+Embedding plots is computationally expensive.
+Such plots require much of both the computational engine of R and the graphics device.
+Hence, it is not unreasonable to expect \texttt{glyphmaps} to be slow, which it can be.
+However, relative to similar visual tasks \texttt{glyphmaps} is surprisingly fast.
+The final code chunk compares the speed of plotting forty thousand points in an embedded plot to plotting the same points in a facetted plot.
+\texttt{glyphmaps} performs quite well in comparison.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{flushleft}
+\ttfamily\noindent
+\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }embedded{\ }plot}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{system.time}\hlkeyword{(}\hlfunctioncall{print}\hlkeyword{(}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{long}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{,}{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{lat}\hlkeyword{[}\hlnumber{1}\hlkeyword{]}\hlkeyword{)}\hlkeyword{,}{\ }\hlfunctioncall{c}\hlkeyword{(}\hlstring{"{}lat"{}}\hlkeyword{,}{\ }\hlstring{"{}long"{}}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+##    user  system elapsed 
+##  12.460   2.282  15.981 
+\end{verbatim}
+\begin{flushleft}
+\ttfamily\noindent
+\hspace*{\fill}\\
+\hlstd{}\hlcomment{\usebox{\hlnormalsizeboxhash}\usebox{\hlnormalsizeboxhash}{\ }facetted{\ }plot}\hspace*{\fill}\\
+\hlstd{}\hlfunctioncall{system.time}\hlkeyword{(}\hlfunctioncall{print}\hlkeyword{(}\hlfunctioncall{ggplot}\hlkeyword{(}\hlsymbol{nasa}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{glyph}\hlkeyword{(}\hlfunctioncall{geom\usebox{\hlnormalsizeboxunderscore}point}\hlkeyword{(}\hlfunctioncall{aes}\hlkeyword{(}\hlargument{x}{\ }\hlargument{=}{\ }\hlsymbol{surftemp}\hlkeyword{,}\hspace*{\fill}\\
+\hlstd{}{\ }{\ }\hlargument{y}{\ }\hlargument{=}{\ }\hlsymbol{temperature}\hlkeyword{)}\hlkeyword{,}{\ }\hlargument{size}{\ }\hlargument{=}{\ }\hlnumber{1}\hlkeyword{/}\hlnumber{2}\hlkeyword{)}{\ }\hlkeyword{+}{\ }\hlfunctioncall{facet\usebox{\hlnormalsizeboxunderscore}grid}\hlkeyword{(}\hlsymbol{long}\hlkeyword{\urltilda{}}\hlsymbol{lat}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\hlkeyword{)}\mbox{}
+\normalfont
+\end{flushleft}
+\begin{verbatim}
+## Error: argument "major.aes" is missing, with no default
+## Timing stopped at: 0.012 0.007 0.019 
+\end{verbatim}
+\end{kframe}
+
+{\centering \includegraphics[width=0.49\textwidth]{figure/minimal-<<speed-test}
+
+}
+
+\end{knitrout}
+
 \bibliographystyle{plainnat}

### 2012-06-21 R/geom_subplot.r
Changed naming scheme from glyphmaps to subplots

@@ -84,17 +84,17 @@ glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95),
     
   if (is.null(reference)) {
     if (ply.aes) {
-  	  ply_aes(glayer(layer))
+  	  ply_aes(sp_layer(layer))
     } else {
-      glayer(layer)
+      sp_layer(layer)
     }
   } else {
   	ref.layer <- reference(layer, "glyph", major.aes, glyph.by, width, height, 
   	  merge.overlaps)
     if (ply.aes) {
-  	  list(ref.layer, ply_aes(glayer(layer)))
+  	  list(ref.layer, ply_aes(sp_layer(layer)))
     } else {
-      list(ref.layer, glayer(layer))
+      list(ref.layer, sp_layer(layer))
     }
   }
 }
@@ -155,10 +155,10 @@ assign_glyphs <- function(., data) {
   data
 }
 
-# Calculate final positions in a glayer
+# Calculate final positions in a sp_layer
 # 
-# combine_glyphs calculates the final positions for every location in a glayer.
-# glayer_build first builds each glyph separately as if it were a facet. If 
+# combine_glyphs calculates the final positions for every location in a sp_layer.
+# sp_layer_build first builds each glyph separately as if it were a facet. If 
 # plotted, these glyphs would overlap with each other. combine_glyph relocates 
 # each glyph based on the global positions stored in the layer's embed variable.
 combine_glyphs <- function(., data) {  

### 2012-06-21 R/geom_subplot2d.r

@@ -60,7 +60,7 @@ grid <- function(layer, grid.aes = ggplot2::aes(), x.nbin = 10, y.nbin = 10,
   x_scale = identity, y_scale = identity, width.adjust = 0.95, 
   height.adjust = 0.95, reference = ref_box(), ply.aes = TRUE, .ref = FALSE) {
   
-  if (is.glayer(layer) || is.list(layer)) {
+  if (is.sp_layer(layer) || is.list(layer)) {
     stop("Cannot grid glayer", call. = TRUE)
   }
  # if ("plyr" %in% ls(layer)) {
@@ -94,17 +94,17 @@ grid <- function(layer, grid.aes = ggplot2::aes(), x.nbin = 10, y.nbin = 10,
   
   if (is.null(reference)) {
     if (ply.aes) {
-      ply_aes(glayer(layer))
+      ply_aes(sp_layer(layer))
     } else {
-      glayer(layer)
+      sp_layer(layer)
     }
   } else {
     ref.layer <- reference(layer, "grid", grid.aes, x.nbin = x.nbin, 
       y.nbin = y.nbin)
     if (ply.aes) {
-      list(ref.layer, ply_aes(glayer(layer)))
+      list(ref.layer, ply_aes(sp_layer(layer)))
     } else {
-      list(ref.layer, glayer(layer))
+      list(ref.layer, sp_layer(layer))
     }
   }
 }

### 2012-06-21 R/ggsubplot-class.r

@@ -1,0 +1,81 @@
+check_ggsubplot <- function(object) {
+	errors <- character()
+	if (!is(object@.Data, "ggplot")) {
+		msg <- "ggsubplot must be a ggplot object."
+		errors <- c(errors, msg)
+	}
+	if (length(errors) == 0) 
+		TRUE
+	else
+		errors
+}
+
+#' list S4 class
+#' 
+#' @name list-class
+#' @aliases list
+#'
+#' @exportClass list
+NULL
+
+#' ggplot S4 class
+#' 
+#' @name ggplot-class
+#' @aliases ggplot
+#'
+#' @exportClass ggplot
+setOldClass(c("ggplot", "list"))
+
+#' ggsubplot class
+#'
+#' a ggsubplot object is a ggplot object that has been extended to include methods 
+#' for embedding subplots when plotting. 
+#'
+#' @name ggsubplot-class
+#' @rdname ggsubplot-class
+#' @exportClass ggsubplot
+#' @aliases show,ggsubplot-method
+#' @aliases print,ggsubplot-method
+#' @aliases show,ggsubplot-method
+setClass("ggsubplot", contains = c("ggplot"), validity = check_ggsubplot)
+
+#' @export
+setMethod("show", signature(object = "ggsubplot"), function(object){
+	print(object)
+})
+
+#' @S3method print ggsubplot
+print.ggsubplot <- function(x, newpage = is.null(vp), vp = NULL, ...) {
+    ggplot2:::set_last_plot(x)
+    if (newpage) 
+        grid::grid.newpage()
+    data <- ggsubplot_build(x)
+    gtable <- ggplot2::ggplot_gtable(data)
+    if (is.null(vp)) {
+        grid::grid.draw(gtable)
+    }
+    else {
+        if (is.character(vp)) 
+            grid::seekViewport(vp)
+        else grid::pushViewport(vp)
+        grid::grid.draw(gtable)
+        grid::upViewport()
+    }
+    invisible(data)
+}
+
+
+#' Create a ggsubplot object
+#' 
+#' glyph_plot gives a ggplot object the S4 class `ggsubplot', see 
+#' \code{\link{ggsubplot-class}}. ggsubplot denotes ggplot objects that contain 
+#' extra information to be used to embed subplots when plotting. ggsubplot 
+#' objects have similar, but different print and build methods than ggplot2 
+#' objects.
+#' 
+#' @param ggplot a ggplot object
[…]

### 2012-06-21 R/ggsubplot_build.r

@@ -1,33 +1,33 @@
-#' Build a glyphs object for rendering
+#' Build a ggsubplot object for rendering
 #' 
-#' glyph_build takes a glyph plot object (class glyphs), and performs all steps 
+#' ggsubplot_build takes a ggsubplot object and performs all steps 
 #' necessary to produce an object that can be rendered. This function outputs 
 #' two pieces: a list of data frames (one for each layer), and a panel object, 
 #' which contain all information about axis limits, breaks, etc.
 #' 
 #' @keywords internal
-#' @param layer an object of class glayer
-#' @seealso print.glyphs and \code{\link{glayer_build}} for functions that contain the complete 
-#' set of steps for generating a glyphs plot
+#' @param layer an object of class sp_layer
+#' @seealso print.ggsubplot and \code{\link{sp_layer_build}} for functions that 
+#' contain the complete set of steps for generating a ggsubplot plot
 #' @export
-glyph_build <- function(plot){
+ggsubplot_build <- function(plot){
   if (length(plot$layers) == 0) stop("No layers in plot", call.=FALSE)
   if (!identical(plot$facet, ggplot2::facet_null())) {
-  	stop("glyphs do not support facetting", call. = FALSE)
+  	stop("ggsubplots do not support facetting", call. = FALSE)
   }
 	
   plot <- ggplot2:::plot_clone(plot)
   layers <- plot$layers
   layers <- propogate_data(layers, plot$data)
 	
-  # separate into glayers and normal layers
-  gls <- unlist(lapply(layers, is.glayer))
-  if (all(!gls)) return(ggplot2::ggplot_build(plot))
-  if (all(gls) && sum(gls) == 1) return(glayer_build(layers[[gls]]))
-  glayers <- layers[gls]
-  plot$layers <- layers[!gls]
-  gl.order <- seq_along(layers)[gls]
-  nl.order <- seq_along(layers)[!gls]
+  # separate into sp_layers and normal layers
+  spls <- unlist(lapply(layers, is.sp_layer))
+  if (all(!spls)) return(ggplot2::ggplot_build(plot))
+  if (all(spls) && sum(spls) == 1) return(sp_layer_build(layers[[spls]]))
+  splayers <- layers[spls]
+  plot$layers <- layers[!spls]
+  spl.order <- seq_along(layers)[spls]
[…]

### 2012-06-21 R/glayer-class.r

@@ -1,146 +1,0 @@
-#' @include glyphs-class.r
-NULL
-
-check_glayer <- function(object) {
-	errors <- character()
-	if (!proto::is.proto(object@layer)) {
-		msg <- "glayer must be a proto object."
-		errors <- c(errors, msg)
-	}
-  if (!("embed" %in% ls(object@layer))) {
-    msg <- "glayers should contain an `embed' variable. Try building with glyph() or grid()"
-    errors <- c(errors, msg)
-  }
-	if (length(errors) == 0) 
-		TRUE
-	else
-		errors
-}
-
-#' environment S4 class
-#' 
-#' @name environment-class
-#' @aliases environment
-#' 
-#' @exportClass environment
-NULL
-
-#' proto S4 class
-#' 
-#' @name proto-class
-#' @aliases proto
-#' 
-#' @exportClass proto
-setOldClass(c("proto", "environment"))
-
-#' glayer class
-#'
-#' glayers are layers made with glyphmaps methods. They are equivalent to the 
-#' layers made by ggplot2 functions in all ways except that they contain extra 
-#' information that is used to divide the data into subplots and locate those 
-#' subplots witihn the layer when plotting.
-#'
-#' @name glayer-class
-#' @rdname glayer-class
-#' @exportClass glayer
-#' @aliases show,glayer-method
-#' @aliases c,glayer-method
-#' @aliases rep,glayer-method
-#' @aliases ls,glayer-method
-#' @aliases [,glayer-method
-#' @aliases [<-,glayer-method
-#' @aliases $,glayer-method
-#' @aliases $<-,glayer-method
-#' @aliases +,ggplot,glayer-method
-#' @aliases +,glyphs,glayer-method
-#' @aliases ggtransform,glayer-method
-setClass("glayer", representation(layer = "proto"), validity = check_glayer)
-
-#' @export
-setMethod("show", signature(object = "glayer"), function(object) {
-	print(object@layer) 
-})
-
-#' @export
-setMethod("c", signature(x = "glayer"), function(x, ...){
-	# c(get_layer(x), unlist(lapply(list(...), get_layer)))
-	stop("object of type 'glayer' is not subsettable")
-})
-
-#' @export
-setMethod("rep", signature(x = "glayer"), function(x, ...){
-	stop("object of type 'glayer' is not subsettable")
-})
-
[…]

### 2012-06-21 R/glyphs-class.r

@@ -1,80 +1,0 @@
-check_glyphs <- function(object) {
-	errors <- character()
-	if (!is(object@.Data, "ggplot")) {
-		msg <- "glyphs must be a ggplot object."
-		errors <- c(errors, msg)
-	}
-	if (length(errors) == 0) 
-		TRUE
-	else
-		errors
-}
-
-#' list S4 class
-#' 
-#' @name list-class
-#' @aliases list
-#'
-#' @exportClass list
-NULL
-
-#' ggplot S4 class
-#' 
-#' @name ggplot-class
-#' @aliases ggplot
-#'
-#' @exportClass ggplot
-setOldClass(c("ggplot", "list"))
-
-#' glyphs class
-#'
-#' a glyphs object is a ggplot object that has been extended to include methods 
-#' for embedding subplots when plotting. 
-#'
-#' @name glyphs-class
-#' @rdname glyphs-class
-#' @exportClass glyphs
-#' @aliases show,glyphs-method
-#' @aliases print,glyphs-method
-#' @aliases show,glyphs-method
-setClass("glyphs", contains = c("ggplot"), validity = check_glyphs)
-
-#' @export
-setMethod("show", signature(object = "glyphs"), function(object){
-	print(object)
-})
-
-#' @S3method print glyphs
-print.glyphs <- function(x, newpage = is.null(vp), vp = NULL, ...) {
-    ggplot2:::set_last_plot(x)
-    if (newpage) 
-        grid::grid.newpage()
-    data <- glyph_build(x)
-    gtable <- ggplot2::ggplot_gtable(data)
-    if (is.null(vp)) {
-        grid::grid.draw(gtable)
-    }
-    else {
-        if (is.character(vp)) 
-            grid::seekViewport(vp)
-        else grid::pushViewport(vp)
-        grid::grid.draw(gtable)
-        grid::upViewport()
-    }
-    invisible(data)
-}
-
-
-#' Create a glyphs object
-#' 
-#' glyph_plot gives a ggplot object the S4 class `glyphs', see 
-#' \code{\link{glyphs-class}}. glyphs denotes ggplot objects that contain extra 
-#' information to be used to embed subplots when plotting. glyphs objects have 
-#' similar, but different print and build methods than ggplot2 objects.
-#' 
-#' @param ggplot a ggplot object
-#' #' @export glyph_plot
-glyph_plot <- function(ggplot) {
-	new("glyphs", ggplot)
-}
-

### 2012-06-21 R/ply_aes.r

@@ -9,10 +9,10 @@
 #' 
 #' Users may specify which groupings to use through the .vars 
 #' argument. If this argument is left NULL, ply_aes will search for and use a 
-#' group aes, a glyphing or gridding criteria (in a glayer), a facetting 
+#' group aes, a glyphing or gridding criteria (in a sp_layer), a facetting 
 #' criteria, or any combination of these that it finds. 
 #' 
-#' @param layer a ggplot2 layer or glayer object. This layer's aesthetics will 
+#' @param layer a ggplot2 layer or sp_layer object. This layer's aesthetics will 
 #' be computed groupwise, but the layer will remain the same in every other 
 #' respect.
 #' @param .vars variable names to group by (optional), stored as a character 
@@ -27,8 +27,8 @@ ply_aes.list <- function(layer, .vars = NULL) {
   lapply(layer, ply_aes, .vars)
 }
 
-#' @S3method ply_aes glayer
-ply_aes.glayer <- function(layer, .vars = NULL) {
+#' @S3method ply_aes sp_layer
+ply_aes.sp_layer <- function(layer, .vars = NULL) {
   if (!is.null(.vars)) {
     layer$plyr <- list(ply.by = .vars)
   } else {

### 2012-06-21 R/sp_layer-class.r

@@ -1,0 +1,146 @@
+#' @include ggsubplot-class.r
+NULL
+
+check_sp_layer <- function(object) {
+	errors <- character()
+	if (!proto::is.proto(object@layer)) {
+		msg <- "sp_layer must be a proto object."
+		errors <- c(errors, msg)
+	}
+  if (!("embed" %in% ls(object@layer))) {
+    msg <- "sp_layer should contain an `embed' variable. Try building with geom_subplot() or geom_subplot2d()"
+    errors <- c(errors, msg)
+  }
+	if (length(errors) == 0) 
+		TRUE
+	else
+		errors
+}
+
+#' environment S4 class
+#' 
+#' @name environment-class
+#' @aliases environment
+#' 
+#' @exportClass environment
+NULL
+
+#' proto S4 class
+#' 
+#' @name proto-class
+#' @aliases proto
+#' 
+#' @exportClass proto
+setOldClass(c("proto", "environment"))
+
+#' sp_layer class
+#'
+#' sp_layers are layers made with ggsubplot methods. They are equivalent to the 
+#' layers made by ggplot2 functions in all ways except that they contain extra 
+#' information that is used to divide the data into subplots and locate those 
+#' subplots witihn the layer when plotting.
+#'
+#' @name sp_layer-class
+#' @rdname sp_layer-class
+#' @exportClass sp_layer
+#' @aliases show,sp_layer-method
+#' @aliases c,sp_layer-method
+#' @aliases rep,sp_layer-method
+#' @aliases ls,sp_layer-method
+#' @aliases [,sp_layer-method
+#' @aliases [<-,sp_layer-method
+#' @aliases $,sp_layer-method
+#' @aliases $<-,sp_layer-method
+#' @aliases +,ggplot,sp_layer-method
+#' @aliases +,glyphs,sp_layer-method
+#' @aliases ggtransform,sp_layer-method
+setClass("sp_layer", representation(layer = "proto"), validity = check_sp_layer)
+
+#' @export
+setMethod("show", signature(object = "sp_layer"), function(object) {
+	print(object@layer) 
+})
+
+#' @export
+setMethod("c", signature(x = "sp_layer"), function(x, ...){
+	# c(get_layer(x), unlist(lapply(list(...), get_layer)))
+	stop("object of type 'sp_layer' is not subsettable")
+})
+
+#' @export
[…]

### 2012-06-21 R/sp_layer_build.r

@@ -1,19 +1,21 @@
-#' Build a glayer for rendering
+#' Build a sp_layer for rendering
 #' 
-#' glayer_build takes a glyph layer (class glayer), and performs all steps 
-#' necessary to produce an object that can be rendered. This function outputs 
-#' two pieces: a list of data frames (one for each layer), and a panel object, 
-#' which contain all information about axis limits, breaks, etc.
+#' sp_layer_build takes a layer of subplots (class sp_layer), and performs all 
+#' steps necessary to produce an object that can be rendered. This function 
+#' outputs two pieces: a list of data frames (one for each layer), and a panel 
+#' object, which contains all information about axis limits, breaks, etc.
 #' 
-#' If the glayer is accompanied by regular layers, glayer_build will be used in 
-#' conjunction with \code{\link{glyph_build}} to build the plot for rendering.
+#' If the sp_layer is accompanied by regular layers, sp_layer_build will be used 
+#' in conjunction with \code{\link{ggsubplot_build}} to build the plot for 
+#' rendering.
 #' 
 #' @keywords internal
-#' @param layer an object of class glayer
-#' @seealso print.glyphs and \code{\link{glyph_build}} for 
-#' functions that contain the complete set of steps for generating a glyphs plot
+#' @param layer an object of class sp_layer
+#' @seealso print.ggsubplot and \code{\link{ggsubplot_build}} for 
+#' functions that contain the complete set of steps for generating a ggsubplot 
+#' plot
 #' @export
-glayer_build <- function(layer) {
+sp_layer_build <- function(layer) {
   if (!("embed" %in% ls(layer))) {
     stop("layer does not have embedded subplots")
   }
@@ -54,7 +56,7 @@ glayer_build <- function(layer) {
   minor
 }
 
-#' Format data from a facet plot to use in a glyph plot
+#' Format data from a facet plot to use in a ggsubplot plot
 #' 
 #' unpanel replaces the PANEL variable of a data frame with a GLYPH variable. It 
 #' adjusts the data frame's group variable to retain the grouping information 

### 2012-06-21 demo.r

@@ -2,7 +2,7 @@
 ###             geom_scatterplots               ###
 ###################################################
 library(devtools)
-install_github("ggplyr", "garrettgman", "glyphmaps")
+install_github("ggplyr", "garrettgman", "ggsubplot")
 library(glyphmaps)
 load_all("../ggplyr")
 load("data/nasa.RData")
@@ -172,12 +172,12 @@ ggplot(mpg) + GeomCoxcomb$new(mapping = aes(angle = trans, fill = lat,
 ggplot(mpg) + geom_coxcomb(aes(angle = trans, fill = lat))
 
 
-ggplot(seasons) + 
+system.time(print(ggplot(seasons) + 
   glyph(
     geom_line(aes(x = time, y = pred)), 
     major = aes(lon[1], lat[1]), glyph.by = "stn", 
     height = 1, width = 2,
-    ref = ref_box(aes(fill = avg)), merge = TRUE)
+    ref = ref_box(aes(fill = avg)), merge = TRUE)))
 
 ###########################################
 ###          not yet working            ###

### 2012-06-21 R/geom_subplot.r
wrote geom_subplot

@@ -99,24 +99,129 @@ glyph <- function(layer, major.aes, glyph.by = NULL, width = rel(0.95),
   }
 }
 
-# Assigns glyph membership to rows
+#' Create a layer of embedded subplots
+#' 
+#' @param mapping An aesthetic mapping, usually constructed with 
+#' \code{\link[ggplot2]{aes}}. This mapping determines where in the major x and 
+#' y axes each glyph will be position. Only x, y, group, and subplot aesthetics 
+#' will be used. All other aesthetics will be ignored - consider placing them in 
+#' the subplot's mapping instead.
+#' @param width numeric or rel object. The width of each glyph. If width is 
+#' numeric, the glyph will be drawn with a width equal to width units on the x 
+#' axis. If width is of class \code{\link{rel}}, glyph will 
+#' attempt to assign an inuitive width based on the number of total glyphs and 
+#' their placement within the plot. The width can be scaled relative to this 
+#' intuitive width by changing the value of the rel object.
+#' @param height numeric or rel object. The height of each glyph. Height behaves 
+#' the same way as width, but applies to the y dimension.
+#' @param x_scale function. The scaling to use for the x axis within each glyph. 
+#' If x_scale equals \code{\link{identity}}(default), the x limits within each 
+#' glyph will correspond to the range of x across all glyphs. This aids 
+#' comparison because each glyph will use the same scale. If x_scale equals 
+#' \code{\link{free}}, each glyph will use its own x scale. The limits of this 
+#' scale will be set to the range of x values in that glyph.
+#' @param y_scale function. y_scale behaves the same as x_scale but controls the 
+#' scales for the y axis within each glyph.
+#' @param position character. A string that specifies which position adjustment 
+#' should be used. geom_subplot only recognizes "identity" and "merge". 
+#' @param reference function. Function used to create reference objects for 
+#' glyphs. If NULL, no reference objects are used. Reference objects are plotted 
+#' on a layer beneath the glyphs. They provide a consistent frame of reference 
+#' to aid comparisons between the glyphs. Functions that create reference 
[…]

### 2012-06-21 R/ggsubplot_build.r

@@ -19,11 +19,12 @@ ggsubplot_build <- function(plot){
   plot <- ggplot2:::plot_clone(plot)
   layers <- plot$layers
   layers <- propogate_data(layers, plot$data)
+  env <- plot$plot_env
 	
   # separate into sp_layers and normal layers
   spls <- unlist(lapply(layers, is.sp_layer))
   if (all(!spls)) return(ggplot2::ggplot_build(plot))
-  if (all(spls) && sum(spls) == 1) return(sp_layer_build(layers[[spls]]))
+  if (all(spls) && sum(spls) == 1) return(sp_layer_build(layers[[spls]], env))
   splayers <- layers[spls]
   plot$layers <- layers[!spls]
   spl.order <- seq_along(layers)[spls]
@@ -38,7 +39,7 @@ ggsubplot_build <- function(plot){
   # build glyph layers (embedded plots)
   embedded <- list()
   for (i in seq_along(splayers)) {
-    embedded[[i]] <- sp_layer_build(splayers[[i]])
+    embedded[[i]] <- sp_layer_build(splayers[[i]], env)
   }
 	
 	
@@ -121,4 +122,18 @@ propogate_data <- function(layers, plot_data) {
 		layer
 	}
 	lapply(layers, ensure_data)
+}
+
+
+#' Ensure each layer contains all aesthetic mappings that affect it.
+#' 
+#' propogate_aes checks for aesthetics defined at the global level of a plot 
+#' that affect a layer. Propogate_aes moves such aesthetics into the layer's 
+#' mapping. 
+#' @param layer ggplot2 layer objects
+#' @param plot_mapping the global data set for a ggplot2 plot
+propogate_aes <- function(layer, plot_mapping) {
+  layer$mapping <- c(layer$mapping, 
+    plot_mapping[setdiff(names(plot_mapping), names(layer$mapping))])
+  layer
 }

### 2012-06-21 R/sp_layer_build.r

@@ -15,20 +15,20 @@
 #' functions that contain the complete set of steps for generating a ggsubplot 
 #' plot
 #' @export
-sp_layer_build <- function(layer) {
+sp_layer_build <- function(layer, plot.env) {
   if (!("embed" %in% ls(layer))) {
     stop("layer does not have embedded subplots")
   }
-	 
+  
   layer <- layer_clone(layer)
-  layer$data <- layer$assign_glyphs(layer$data)
+  layer$data <- layer$assign_subplots(layer$data, plot.env)
   minor <- ggplot2::ggplot_build(ggplot2::ggplot() + layer + 
-    ggplot2::facet_wrap("GLYPH")) 
+    ggplot2::facet_wrap("SUBPLOT")) 
 
   ### combine subplots (minor) into single plot
   # data
   data <- unpanel(minor$data[[1]])
-  data <- layer$combine_glyphs(data)
+  data <- layer$combine_subplots(data)
   data$PANEL <- 1L
 	
   # panel
@@ -69,7 +69,7 @@ unpanel <- function(df) {
   if (!is.null(df$group)) {
     df$group <- interaction(df$group, df$PANEL)
   } 
-  df$GLYPH <- as.numeric(as.character(df$PANEL))
+  df$SUBPLOT <- as.numeric(as.character(df$PANEL))
   df$PANEL <- NULL
   df
 }

### 2012-06-21 demo.r

@@ -28,11 +28,21 @@ ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture,
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
   major = aes(mean(Fertility), mean(Education)))
 
+ggplot(test.data) + geom_subplot(aes(mean(Fertility), mean(Education), 
+  group = interaction(long, lat), subplot = geom_point(aes(Fertility, 
+  Agriculture, color = rank(Catholic)), size = 3)))
+
 # embed bars (categorical with stats)
 ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
   aes(x = mean(displ), y = mean(cty)), c("year"), y_scale = free, 
   width = 1/4, height = 1/4)
+
+ggplot(mpg) + geom_subplot(aes(mean(displ), mean(cty), group = year,
+  subplot = geom_bar(aes(trans, fill = year))), y_scale = free, width = 1/4, 
+  height = 1/4)
   
+
+
 # reference boxes
 # boxes
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 

### 2012-06-21 R/geom_subplot.r
converted ref_boxes to use geom_subplot and geom_subplot2d. Removed ref_points because they seemed too hard to interpret visually.

@@ -157,7 +157,7 @@ geom_subplot <- function(mapping, width = rel(0.95), height = rel(0.95),
     merge.overlaps <- FALSE
   }
   
-  layer <- extract_layer(mapping$subplot)
+  layer <- extract_layer(mapping$subplot, parent.frame())
   mapping$subplot <- NULL
   layer$embed <- list(width = width, height = height, x_scale = x_scale, 
     y_scale = y_scale, merge.overlaps = merge.overlaps, major.aes = mapping)
@@ -172,7 +172,7 @@ geom_subplot <- function(mapping, width = rel(0.95), height = rel(0.95),
       sp_layer(layer)
     }
   } else {
-    ref.layer <- reference(layer, "glyph", mapping, width, height, position)
+    ref.layer <- reference(layer, "subplot", mapping, width, height, position)
     if (ply.aes) {
       list(ref.layer, ply_aes(sp_layer(layer)))
     } else {
@@ -181,8 +181,8 @@ geom_subplot <- function(mapping, width = rel(0.95), height = rel(0.95),
   }
 }
 
-extract_layer <- function(subplot_aes) {
-  layer <- eval(subplot_aes, envir = parent.frame(), enclos = parent.frame(3))
+extract_layer <- function(subplot_aes, env) {
+  layer <- eval(subplot_aes, env)
   if ("sp_plot" %in% class(layer) | "ggsubplot" %in% class(layer)) {
     stop("Cannot place subplots inside subplots", call. = FALSE)
   }

### 2012-06-21 R/ggsubplot_build.r

@@ -133,7 +133,9 @@ propogate_data <- function(layers, plot_data) {
 #' @param layer ggplot2 layer objects
 #' @param plot_mapping the global data set for a ggplot2 plot
 propogate_aes <- function(layer, plot_mapping) {
-  layer$mapping <- c(layer$mapping, 
+  mapping <- c(layer$mapping, 
     plot_mapping[setdiff(names(plot_mapping), names(layer$mapping))])
+  class(mapping) <- "uneval"
+  layer$mapping <- mapping
   layer
 }

### 2012-06-21 R/ref_box.r

@@ -26,8 +26,8 @@
 #' \code{\link{ref_points}}
 #' @export
 ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {	
-  function(layer, type, major.aes, glyph.by = NULL, width = rel(1), 
-  height = rel(1), merge.overlaps = FALSE, x.nbin = 10, y.nbin = 10) {
+  function(layer, type, major.aes, width = rel(1), height = rel(1), 
+    position = "identity", x.nbin = 10, y.nbin = 10) {
   	
   	def_aes <- ggplot2::aes(xmin = -1, xmax = 1, ymin = -1, ymax = 1)
   	mapping <- c(mapping, def_aes[setdiff(names(def_aes), names(mapping))])
@@ -38,13 +38,14 @@ ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {
   	if (is.null(mapping$colour)) rlayer$geom_params$colour <- color
   	if (!inherits(layer$data, "waiver")) rlayer$data <- layer$data
   	
+    major.aes$subplot <- rlayer
+    
     switch(type,
-      glyph = glyph(rlayer, major.aes = major.aes, glyph.by = glyph.by, 
-        width = width, height = height, merge.overlaps = merge.overlaps, 
-        reference = NULL, .ref = TRUE), 
-      grid = grid(rlayer, grid.aes = major.aes, x.nbin = x.nbin, 
-        y.nbin = y.nbin, width.adjust = 1, height.adjust = 1, 
-        reference = NULL, .ref = TRUE)
+      subplot = geom_subplot(mapping = major.aes, width = width, 
+        height = height, position = position, reference = NULL, .ref = TRUE), 
+      subplot2d = geom_subplot2d(mapping = major.aes, x.nbin = x.nbin, 
+        y.nbin = y.nbin, width.adjust = 1, height.adjust = 1, reference = NULL, 
+        .ref = TRUE)
     )
   }
 } 
@@ -76,8 +77,8 @@ ref_box <- function(mapping = NULL, fill = "grey90", color = "white", ...) {
 #' @seealso \code{\link{ref_box}}, \code{\link{ref_vline}} and \code{\link{ref_points}}
 #' @export
 ref_hline <- function(mapping = NULL, thickness = 0.2, fill = "white", ...) {	
-  function(layer, type, major.aes, glyph.by = NULL, width = rel(1), 
-  height = rel(1), merge.overlaps = FALSE, x.nbin = 10, y.nbin = 10) {
[…]

### 2012-06-21 demo.r

@@ -41,6 +41,13 @@ ggplot(mpg) + geom_subplot(aes(mean(displ), mean(cty), group = year,
   subplot = geom_bar(aes(trans, fill = year))), y_scale = free, width = 1/4, 
   height = 1/4)
   
+bars <- geom_bar(aes(trans, fill = year))
+bars3 <- bars
+test <- function(bars2) {
+ggplot(mpg) + geom_subplot(aes(mean(displ), mean(cty), group = year,
+  subplot = bars2), y_scale = free, width = 1/4, height = 1/4)
+}
+test(bars3)
 
 
 # reference boxes
@@ -51,23 +58,33 @@ ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture,
   major = aes(mean(Fertility), mean(Education)), ref = ref_box(aes(fill = 
   mean(Catholic)), alpha = 0.2))
 
+ggplot(test.data) + geom_subplot(aes(mean(Fertility), mean(Education), 
+  group = interaction(long, lat), subplot = geom_point(aes(Fertility, 
+  Agriculture, color = rank(Catholic)), size = 3)), 
+  ref = ref_box(aes(fill = mean(Catholic)), alpha = 0.2))
+
+
 # hlines
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
   major = aes(mean(Fertility), mean(Education)), ref = ref_hline(aes(fill = 
   mean(Catholic))))
 
+ggplot(test.data) + geom_subplot(aes(mean(Fertility), mean(Education), 
+  group = interaction(long, lat), subplot = geom_point(aes(Fertility, 
+  Agriculture, color = rank(Catholic)), size = 3)), 
+  ref = ref_hline(aes(fill = mean(Catholic))))
+
 # vlines
 ggplot(test.data) + glyph(geom_point(aes(Fertility, Agriculture, 
   color = rank(Catholic)), size = 3), glyph.by = c("lat", "long"), 
   major = aes(mean(Fertility), mean(Education)), ref = ref_vline(aes(fill = 
   mean(Catholic))))
-  
-# points
-ggplot(mpg) + glyph(geom_bar(aes(x = trans, fill = year)), 
-  aes(x = mean(displ), y = mean(cty)), c("year"), y_scale = free, 
[…]
